### 2025-01-20 pre_print/pre_print.qmd
Plots updated, added model and prior eqs

---
articletitle: |
  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members.
# shorttitle: "Doe et al. (2022)" # usually auto-generated
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
author:
  - name: James Steele
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
  - name: James Fisher
    affiliations:
      - ref: 3
    orcid: 0000-0002-6013-8402
    email: james.fisher.phd@outlook.com
  - name: Tim Dettmann
    affiliations:
      - ref: 1
    email: tim.dettmann@gmail.com
affiliations:
      - id: 1
        name: Kieser Australia
      - id: 2
        name: Steele Research Limited
      - id: 3
        name: Department of Sport and Health, Solent University
abstract: |
  Place your abstract here. Use Markdown to style the abstract, e.g. write `*word*` to print a *word* in italics. You can use the Latex command `\newline\` to create a linebreak in the abstract. \newline\ *Explanation*: The second backslash ensures that styling can be applied directly after the linebreak. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [template, demo, exercise science] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: bibliography.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(patchwork)
library(tidyverse)
library(grid)
library(kableExtra)

targets::tar_config_set(store = here::here('_targets'))
```

# Introduction

Recent work has highlighted the importance of supervision in optimising strength outcomes from resistance training (RT). In a recent systematic review and meta-analysis from Fisher et al. [-@fisherRoleSupervisionResistance2022a] there was a moderate standardised mean effect (0.40 \[95%CI: 0.06, 0.74\]) of supervised vs unsupervised RT on strength outcomes synthesised from ten studies. Since then, a further study has reported that even in previously trained participants there may be benefit to strength outcomes from supervision [@colemanSupervisionResistanceTraining2023].

It has been speculated [@fisherRoleSupervisionResistance2022a] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort. Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022]. Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024]). Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.

As such it has also been suggested that a supervising trainer plays the role of providing motivation and enhancing trainee effort [@fisherRoleSupervisionResistance2022a]. Indeed, recent survey studies highlight that trainees perceive supervision to have an important role in determining their motivation and resultant effort during RT which they also perceive to be important to achieve their training goals [@fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024].

However, a recent meta-analysis of the dose-response relationship of proximity to momentary failure for strength and hypertrophy outcomes highlights that, while there is increased hypertrophy with closer proximity to failure, there is not a clear relationship for strength [@robinsonExploringDoseResponse2024]. But a caveat is that there was limited data for proximities to failure \>10 repetitions, and that the models were adjusted for load (average loads were typically \~75-85% of one repetition maximum). Thus, the results of this meta-analysis apply to proximity to failure *after* selecting load (i.e., intraset effort) and a higher relative load requires a greater effort all else being equal. When considering the average self-selected load of \~53% 1RM coupled with a typical \~10 repetitions per set equating to proximities to failure \>10 repetitions [@steeleAreTraineesLifting2022] it seems plausible that unsupervised trainees may train with a less-than-optimal load, repetitions, and resultant effort to optimise strength outcomes. Ultimately, and in combination, supervision should provide trainees with feedback of their past performance, whether technical, effort- or program-based, and guidance toward their future performance (Fisher, under review).

We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership. As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members we investigated experimentally the impact of supervision from a qualified exercise scientist.

[^1]: Note, Core members do have a "Review" session every twentieth session with an exercise scientist to complete strength testing, and review their training programme cards including selection of exercises and weights.

# Methods

```{r}
targets::tar_load(data)
```

## Experimental approach to the problem

The study was conducted at a selection of strength clinics operated by Kieser Australia recruiting from the existing pool of Core members at these locations. All Kieser Australia members are prescribed the same protocol. This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action). Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist. Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session. An acute randomised cross-over design was employed to examine participants completing an RT session with and without direct supervision with one week apart (i.e., a Core vs Assisted session). Participants were instructed to complete a single set of the resistance machine exercises on their current training programme card using their current training loads to momentary failure. The time under load (TUL) performed, and perceptions of effort and discomfort were recorded and compared between both conditions. This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.

In addition to the experimental study design noted above, we also took samples of Core and Assisted trainees from historical data records in Kieser Australia's database. We took separate random samples to obtain both TUL and session RPE data. There were two reasons for this. Firstly, we used this data to generate empirical informative prior distributions to use in our Bayesian modelling of the experimental study data. Secondly, they allowed us to also explore in the descriptive[^2] difference between Core and Assisted training *in situ*, but also that we could compare the prior distribution for the random sample of Core members *in situ* with the posterior distribution of Core members completing their Core session in the experimental study to understand the extent to which study effects might influence outcomes.

[^2]: Note that we do not claim this difference to be a causal effect in this data given that we have not adjusted for confounders, for example that different types of people might self-select into either Core or Assisted memberships and that this might confound comparisons of either TUL or session RPE.

## Participants

### Experimental

The study was advertised at five Kieser Australia strength clinics and we sought to recruit members who had \> 6 months prior training experience at the strength clinics, were healthy (no clinical conditions on their member record as recorded by a physiotherapist or exercise physiologist), both males and females (age +18 years), and without any current condition for which RT would be contraindicated. They were recruited from the existing client pool of Core members. We recruited a sample size based on considerations of what was logistically feasible given the availability of the exercise scientist staff at each clinic to conduct data collection and perform the supervised Assisted sessions and wanting to minimise the burden on day-to-day operation of the clinics. As such it was decided that a target of 50 participants across clinics would be acceptable. We ended up with ($k=$`r length(unique(data$id))`) participants recruited across the five clinics.

### Prior Sample

We queried the Kieser Australia database to generate random samples from historical data. We opted to generate reasonably large samples to ensure precise prior estimates, though did not use the entirety of the historical data (\~50000 members, \>10 years of data) so as to reduce computation time for Bayesian modelling. For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise. For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.

[^3]: Whilst clients can have Core or Assisted memberships the former also have sessions that are supervised (i.e., their "Review" sessions noted in footnote$^1$) and the latter can have unsupervised sessions (i.e., they can attend the clinic outside of their scheduled sessions with the exercise scientist and train unsupervised). Thus we limited our sample to only normal training sessions which had been either unsupervised or supervised for both Core and Assisted members respectively.

## Protocols

All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland). For the experimental study participants were instructed to attend two training sessions at least one-week apart where they completed either their current Core session (i.e., unsupervised) or an Assisted session (i.e., supervised). During the Core session the participant completed their training session as prescribed with the only difference for the Assisted session being that an exercise scientist supervised them providing instruction and motivation, however, they did not interfere in any physical way during the set (e.g. to spot or assist in completing a failing repetition). Each participant had an existing training programme card. In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card prior to participation in the study. Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the time-under-load achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL. Momentary failure was defined as per Steele et al. [-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was communicated to the participants. During each session for each exercise the trainees/trainers recorded their time-under-load achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose. In addition, they recorded their perceptions of effort and discomfort in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).

The data generated from the sample of historical members was generated according to the standard protocol described above that all Kieser Australia members are prescribed. TUL was also recorded similarly using timers available about the clinic, r for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which has placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application. Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale with...

## Statistical Analysis

```         
    UPDATE LINKS!!!!!!!!
```

All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/dqwh5/> or the corresponding GitHub repository <https://github.com/jamessteeleii/self_talk_meta_analysis_update>. We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://osf.io/ftajc>.

As noted, the project was not pre-registered but involved exploratory analysis of the experimental and prior datasets. All analyses have been conducted within a Bayesian framework and all posterior estimates and their precision, along with conclusions based upon them, are interpreted continuously and probabilistically, considering priors, data quality, and all within the context of each outcome and the assumptions of the model employed as the estimator [@kruschkeBayesianNewStatistics2018]. All models were run with 2000 warmup and 2000 sampling iterations and four Monte Carlo Markov Chains. Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions. We fit two sets of models for the TUL and RPE outcomes, one on the prior sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the rating of perceived discomfort outcome using weakly regularising priors. These models are described below. For all models we calculated average marginal effects as the global grand means for both predictions under each condition as well as contrasts between conditions and visualise both prior and posterior distributions along with mean and 95% quantile estimates from these.

### Time Under Load Analysis

Upon inspecting the raw distribution of the prior sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (ADD LINK). Anecdotal reports from Kieser Australia staff had also prior to exploring the data suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure. We did not however expect this to be the same for the prior sample of Assisted members. As such, it suggested to us that there might be two processes underlying the observed data: a tendency to target a specific TUL threshold and to stop the exercise at that point, or to otherwise continue the exercise to momentary failure (or to stop the exercise based on some other process such as tolerable proximity to failure). Thus, we opted to use a hurdle type model comprising a Bernoulli distribution for the probability of having a TUL of 120 seconds, and a student $t$ distribution for all other values of TUL[^4]. We expected this may also be the case in our experimental dataset given we recruited Core members and so utilised the same type of model for this. As such this allowed us to examine both the expectations of the global grand mean for TUL, in addition to the probabilities of having a TUL of 120 seconds.

[^4]: A custom family for the hurdle-student $t$ model was produced for use with the `brms` R package, adapted from the hurdle-normal distribution developed by Heiss [-@heissGuideModelingOutcomes].

For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept. The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used (note, due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset). These fixed and random effects were modelled as predictors for both the hurdle and student $t$ components of the model. TUL was centred at zero for modelling (i.e., all values of TUL$_i$ has 120 subtracted from them prior to modelling). Formally the model(s) were as follows:

$$
\begin{aligned}
& \textbf{Hurdle model of}\ i^{th}\ \textbf{TUL across condition}\ c\  \textbf{within}\ j^{th}\ \textbf{location,}\ k^{th}\ \textbf{member, and}\ l^{th} \textbf{ machine} \\
\text{TUL}_{ic} &\sim \operatorname{Hurdle\,log-student}t(\pi_{ic_{j[i],k[i],l[i]}}, \nu_y, \mu_{ic_{j[i],k[i],l[i]}}, \sigma_y) \ ...\ \text{or alternatively,} \\
\text{TUL}_{ic} &\sim
\begin{Bmatrix}
0 & \text{with probability } \pi_{ic_{j[i],k[i],l[i]}} \\[4pt]
\operatorname{Student}t(\mu_{ic_{j[i],k[i],l[i]}}, \nu_y, \sigma_y) & \text{with probability } 1 - \pi_{ic_{j[i],k[i],l[i]}}
\end{Bmatrix}
\\
\\
& \textbf{Models for distribution parameters} \\
\operatorname{logit}(\pi_{ic}) &= (\gamma_0 + \gamma_{0_{j[i],k[i],l[i]}}) + \gamma_1 \text{Condition}_{ic} & \text{120 seconds/not-120 seconds process} \\[4pt]
\mu_{ic} &= (\beta_0 + b_{0_{j[i],k[i],l[i]}}) + \beta_1\text{Condition}_{ic} & \text{Location parameter in student}\ t\ \text{process} \\[4pt]
\gamma_{0_j} &\sim \mathcal{N}(0, \sigma_{\gamma_{0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in hurdle model} \\[4pt]
\gamma_{0_k} &\sim \mathcal{N}(0, \sigma_{\gamma_{0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in hurdle model} \\[4pt]
\gamma_{0_l}, &\sim \mathcal{N}(0, \sigma_{\gamma_{0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in hurdle model} \\[4pt]
b_{0_j} &\sim \mathcal{N}(0, \sigma_{b_{0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in student}\ t\ \text{model model} \\[4pt]
b_{0_k} &\sim \mathcal{N}(0, \sigma_{b_{0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in student}\ t\ \text{model model} \\[4pt]
b_{0_l} &\sim \mathcal{N}(0, \sigma_{b_{0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in student}\ t\ \text{model model} \\[4pt]
\\
\end{aligned}
$$ {#eq-hurdle-model}
Weakly regularising priors were employed for prior sample dataset. These were the default priors in the `brms` package for all intercept terms which are scaled to the expected response value when all predictors are at their means and use a student $t$ distribution with degrees of freedom $\nu=3$, and variance terms similarly scaled with $\mu=0$. The only change to this was to set reasonable upper and lower bounds for the prior condition effects to limit prediction of impossible TUL values. The priors were as follows:

$$
\begin{aligned}
& \textbf{Priors (prior sample model)} \\
\gamma_0 &\sim \text{Logistic}(2, 0.1) & \text{Prior for intercept in hurdle model i.e., Assisted condition} \\
\beta_0 &\sim \text{Student}\ t(3, 3, 13.3) & \text{Prior for intercept in student}\ t\ \text{model i.e., Assisted condition} \\
\gamma_1,\beta_1 &\sim \text{Student}\ t(3, 0, 15, \text{lb}=-60, \text{lb}=60) & \text{Prior for Condition effects in both models i.e., Core minus Assisted condition} \\
\sigma_y,\sigma_{\gamma_{0_{j}}},\sigma_{\gamma_{0_{k}}},\sigma_{\gamma_{0_{l}}},\sigma_{b_{0_{j}}},\sigma_{b_{0_{k}}},\sigma_{b_{0_{l}}} \ &\sim \text{Student}\ t(3, 0, 13.3, \text{lb}=0) & \text{Prior for all variability parameters} \\
\nu_y &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
\end{aligned}
$$ {#eq-hurdle-prior}

We then took posterior parameter estimates from the prior sample model and used these to generate informative priors for the experimental data models which were as follows:

$$
\begin{aligned}
& \textbf{Priors (experimental data model)} \\
\gamma_0 &\sim \text{Student}\ t(3.28, -3.03, 0.12) & \text{Prior for intercept in hurdle model i.e., Assisted condition} \\
\beta_0 &\sim \text{Student}\ t(3.28, 2.79, 1.30) & \text{Prior for intercept in student}\ t\ \text{model i.e., Assisted condition} \\
\gamma_1 &\sim \text{Student}\ t(3.28, -0.19, 0.11) & \text{Prior for Condition effects in hurdle model i.e., Core minus Assisted condition} \\
\beta_1 &\sim \text{Student}\ t(3.28, 0.07, 0.58) & \text{Prior for Condition effects in student}\ t\ \text{model i.e., Core minus Assisted condition} \\
\sigma_y &\sim \text{Student}\ t(3, 10.59, 0.19, \text{lb}=0) & \text{Prior for residual variability parameter} \\
\sigma_{\gamma_{0_{j}}} &\sim \text{Student}\ t(3, 0.31, 0.08, \text{lb}=0) & \text{Prior for location variability parameter in hurdle model} \\
\sigma_{\gamma_{0_{k}}} &\sim \text{Student}\ t(3, 1.16, 0.07, \text{lb}=0) & \text{Prior for member variability parameter in hurdle model} \\
\sigma_{\gamma_{0_{l}}} &\sim \text{Student}\ t(3, 0.15, 0.08, \text{lb}=0) & \text{Prior for machine variability parameter in hurdle model} \\
\sigma_{b_{0_{j}}} &\sim \text{Student}\ t(3, 3.43, 0.62, \text{lb}=0) & \text{Prior for location variability parameter in student}\ t\ \text{model} \\
\sigma_{b_{0_{k}}} &\sim \text{Student}\ t(3, 10.2, 0.23, \text{lb}=0) & \text{Prior for member variability parameter in student}\ t\ \text{model} \\
\sigma_{b_{0_{l}}} &\sim \text{Student}\ t(3, 3.72, 0.93, \text{lb}=0) & \text{Prior for machine variability parameter in student}\ t\ \text{model} \\
\nu_y &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
\end{aligned}
$$ {#eq-hurdle-prior-exp}

### Rating of Perceived Effort and Discomfort Analysis

Despite the use of two different scales and protocols for data collection of perception of effort in our samples i.e., a session RPE using the 6-20 Borg scale for the prior sample and immediately post each exercise RPE using a 0-10 scale, both scales were anchored at their limits as in essence no effort (6 or 0) or maximal effort (20 or 10). As such, and as the intention was to utilise the prior sample data to assist in producing empirical prior distributions for our experimental data models, we opted to rescale both outcomes to lie on the $(0,1)$ interval such that they reflected the percent effort that was reported as perceived [@steeleWhatPerceptionEffort2020]. We then employed an ordered beta regression model [@kubinecOrderedBetaRegression2022] which employs a cutpoint process similar to ordered logistic models in order to model both the continuous responses on the $(0,1)$ interval reflecting some effort though not maximal, and the degenerate response on the bounds $[0,1]$ reflecting both no and maximal effort respectively. We approached the rating of perceived discomfort similarly as this scale was also anchored between no discomfort and the maximum imaginable.

For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept. The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used for the experimental data model only (note, due to the session RPE reported in the prior sample data and that we sampled only one session we do not have member or machine random effects, and due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset). These fixed and random effects were modelled as predictors for both the mean and precision parameters of the ordered beta regression model. As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval. Formally the model(s) were as follows (note that the model for the prior dataset omits the random effect for machine):

$$
\begin{aligned}
& \textbf{Ordered beta regression model of}\ i^{th}\ \textbf{RPE across condition}\ c\  \textbf{within}\ j^{th}\ \textbf{location,}\ k^{th}\ \textbf{member, and}\ l^{th} \textbf{ machine} \\
\text{RPE}_{ic} &\sim
\begin{Bmatrix}
0 & \text{with probability } \alpha_{ic_{j[i],k[i],l[i]}} \\[4pt]
\in (0,1) & \text{with probability } \delta_{ic_{j[i],k[i],l[i]}} \\[4pt]
1 & \text{with probability } \gamma_{ic_{j[i],k[i],l[i]}} \\[4pt]
\end{Bmatrix}
\\
\\
\text{logit}(\alpha_{ic}) &= 1-(X'\beta-k_1) & \text{Probability of obtaining a 0} \\[4pt]
\text{logit}(\delta_{ic}) &= \bigr[(X'\beta-k_1) - (X'\beta-k_2)\bigr]\text{Beta}(X'\beta, X'
\beta_\phi) & \text{Probability of obtaining a value between 0 and 1} \\[4pt]
\text{logit}(\gamma_{ic}) &= (X'\beta-k_2) & \text{Probability of obtaining a 1} \\[4pt]
\\
X'\beta&=(\beta_0 + b_{0_{j[i],k[i],l[i]}}) + \beta_1 \text{Condition}_{ic} & \text{Vector of predictors} \\[4pt] 
b_{0_j} &\sim \mathcal{N}(0, \sigma_{b_{0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in student}\ t\ \text{model model} \\[4pt]
b_{0_k} &\sim \mathcal{N}(0, \sigma_{b_{0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in student}\ t\ \text{model model} \\[4pt]
b_{0_l} &\sim \mathcal{N}(0, \sigma_{b_{0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in student}\ t\ \text{model model} \\[4pt]
X'\beta_\phi&=(\beta_{\phi0} + b_{\phi0_{j[i],k[i],l[i]}}) + \beta_{\phi1} \text{Condition}_{ic} & \text{Vector of predictors} \\[4pt] 
b_{\phi0_j} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in student}\ t\ \text{model model} \\[4pt]
b_{\phi0_k} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in student}\ t\ \text{model model} \\[4pt]
b_{\phi0_l} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in student}\ t\ \text{model model} \\[4pt]
\\
\end{aligned}
$$ {#eq-ordbeta-model}

Weakly regularising priors were employed for prior sample dataset. These were the default priors in the `ordbetareg` package for all intercept terms which are scaled to the expected response value when all predictors are at their means and use a student $t$ distribution with degrees of freedom $\nu=3$, and variance terms similarly scaled with $\mu=0$. We left the flat prior on the fixed effect for condition given we were primarily interested in letting the prior dataset speak for itself in generating a posterior distribution to use as prior in the experimental data model. As noted, the prior sample model only included the random effects for location. The priors were as follows:

$$
\begin{aligned}
& \textbf{Priors (prior sample model)} \\
\beta_0 &\sim \text{Student}\ t(3, 0.6, 2.5) & \text{Prior for intercept of mean parameter i.e., Assisted condition} \\
\beta_{\phi0} &\sim \text{Student}\ t(3, 0, 2.5) & \text{Prior for intercept of precision parameter i.e., Assisted condition} \\
\beta_1,\beta_{\phi1} &\sim \text{Uniform}(\text{lb}=-\infty,u=\infty) & \text{Prior for Condition effects in both location and precision parameters i.e., Core minus Assisted condition} \\
\sigma_{b_{0_{j}}},\sigma_{b_{\phi0_{j}}} \ &\sim \text{Student}\ t(3, 0, 2.5, \text{lb}=0) & \text{Prior for location variability parameters} \\
k_1,k_2 &\sim \text{Induced dirichlet}(1,1,1) & \text{Prior for cutpoint probabilities} \\
\end{aligned}
$$ {#eq-ordbeta-prior}

We then took posterior parameter estimates from the prior sample model and used these to generate informative priors for the experimental data model which were as follows:

$$
\begin{aligned}
& \textbf{Priors (experimental sample model)} \\
\beta_0 &\sim \text{Student}\ t(3, 0.35, 0.05) & \text{Prior for intercept of mean parameter i.e., Assisted condition} \\
\beta_{\phi0} &\sim \text{Student}\ t(3, 2.01, 0.07) & \text{Prior for intercept of precision parameter i.e., Assisted condition} \\
\beta_1 &\sim \text{Student}\ t(3, -0.24, 0.04) & \text{Prior for Condition effects in mean parameter i.e., Core minus Assisted condition} \\
\beta_{\phi1} &\sim \text{Student}\ t(3, -0.48, 0.06) & \text{Prior for Condition effects in precision parameter i.e., Core minus Assisted condition} \\
\sigma_{b_{0_{j}}} \ &\sim \text{Student}\ t(3, 0.19, 0.04, \text{lb}=0) & \text{Prior for location variability parameters} \\
\sigma_{b_{\phi0_{j}}} \ &\sim \text{Student}\ t(3, 0.26, 0.05, \text{lb}=0) & \text{Prior for all other variability parameters} \\
\sigma_{b_{0_{k}}},\sigma_{b_{0_{l}}},\sigma_{b_{\phi0_{k}}},\sigma_{b_{\phi0_{l}}} \ &\sim \text{Student}\ t(3, 0, 2.5, \text{lb}=0) & \text{Prior for all variability parameters} \\
k_1,k_2 &\sim \text{Induced dirichlet}(1,1,1) & \text{Prior for cutpoint probabilities} \\
\end{aligned}
$$ {#eq-ordbeta-prior-exp}

We did not have prior sample data for rating of perceived discomfort and so for this outcome we used the same model as above for the RPE outcomes but with wholly default weakly regularising priors.

# Results

ADD NO OBS FOR DATASETS

## Time Under Load

## Rating of Perceived Effort and Discomfort

# Contributions

Authors should report the contributions of each author in the a specific contribution section based on the guidelines set forth by the International Committee of Medical Journal Editors.

Please indicate author contributions as clearly as possible, according to the following criteria:

-   Substantial contributions to conception and design
-   Acquisition of data
-   Analysis and interpretation of data
-   Drafting the article or revising it critically for important intellectual content
-   Final approval of the version to be published

# Acknowledgements

People who contributed to the work but do not fit our author criteria should be listed in the acknowledgements, along with their contributions. You must ensure that anyone named in the acknowledgements agrees to being so named.

Funding sources should not be included in the acknowledgements, but in the section below.

# Funding Information

Please provide a list of the sources of funding, as well as the relevant grant numbers, where possible. List the authors associated with specific funding sources. You will also enter this information in a form during the submission process, but it must be repeated here.

# Data and Supplementary Material Accessibility

This should list the database(s) and, if appropriate, the respective accession numbers and DOIs for all data or supplementary material for the manuscript that has been made publicly available on a trusted digital repository. If no data, code, or supplementary material are available for this manuscript then the reason for this should be explained here.

### 2025-01-20 R/functions.R

@@ -70,10 +70,10 @@ fit_model_prior_sample_tul <-
       ),
       family = hurdle_student_t,
       stanvars = stan_vars,
-      # prior = c(
-      #   # set wide but constrained priors for just the fixed student t effects to realistic range for posterior predictions
-      #   set_prior("student_t(3, 0, 15)", class = "b", lb = -60, ub = 60)
-      # ),
+      prior = c(
+        # set wide but constrained priors for just the fixed effects
+        set_prior("student_t(3, 0, 15)", class = "b", lb = -60, ub = 60)
+      ),
       data = data,
       chains = 4,
       cores = 4,
@@ -579,6 +579,7 @@ make_plot_prior_sample_tul <- function(data) {
                linewidth = 0.25
     ) +
     geom_histogram(binwidth = 1, alpha = 0.5) +
+    scale_x_continuous(breaks = seq(60,180,10)) +
     facet_grid(core_assisted ~ .) +
     labs(x = "Time Under Load (seconds)") +
     theme_classic(base_size = 8) +
@@ -614,6 +615,7 @@ make_plot_data_tul <- function(data) {
       alpha = 0.5,
       linewidth = 0.25
     ) + 
+    scale_y_continuous(breaks = seq(60,600,30)) +
     labs(x = "Condition",
          y = "Time Under Load (Seconds)") +
     theme_classic(base_size = 8)
@@ -742,7 +744,7 @@ make_plot_combined_data <- function(plot1, plot2, plot3, plot4, plot5) {
   ) +
     plot_annotation(title = "Current Experimental Data Distributions",
                     subtitle = "Time Under Load, Rating of Perceived Effort, and Rating of Perceived Discomfort") +
-    plot_layout(axes = "collect")
+    plot_layout(axes = "collect_x")
   
   
   wrap_elements(prior_plots) / wrap_elements(data_plots) 
@@ -787,7 +789,7 @@ make_plot_preds_tul <- function(pred_draws) {
     ) +
     scale_color_manual(values = c("#56B4E9", "#E69F00"), limits = rev) +
     scale_fill_manual(values = c("#56B4E9", "#E69F00"), limits = rev) +
-    scale_x_continuous(limits = c(60, 180)) +
+    scale_x_continuous(limits = c(60, 180), breaks = seq(60,180,10)) +
     labs(
       y = "Condition",
       x = "Time Under Load (Seconds)",

### 2025-01-20 _targets.R

@@ -240,5 +240,42 @@ list(
     plot_combined_rpe_discomfort,
     make_plot_combined_rpe_discomfort(plot_preds_rpe, plot_contrasts_rpe,
                            plot_preds_discomfort, plot_contrasts_discomfort)
+  ),
+  
+  # Make plots into tiffs
+  tar_target(
+    plot_combined_data_tiff,
+    ggsave(
+      "plots/plot_combined_data.tiff", 
+      plot_combined_data, 
+      device = "tiff",
+      dpi = 300,
+      w = 7.5,
+      h = 7.5
+    )
+  ),
+  
+  tar_target(
+    plot_combined_tul_tiff,
+    ggsave(
+      "plots/plot_combined_tul.tiff", 
+      plot_combined_tul, 
+      device = "tiff",
+      dpi = 300,
+      w = 7.5,
+      h = 7
+    )
+  ),
+  
+  tar_target(
+    plot_combined_rpe_discomfort_tiff,
+    ggsave(
+      "plots/plot_combined_rpe_discomfort.tiff", 
+      plot_combined_rpe_discomfort, 
+      device = "tiff",
+      dpi = 300,
+      w = 7.5,
+      h = 7
+    )
   )
 )

### 2025-01-21 pre_print/pre_print.qmd
Analysis and results sections finished

@@
        \usepackage[font=scriptsize]{caption}
+       \usepackage{caption}
 author:
@@
 library(tidyverse)
-library(grid)
-library(kableExtra)
+library(marginaleffects)
+library(ggdist)
 
 targets::tar_config_set(store = here::here('_targets'))
+
+# posterior predict functions for hurdle model
+posterior_predict_hurdle_student_t <- function(i, prep, ...) {
+  nu <- brms::get_dpar(prep, "nu", i = i)
+  mu <- brms::get_dpar(prep, "mu", i = i)
+  sigma <- brms::get_dpar(prep, "sigma", i = i)
+  theta <- brms::get_dpar(prep, "hu", i = i)
+
+  hu <- runif(prep$ndraws, 0, 1)
+  ifelse(hu < theta, 0, brms::rstudent_t(prep$ndraws, nu, mu, sigma))
+}
+
+posterior_epred_hurdle_student_t <- function(prep) {
+  with(prep$dpars, mu * (1 - hu))
+}
 ```
@@
 targets::tar_load(data)
+targets::tar_load(prior_data_tul)
+targets::tar_load(prior_data_rpe)
+
 ```
@@
 As such it was decided that a target of 50 participants across clinics would be acceptable.
-We ended up with ($k=$`r length(unique(data$id))`) participants recruited across the five clinics.
+We ended up with n = `r length(unique(data$id))` participants recruited across the five clinics.
 
@@
 For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.
+Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`.
 For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.
@@
 TUL was also recorded similarly using timers available about the clinic, r for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which has placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.
-Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale with...
+Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale with (where the scale is displayed as scrollable and the verbal anchors are "6 - None", "7- Very, very light", "13 – Somewhat hard", "19 - Very, very hard", and "20 – Maximal Exertion"
+)
 
@@
 These models are described below.
-For all models we calculated average marginal effects as the global grand means for both predictions under each condition as well as contrasts between conditions and visualise both prior and posterior distributions along with mean and 95% quantile estimates from these.
+For all models we calculated average marginal effects as the global grand means for both predictions under each condition as well as contrasts between conditions and visualise both prior and posterior distributions along with median and 95% quantile interval (QI) estimates from these.
 
@@
 
-Upon inspecting the raw distribution of the prior sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (ADD LINK).
+Upon inspecting the raw distribution of the prior sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (see @fig-raw).
 Anecdotal reports from Kieser Australia staff had also prior to exploring the data suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.
@@
 TUL was centred at zero for modelling (i.e., all values of TUL$_i$ has 120 subtracted from them prior to modelling).
-Formally the model(s) were as follows:
+Formally the model(s) for $i^{th}$ TUL across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows:
 
 $$
+\tiny
 \begin{aligned}
-& \textbf{Hurdle model of}\ i^{th}\ \textbf{TUL across condition}\ c\  \textbf{within}\ j^{th}\ \textbf{location,}\ k^{th}\ \textbf{member, and}\ l^{th} \textbf{ machine} \\
 \text{TUL}_{ic} &\sim \operatorname{Hurdle\,log-student}t(\pi_{ic_{j[i],k[i],l[i]}}, \nu_y, \mu_{ic_{j[i],k[i],l[i]}}, \sigma_y) \ ...\ \text{or alternatively,} \\
@@
 $$
+\tiny
 \begin{aligned}
-& \textbf{Priors (prior sample model)} \\
+& \textbf{Priors (prior sample data model)} \\
 \gamma_0 &\sim \text{Logistic}(2, 0.1) & \text{Prior for intercept in hurdle model i.e., Assisted condition} \\
@@
 $$
+\tiny
 \begin{aligned}
-& \textbf{Priors (experimental data model)} \\
+& \textbf{Priors (experimental sample data model)} \\
 \gamma_0 &\sim \text{Student}\ t(3.28, -3.03, 0.12) & \text{Prior for intercept in hurdle model i.e., Assisted condition} \\
@@
 As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval.
-Formally the model(s) were as follows (note that the model for the prior dataset omits the random effect for machine):
+Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows:were as follows (note that the model for the prior dataset omits the random effect for machine):
 
 $$
+\tiny
 \begin{aligned}
-& \textbf{Ordered beta regression model of}\ i^{th}\ \textbf{RPE across condition}\ c\  \textbf{within}\ j^{th}\ \textbf{location,}\ k^{th}\ \textbf{member, and}\ l^{th} \textbf{ machine} \\
 \text{RPE}_{ic} &\sim
@@
 $$
+\tiny
 \begin{aligned}
-& \textbf{Priors (prior sample model)} \\
+&\textbf{Priors (prior sample data model)} \\
 \beta_0 &\sim \text{Student}\ t(3, 0.6, 2.5) & \text{Prior for intercept of mean parameter i.e., Assisted condition} \\
@@
 $$
+\tiny
 \begin{aligned}
-& \textbf{Priors (experimental sample model)} \\
+&\textbf{Priors (experimental sample data model)} \\
 \beta_0 &\sim \text{Student}\ t(3, 0.35, 0.05) & \text{Prior for intercept of mean parameter i.e., Assisted condition} \\
@@
 # Results
+In the experimental dataset we ended up with `r nrow(data)/2` observations for each outcome across for each of the two conditions.
+Our prior sample for TUL after filtering to the same sample of machines came from `r length(unique(prior_data_tul$id))` members (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`) from `r length(unique(prior_data_tul$location))` locations encompassing `r nrow(prior_data_tul)` observations.
+The distributions of raw data from both the prior sample and experimental datasets can be seen in @fig-raw.
+```{r}
+#| label: fig-raw
+#| fig-width: 7.5
+#| fig-height: 7
+#| fig-cap: The top two panels show the distributions as histograms of both time under load and session rating of perceived effort in the prior sample of data, and the bottom three panels show the paired responses for time under load in addition to the paired responses and histograms for the rating of perceived effort and discomfort in the current experimental data (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel and horizontal dashed lines in the bottom left panel).
 
-ADD NO OBS FOR DATASETS
+targets::tar_load(plot_combined_data)
+
+plot_combined_data
+
+```
 
 ## Time Under Load
+```{r}
+targets::tar_load(model_prior_sample_tul)
+targets::tar_load(model_tul)
+
+preds_prior_tul <- avg_predictions(model_prior_sample_tul,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+preds_tul <- avg_predictions(model_tul,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+contrast_prior_tul <- avg_comparisons(model_prior_sample_tul,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+contrast_tul <- avg_comparisons(model_tul,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+preds_prior_tul_hu <- avg_predictions(model_prior_sample_tul,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "hu")
+
+preds_tul_hu <- avg_predictions(model_tul,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "hu")
+
+contrast_prior_tul_hu <- avg_comparisons(model_prior_sample_tul,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "hu")
+
+contrast_tul_hu <- avg_comparisons(model_tul,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "hu")
+
+```
+In our prior sample model there was little difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds.
+The estimated TUL for Core sessions was `r preds_prior_tul$estimate[2]+120` [95%QI: `r preds_prior_tul$conf.low[2]+120`, `r preds_prior_tul$conf.high[2]+120`] seconds, for Assisted sessions was `r preds_prior_tul$estimate[1]+120` [95%QI: `r preds_prior_tul$conf.low[1]+120`, `r preds_prior_tul$conf.high[1]+120`] seconds, and the between condition contrast (Core minus Assisted) was `r contrast_prior_tul$estimate` [95%QI: `r contrast_prior_tul$conf.low`, `r contrast_prior_tul$conf.high`] seconds.
+The probability of stopping at 120 seconds for Core sessions was `r scales::percent(preds_prior_tul_hu$estimate[2])` [95%QI: `r scales::percent(preds_prior_tul_hu$conf.low[2])`, `r scales::percent(preds_prior_tul_hu$conf.high[2])`], for Assisted sessions was `r scales::percent(preds_prior_tul_hu$estimate[1])` [95%QI: `r scales::percent(preds_prior_tul_hu$conf.low[1])`, `r scales::percent(preds_prior_tul_hu$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_tul_hu$estimate)` [95%QI: `r scales::percent(contrast_prior_tul_hu$conf.low)`, `r scales::percent(contrast_prior_tul_hu$conf.high)`].
+
+Our experimental sample model showed far greater difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds, in addition to the posterior estimates differing considerably from the prior sample estimates in terms of between condition contrasts for TUL and probability of stopping t 120 seconds, Assisted condition TUL, and both conditions probability of stopping at 120 seconds.
+The estimated TUL for Core sessions was `r preds_tul$estimate[2]+120` [95%QI: `r preds_tul$conf.low[2]+120`, `r preds_tul$conf.high[2]+120`] seconds, for Assisted sessions was `r preds_tul$estimate[1]+120` [95%QI: `r preds_tul$conf.low[1]+120`, `r preds_tul$conf.high[1]+120`] seconds, and the between condition contrast (Core minus Assisted) was `r contrast_tul$estimate` [95%QI: `r contrast_tul$conf.low`, `r contrast_tul$conf.high`] seconds.
+The probability of stopping at 120 seconds for Core sessions was `r scales::percent(preds_tul_hu$estimate[2])` [95%QI: `r scales::percent(preds_tul_hu$conf.low[2])`, `r scales::percent(preds_tul_hu$conf.high[2])`], for Assisted sessions was `r scales::percent(preds_tul_hu$estimate[1])` [95%QI: `r scales::percent(preds_tul_hu$conf.low[1])`, `r scales::percent(preds_tul_hu$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_tul_hu$estimate)` [95%QI: `r scales::percent(contrast_tul_hu$conf.low)`, `r scales::percent(contrast_tul_hu$conf.high)`].
+
+All marginal predictions (i.e., global grand means) for predictions and contrasts can be seen in @fig-tul.
+
+```{r}
+#| label: fig-tul
+#| fig-width: 7.5
+#| fig-height: 7
+#| fig-cap: The top two panels show the prior (i.e., generated from the prior sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of time under load (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel) and the bottom panels show the probabilities of stopping at 120 seconds.
+
+targets::tar_load(plot_combined_tul)
+
+plot_combined_tul
+
+```
 
 ## Rating of Perceived Effort and Discomfort
+```{r}
+targets::tar_load(model_prior_sample_rpe)
+targets::tar_load(model_rpe)
+targets::tar_load(model_prior_sample_discomfort)
+targets::tar_load(model_discomfort)
+
+preds_prior_rpe <- avg_predictions(model_prior_sample_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+preds_rpe <- avg_predictions(model_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+contrast_prior_rpe <- avg_comparisons(model_prior_sample_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+contrast_rpe <- avg_comparisons(model_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+preds_discomfort <- avg_predictions(model_discomfort,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+contrast_discomfort <- avg_comparisons(model_discomfort,
+                               variables = "core_assisted",
+                               re_formula = NA)
+
+```
+In our prior sample model there was a small difference between Core and Assisted sessions in RPE.
+The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_rpe$estimate)` [95%QI: `r scales::percent(contrast_prior_rpe$conf.low)`, `r scales::percent(contrast_prior_rpe$conf.high)`].
+
+Our experimental sample model showed a similar magnitude of difference between Core and Assisted sessions in RPE, though the posterior estimates for RPEs were considerably greater compared with the prior sample estimates.
+The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_rpe$estimate)` [95%QI: `r scales::percent(contrast_rpe$conf.low)`, `r scales::percent(contrast_rpe$conf.high)`].
+
+The rating of perceived discomfort results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions.
+The estimated rating of perceived discomfort as a percentage for Core sessions was `r scales::percent(preds_discomfort$estimate[2])` [95%QI: `r scales::percent(preds_discomfort$conf.low[2])`, `r scales::percent(preds_discomfort$conf.high[2])`], for Assisted sessions `r scales::percent(preds_discomfort$estimate[1])` [95%QI: `r scales::percent(preds_discomfort$conf.low[1])`, `r scales::percent(preds_discomfort$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_discomfort$estimate)` [95%QI: `r scales::percent(contrast_discomfort$conf.low)`, `r scales::percent(contrast_discomfort$conf.high)`].
+
+All marginal predictions (i.e., global grand means) for predictions and contrasts can be seen in @fig-rpe.
+
+```{r}
+#| label: fig-rpe
+#| fig-width: 7.5
+#| fig-height: 7
+#| fig-cap: The top two panels show the prior (i.e., generated from the prior sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of rating of perceived effort on the percentage scale, and the bottom panels show posterior disributions only for rating of perceived discomfort transformed back to the arbitrary unit scale (0,10).
+
+targets::tar_load(plot_combined_rpe_discomfort)
+
+plot_combined_rpe_discomfort
+
+```
+
+# Discussion
+
+ADD DISCUSSION
 

### 2025-01-21 R/functions.R

@@ -767,7 +767,7 @@ make_plot_preds_tul <- function(pred_draws) {
         fill = distribution
       ),
       data = filter(pred_draws, core_assisted == "Core"),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -781,7 +781,7 @@ make_plot_preds_tul <- function(pred_draws) {
         fill = distribution
       ),
       data = filter(pred_draws, core_assisted == "Assisted"),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -814,7 +814,7 @@ make_plot_contrasts_tul <- function(contrast_draws) {
         color = distribution,
         fill = distribution
       ),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -851,7 +851,7 @@ make_plot_preds_tul_hu <- function(pred_draws) {
         fill = distribution
       ),
       data = filter(pred_draws, core_assisted == "Core"),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -865,7 +865,7 @@ make_plot_preds_tul_hu <- function(pred_draws) {
         fill = distribution
       ),
       data = filter(pred_draws, core_assisted == "Assisted"),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -898,7 +898,7 @@ make_plot_contrasts_tul_hu <- function(contrast_draws) {
         color = distribution,
         fill = distribution
       ),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
       position = position_dodge(w = -0.1),
@@ -956,7 +956,7 @@ make_plot_preds_rpe <- function(pred_draws) {
         fill = distribution
       ),
       data = filter(pred_draws, core_assisted == "Core"),
-      point_interval = "mean_qi",
+      point_interval = "median_qi",
       .width = .95,
       slab_alpha = 0.5,
[…]

### 2025-01-21 pre_print/pre_print.qmd
First draft preprint

@@
 articletitle: |
-  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members.
+  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
 # shorttitle: "Doe et al. (2022)" # usually auto-generated
@@
       - id: 1
-        name: Kieser Australia
+        name: Kieser Australia, Melbourne, Australia
       - id: 2
-        name: Steele Research Limited
+        name: Steele Research Limited, Eastleigh, UK
       - id: 3
-        name: Department of Sport and Health, Solent University
+        name: Department of Sport and Health, Solent University, Southampton, UK
+      - id: 4
+        name: The Exercise Coach, Tennessee, USA
 abstract: |
-  Place your abstract here.
-Use Markdown to style the abstract, e.g.
-write `*word*` to print a *word* in italics.
-You can use the Latex command `\newline\` to create a linebreak in the abstract.
-\newline\ *Explanation*: The second backslash ensures that styling can be applied directly after the linebreak.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
-This document is only a demo explaining how to use the template.
+  Recent work has suggested that supervision during resistance training might enhance outcomes such as strength adaptation due to role that it plays in optimising trainee effort.
+We were afforded an opportunity to examine the role of supervision on effort during resistance training in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership or supervised by a qualified exercise scientist on an "Assisted" membership.
+Both retrospectively through examination of samples of member training records, and through a prospective experimental study, we examined the effects of supervision upon exercise performance as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort.
+We adopted Bayesian methods for modelling and from the prior sample of retrospective data generated empirically informed posterior distributions to use as prior distributions in our modelling of the experimental study.
+The prior sample consisted of training sessions from 1000 each of Core and Assisted members, whilst the experimental study consisted of Core members recruited to in randomised order complete an unsupervised and a supervised session using their current training loads where they were instructed to perform each exercise to momentary failure.
+Our findings suggest that, *in situ* under real world settings (i.e., our prior sample of members) there was little difference in exercise performance, yet in our experimental study there was a clear benefit to TUL performance when under supervision (Core = 125.12 [95%QI: 113.70, 131.90] seconds; Assisted = 147.35 [95%QI: 134.29, 154.81] seconds; Core minus Assisted contrast = -22.10 [95%QI: -26.60, -17.61] seconds).
+There was a small difference in RPE in percentage points reported with and without supervision under both real world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; Core minus Assisted contrast = -6% [95%QI: -8%, -4%]) and experimental conditions (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; Core minus Assisted contrast = -6% [95%QI: -10%, -4%]) suggesting a that under supervision trainees train with greater proximity to failure, which was also supported by greater rating of perceived discomfort under supervision during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3] pts; Assisted = 7.5 [95%QI: 6.5, 8.3] pts; Core minus Assisted contrast = -1.2 [95%QI: -1.6, -0.9] pts).
+These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
 license-type: ccby # change if neccessary
@@
 Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
-Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024]).
+Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024].
 Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.
@@
 
-We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership.
-As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members we investigated experimentally the impact of supervision from a qualified exercise scientist.
+We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership both retrospectively and through a prospective experimental study.
+As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members in which we investigated experimentally the impact of supervision from a qualified exercise scientist.
 
@@
 All Kieser Australia members are prescribed the same protocol.
-This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action).
-Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist.
+This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds (though an upper limit of ~180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action).
+Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist .
 Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session.
@@
 Secondly, they allowed us to also explore in the descriptive[^2] difference between Core and Assisted training *in situ*, but also that we could compare the prior distribution for the random sample of Core members *in situ* with the posterior distribution of Core members completing their Core session in the experimental study to understand the extent to which study effects might influence outcomes.
+
+The experimental study component of this project was approved by the Southampton Solent University Health Exercise and Sport Science Ethics committee (id: STEELEAUG2016).
+All participants provided informed consent to participate in the study.
+Data in the prior sample was used in de-identified form and with member consent for data to be used for research purposes.
 
@@
 For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.
-Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`.
+Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`).
 For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.
@@
 
-All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland).
+All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland) including the A1 (hip extension), A2 (torso flexion), A3 (hip abduction), B1 (leg extension), B6 (leg press), B7 (seated leg curl), C3 (torso arm i.e., pulldown), C7 (seated row), D5 (arm cross i.e., pectoral fly), D6 (chest press), D7 (seated dip), F2/F2.1 (abdominal flexion), and K2 (supported supinated grip pullup).
 For the experimental study participants were instructed to attend two training sessions at least one-week apart where they completed either their current Core session (i.e., unsupervised) or an Assisted session (i.e., supervised).
@@
 ```         
-    UPDATE LINKS!!!!!!!!
+    JAMES TO UPDATE LINKS/CITATION IN PARAGRAPH BELOW!!!!!!!!
 ```
@@
 As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval.
-Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows:were as follows (note that the model for the prior dataset omits the random effect for machine):
+Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows (note that the model for the prior dataset omits the random effect for machine):
 
@@
                                variables = "core_assisted",
-                               re_formula = NA)
+                               re_formula = NA) |>
+  mutate(across(where(is.numeric), round, 2))
 
@@
                                re_formula = NA,
-                               dpar = "hu")
+                               dpar = "hu") |>
+  mutate(across(where(is.numeric), round, 2))
 
@@
                                variables = "core_assisted",
-                               re_formula = NA)
+                               re_formula = NA) |>
+  mutate(across(where(is.numeric), round, 2))
+
+preds_prior_rpe_phi <- avg_predictions(model_prior_sample_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
+
+contrast_prior_rpe_phi <- avg_comparisons(model_prior_sample_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                               dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
+
+preds_rpe_phi <- avg_predictions(model_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                             dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
+
+contrast_rpe_phi <- avg_comparisons(model_rpe,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                             dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
 
@@
                                variables = "core_assisted",
-                               re_formula = NA)
+                               re_formula = NA) |>
+  mutate(across(where(is.numeric), round, 2))
+
+preds_discomfort_phi <- avg_predictions(model_discomfort,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                             dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
+
+contrast_discomfort_phi <- avg_comparisons(model_discomfort,
+                               variables = "core_assisted",
+                               re_formula = NA,
+                             dpar = "phi",
+                               type = "link") |>
+  mutate(across(where(is.numeric), round, 2))
 
@@
 The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_rpe$estimate)` [95%QI: `r scales::percent(contrast_prior_rpe$conf.low)`, `r scales::percent(contrast_prior_rpe$conf.high)`].
+The $\phi$ parameter for precision of the beta distribution was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition.
+The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].
 
@@
 The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_rpe$estimate)` [95%QI: `r scales::percent(contrast_rpe$conf.low)`, `r scales::percent(contrast_rpe$conf.high)`].
+Similarly to the prior data, the $\phi$ parameter for precision of the beta distribution in our experimental sample was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition.
+The $\phi$ parameter for Core sessions was `r preds_rpe_phi$estimate[2]` [95%QI: `r preds_rpe_phi$conf.low[2]`, `r preds_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_rpe_phi$estimate[1]` [95%QI: `r preds_rpe_phi$conf.low[1]`, `r preds_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_rpe_phi$estimate` [95%QI: `r contrast_rpe_phi$conf.low`, `r contrast_rpe_phi$conf.high`].
 
 The rating of perceived discomfort results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions.
-The estimated rating of perceived discomfort as a percentage for Core sessions was `r scales::percent(preds_discomfort$estimate[2])` [95%QI: `r scales::percent(preds_discomfort$conf.low[2])`, `r scales::percent(preds_discomfort$conf.high[2])`], for Assisted sessions `r scales::percent(preds_discomfort$estimate[1])` [95%QI: `r scales::percent(preds_discomfort$conf.low[1])`, `r scales::percent(preds_discomfort$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_discomfort$estimate)` [95%QI: `r scales::percent(contrast_discomfort$conf.low)`, `r scales::percent(contrast_discomfort$conf.high)`].
+The estimated rating of perceived discomfort as a percentage for Core sessions was `r preds_discomfort$estimate[2]*10` [95%QI: `r preds_discomfort$conf.low[2]*10`, `r preds_discomfort$conf.high[2]*10`], for Assisted sessions `r preds_discomfort$estimate[1]*10` [95%QI: `r preds_discomfort$conf.low[1]*10`, `r preds_discomfort$conf.high[1]*10`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort$estimate*10` [95%QI: `r contrast_discomfort$conf.low*10`, `r contrast_discomfort$conf.high*10`].
+The $\phi$ parameter for precision of the beta distribution was higher in the Core condition suggesting that rating of perceived discomfort reported was more clustered around the estimated rating of perceived discomfort values as compared to in the Assisted condition, though the contrast for this was less precise than in the RPE models.
+The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].
 
@@
 # Discussion
+Our study is a first to offer insights into the role of supervision (i.e., Core vs Assisted sessions) during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
+Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and rating of perceived discomfort.
+Our key findings were that under real world settings in our prior sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range.
+There was a small difference in RPE reported with and without supervision under both real world and experimental conditions with members reporting higher RPE in Assisted training conditions, though RPE on the percentage points scale was notably higher during the experimental study compared with the prior sample for both Core and Assisted sessions.
+Lastly, rating of perceived discomfort was also higher in the experimental study under supervised Assisted sessions compared to Core sessions.
+These results in general support prior work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022a; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
 
-ADD DISCUSSION
+We will first discuss the findings from the prior sample of Core and Assisted members.
+In this *in situ* dataset we find for both Core and Assisted members there is clear evidence of targeting a specific TUL based on the prescribed TUL range of 90-120 seconds.
+For Core members this is perhaps unsurprising given that, in repetition count based RT prescription, when given the opportunity to self-select repetition number people tend to target ~10 repetitions regardless of the load they self-select and that also most RT prescriptions tend to emphasise that particular target [@steeleAreTraineesLifting2022].
+Thus, that members are also shooting for targets under this slightly different type of RT protocol suggests that this might be characteristic of most people training under real world unsupervised conditions.
+It is however surprising that this is also the case for Assisted members and that TUL differs very little between Core and Assisted sessions.
+However, the lack of difference in TUL itself is not necessarily indicative of suboptimal training effort, and we did observe slightly higher reported RPE during Assisted training sessions compared to Core sessions.
+Thus it might be that during Assisted sessions trainees might be more likely to use loads more appropriate for the prescribed TUL range and thus train with closer proximity to failure as compared to under Core sessions in addition to greater clustering of RPE around higher values (indicated by the $\phi$ parameter in the ordered beta regression).
+This may be due to load progression protocols being more closely followed by supervising exercise scientists as compared to those training unsupervised where, unless using initially heavy loads or in previously well trained persons, most unsupervised trainees do not progress load sufficiently [@steeleAreTraineesLifting2022; @fisherRoleSupervisionResistance2022a].
+In future work we plan to explore differences longitudinally in load progression between Core and Assisted members.
+But, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our prior sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort and many perhaps "going through the motions" [@steeleHigherEffortbasedParadigm2017].
+
+In contrast to the prior sample, our experimental study clearly showed a benefit to TUL performance with supervision suggesting that it might enhance trainee effort.
+Further, given we recruited existing Core members and had them use their current training loads, it also reinforces that members may be typically selecting loads that are too light and also training with considerable distance from momentary failure.
+Though the prescribed protocol *in situ* in the Kieser Australia clinics is for trainees to perform each exercise to momentary failure, despite the average TUL under Core conditions being similar between prior sample and experimental sample (see @fig-tul), we can see in @fig-raw that many were capable of achieving very long TUL before reaching momentary failure in both the Core and Assisted sessions.
+Interestingly whilst the probability of targeting the 120 second threshold was similar in the prior sample between session types, in the experimental study this was far greater in the Core session condition.
+We speculate that it might be a result of many members, whom were typically not achieving the 120 second TUL upper threshold in their prior sessions, interpreting our explicit instructions in the study to continue to momentary failure as being to ensure that they continue to this point.
+Whereas the wide range of TULs, and particularly some of the very high ones, might suggest that others more literally interpreted the instruction to continue to momentary failure.
+That under the explicit experimental study instructions to train to momentary failure members were likely to train with a greater effort under both conditions is reinforced by the greater RPE in percentage points compared to that reported in the prior sample.
+In addition to greater average RPE in both conditions, there was also greater clustering of RPE reported under the Assisted condition suggesting that members might have achieved closer proximity to failure whilst supervised.
+It was also the case that ratings of perceived discomfort were higher under supervision which, as previous evidence has suggested increases with greater proximity to failure [@refaloImprovedUnderstandingProximitytofailure2022a; @refaloInfluenceResistanceTraining2023], also supports that members trained with greater effort whilst supervised.
+
+The limitations of the present results and conclusions should be noted.
+Firstly, the lack of difference in our prior sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors causal of the TUL during training (member characteristics, load used and load progression).
+However, the session RPE data does perhaps corroborate the inference that, whilst training with slightly greater effort during Assisted training, members *in situ* may still be training with suboptimal effort compared to what occurs under experimental study conditions when explicitly the instruction to train to momentary failure is reinforced.
+But, a further issue with this is that in that two separate tools/protocols for capturing RPE was used in the prior sample (session RPE with and adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
+[-@steeleDifferentiationPerceivedEffort2017]).
+The adaptation to the session RPE scale in particular involved changes to verbal anchors and it was noticed in the raw distribution of data (see @fig-raw) that there was an abundance of values of 13 reported.
+Whether or not the tendency for lower RPE in percentage points in the prior sample is necessarily caused by reporting behaviours due to scale construction or due to members actually training with lower effort *in situ* is not clear.
+It is planned that for future member data collection we will move towards an exercise by exercise approach using the Steele et al.
+[-@steeleDifferentiationPerceivedEffort2017] scales.
+The similar degree of contrast between Core and Assisted conditions in both prior sample and experimental study data though does reinforce the role of supervision in influencing trainee effort.
+Lastly, despite the impact of supervision on trainee effort, it is still relatively unclear the extent to which it influences adaptation to training.
+As noted, Robinson et al.
+[-@robinsonExploringDoseResponse2024] suggest that higher effort through load and/or proximity to failure is important for outcomes and  Fisher et al.
+[-@fisherRoleSupervisionResistance2022a] reported a moderate standardised mean effect in favour of supervision for strength adaptation (0.40 \[95%CI: 0.06, 0.74\]) but estimates of these effects are still relatively imprecise.
+Kieser Australia also collect isometric strength test data from their members and so in future work we intend to explore this longitudinally and generate causal estimates regarding the role of supervision during RT from a large sample of members.
+
+# Conclusions
+To our knowledge this study is the first examine the role of supervision during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
+Under real world settings there was little difference in exercise performance, yet in our experimental study there was a clear benefit to performance when under supervision.
+There was a small difference in rating of perceived effort reported with and without supervision under both real world and experimental conditions suggesting a that under supervision trainees train with greater proximity to failure, which was also supported by greater rating of perceived discomfort under supervision.
+These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+
+      JAMES TO UPDATE THESE SECTIONS!!!!! BUT FEEL FREE TO ADD ANYTHING YOU THINK APPROPRIATE
 

### 2025-01-21 R/functions.R

@@ -536,6 +536,24 @@ get_contrast_draws_rpe <- function(model, prior_model) {
   )
 }
 
+get_pred_draws_rpe_phi <- function(model, prior_model) {
+  pred_draws <- bind_rows(
+    predictions(prior_model) |>
+      get_draws() |>
+      mutate(distribution = "prior"),
+    predictions(model) |>
+      get_draws() |>
+      mutate(distribution = "posterior")
+  ) |>
+    mutate(core_assisted = factor(
+      case_when(
+        core_assisted == "core" ~ "Core",
+        core_assisted == "assisted" ~ "Assisted"
+      ),
+      levels = c("Core", "Assisted")
+    ))
+}
+
 # Rating of perceived discomfort outcome
 
 get_pred_draws_discomfort <- function(model) {
@@ -558,6 +576,17 @@ get_contrast_draws_discomfort <- function(model) {
                     variables = "core_assisted",
                     re_formula = NA) |>
       get_draws()
+}
+
+get_pred_draws_discomfort_phi <- function(model) {
+  pred_draws <- predictions(model) |>
+    mutate(core_assisted = factor(
+      case_when(
+        core_assisted == "core" ~ "Core",
+        core_assisted == "assisted" ~ "Assisted"
+      ),
+      levels = c("Core", "Assisted")
+    ))
 }
 
 # Plot results -----

### 2025-01-21 _targets.R

@@ -154,6 +154,11 @@ list(
   ),
   
   tar_target(
+    pred_draws_rpe_phi,
+    get_pred_draws_rpe_phi(model_rpe, model_prior_sample_rpe)
+  ),
+  
+  tar_target(
     pred_draws_discomfort,
     get_pred_draws_discomfort(model_discomfort)
   ),
@@ -161,6 +166,11 @@ list(
   tar_target(
     contrast_draws_discomfort,
     get_contrast_draws_discomfort(model_discomfort)
+  ),
+  
+  tar_target(
+    pred_draws_discomfort_phi,
+    get_pred_draws_discomfort_phi(model_discomfort)
   ),
   
   # Plotting data and results -----

### 2025-02-18 plots/diagnostic_plots.qmd
Add diagnostic plots

---
title: "Diagnostic Plots"
subtitle: "The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
author: "James Steele, James Fisher, Tim Dettmann"
format:
  pdf:
    documentclass: report
    toc: false
    number-sections: false
    highlight-style: github
    fig-width: 10
    fig-height: 10
---
```{r,message=FALSE,warning=FALSE,echo=FALSE}
targets::tar_config_set(store = here::here('_targets'))

```

# Prior Sample Models

## Time Under Load

### $\hat{R}$

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(rhat_prior_sample_tul)

rhat_prior_sample_tul

```

### Trace plots

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(trace_plot_prior_sample_tul)

trace_plot_prior_sample_tul

```

### Posterior predictive check

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(pp_check_prior_sample_tul)

pp_check_prior_sample_tul

```

## Rating of Perceived Effort

### $\hat{R}$

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(rhat_prior_sample_rpe)

rhat_prior_sample_rpe

```

### Trace plots

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(trace_plot_prior_sample_rpe)

trace_plot_prior_sample_rpe

```

### Posterior predictive check

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(pp_check_prior_sample_rpe)

pp_check_prior_sample_rpe

```

# Experimental Study Models

## Time Under Load

### $\hat{R}$

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(rhat_tul)

rhat_tul

```

### Trace plots

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(trace_plot_tul)

trace_plot_tul

```

### Posterior predictive check

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(pp_check_tul)

pp_check_tul

```

## Rating of Perceived Effort

### $\hat{R}$

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(rhat_rpe)

rhat_rpe

```

### Trace plots

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(trace_plot_rpe)

trace_plot_rpe

```

### Posterior predictive check

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(pp_check_rpe)

pp_check_rpe

```

## Rating of Perceived Discomfort

### $\hat{R}$

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(rhat_discomfort)

rhat_discomfort

```

### Trace plots

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(trace_plot_discomfort)

trace_plot_discomfort

```

### Posterior predictive check

```{r,message=FALSE,warning=FALSE,echo=FALSE}

targets::tar_load(pp_check_discomfort)

pp_check_discomfort

```

### 2025-02-18 pre_print/pre_print.qmd
Preprint ready to submit (sans doi)

@@
       - ref: 3
+      - ref: 4
     orcid: 0000-0002-6013-8402
@@
 abstract: |
-  Recent work has suggested that supervision during resistance training might enhance outcomes such as strength adaptation due to role that it plays in optimising trainee effort.
-We were afforded an opportunity to examine the role of supervision on effort during resistance training in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership or supervised by a qualified exercise scientist on an "Assisted" membership.
-Both retrospectively through examination of samples of member training records, and through a prospective experimental study, we examined the effects of supervision upon exercise performance as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort.
-We adopted Bayesian methods for modelling and from the prior sample of retrospective data generated empirically informed posterior distributions to use as prior distributions in our modelling of the experimental study.
-The prior sample consisted of training sessions from 1000 each of Core and Assisted members, whilst the experimental study consisted of Core members recruited to in randomised order complete an unsupervised and a supervised session using their current training loads where they were instructed to perform each exercise to momentary failure.
-Our findings suggest that, *in situ* under real world settings (i.e., our prior sample of members) there was little difference in exercise performance, yet in our experimental study there was a clear benefit to TUL performance when under supervision (Core = 125.12 [95%QI: 113.70, 131.90] seconds; Assisted = 147.35 [95%QI: 134.29, 154.81] seconds; Core minus Assisted contrast = -22.10 [95%QI: -26.60, -17.61] seconds).
-There was a small difference in RPE in percentage points reported with and without supervision under both real world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; Core minus Assisted contrast = -6% [95%QI: -8%, -4%]) and experimental conditions (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; Core minus Assisted contrast = -6% [95%QI: -10%, -4%]) suggesting a that under supervision trainees train with greater proximity to failure, which was also supported by greater rating of perceived discomfort under supervision during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3] pts; Assisted = 7.5 [95%QI: 6.5, 8.3] pts; Core minus Assisted contrast = -1.2 [95%QI: -1.6, -0.9] pts).
-These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+  Supervision during resistance training (RT) may enhance strength gains by optimizing trainee effort.
+We investigated supervision’s role in effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership).
+Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort (RPD).
+Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study.
+The prior sample included 1000 training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
+Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training.
+However, in our experimental study, supervision significantly improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
+RPE was slightly higher with supervision in both real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
+This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]).
+Overall, these results reinforce prior research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
 license-type: ccby # change if neccessary
@@
 # optional link to Supplementary Material:
-suppl-link: https://osf.io/ 
+suppl-link: https://osf.io/etb34/
 reference-section-title: References
@@
 In a recent systematic review and meta-analysis from Fisher et al.
-[-@fisherRoleSupervisionResistance2022a] there was a moderate standardised mean effect (0.40 \[95%CI: 0.06, 0.74\]) of supervised vs unsupervised RT on strength outcomes synthesised from ten studies.
+[-@fisherRoleSupervisionResistance2022] there was a moderate standardised mean effect (0.40 \[95%CI: 0.06, 0.74\]) of supervised vs unsupervised RT on strength outcomes synthesised from ten studies.
 Since then, a further study has reported that even in previously trained participants there may be benefit to strength outcomes from supervision [@colemanSupervisionResistanceTraining2023].
 
-It has been speculated [@fisherRoleSupervisionResistance2022a] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort.
+It has been speculated [@fisherRoleSupervisionResistance2022] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort.
 Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
-Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024].
-Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.
+Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., they select \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024].
+Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will naturally perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.
 
-As such it has also been suggested that a supervising trainer plays the role of providing motivation and enhancing trainee effort [@fisherRoleSupervisionResistance2022a].
+As such it has also been suggested that a supervising trainer plays the role of providing motivation and enhancing trainee effort [@fisherRoleSupervisionResistance2022].
 Indeed, recent survey studies highlight that trainees perceive supervision to have an important role in determining their motivation and resultant effort during RT which they also perceive to be important to achieve their training goals [@fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024].
@@
 We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership both retrospectively and through a prospective experimental study.
-As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members in which we investigated experimentally the impact of supervision from a qualified exercise scientist.
+As such, using Bayesian methods, we examined samples of historical data from both types of member and in a sample of current Core members we investigated experimentally the impact of supervision from a qualified exercise scientist.
 
-[^1]: Note, Core members do have a "Review" session every twentieth session with an exercise scientist to complete strength testing, and review their training programme cards including selection of exercises and weights.
+[^1]: Note, Core members do typically have a "Review" session every ~3 months with an exercise scientist to complete strength testing, and review their training programme cards including selection of exercises and weights.
 
@@
 All Kieser Australia members are prescribed the same protocol.
-This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds (though an upper limit of ~180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action).
-Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist .
+This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action).
+Core trainees are prompted to progress load by \~5% for the next session once a TUL of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist.
 Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session.
@@
 Participants were instructed to complete a single set of the resistance machine exercises on their current training programme card using their current training loads to momentary failure.
-The time under load (TUL) performed, and perceptions of effort and discomfort were recorded and compared between both conditions.
+The TUL performed, and ratings of perceptions of effort (RPE) and discomfort (RPD) were recorded and compared between both conditions.
 This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.
@@
 In addition to the experimental study design noted above, we also took samples of Core and Assisted trainees from historical data records in Kieser Australia's database.
-We took separate random samples to obtain both TUL and session RPE data.
-There were two reasons for this.
+We took separate random samples to obtain both TUL and session RPE data as not all members record their RPE.
+There were two reasons for examining these historical samples.
 Firstly, we used this data to generate empirical informative prior distributions to use in our Bayesian modelling of the experimental study data.
-Secondly, they allowed us to also explore in the descriptive[^2] difference between Core and Assisted training *in situ*, but also that we could compare the prior distribution for the random sample of Core members *in situ* with the posterior distribution of Core members completing their Core session in the experimental study to understand the extent to which study effects might influence outcomes.
+Secondly, they allowed us to also explore the descriptive[^2] difference between Core and Assisted training *in situ*, but also that we could compare the prior distribution for the random samples of members *in situ* with the posterior distribution of members completing their sessions in the experimental study to understand the extent to which study effects might also influence outcomes.
 
@@
 
-[^2]: Note that we do not claim this difference to be a causal effect in this data given that we have not adjusted for confounders, for example that different types of people might self-select into either Core or Assisted memberships and that this might confound comparisons of either TUL or session RPE.
+[^2]: Note that we do not claim this difference to be an estimated causal effect in this data given that we have not adjusted for confounders; for example, that different types of people might self-select into either Core or Assisted memberships and that this might confound comparisons of either TUL or session RPE.
 
@@
 The study was advertised at five Kieser Australia strength clinics and we sought to recruit members who had \> 6 months prior training experience at the strength clinics, were healthy (no clinical conditions on their member record as recorded by a physiotherapist or exercise physiologist), both males and females (age +18 years), and without any current condition for which RT would be contraindicated.
-They were recruited from the existing client pool of Core members.
+They were recruited from the existing member pool of Core members.
 We recruited a sample size based on considerations of what was logistically feasible given the availability of the exercise scientist staff at each clinic to conduct data collection and perform the supervised Assisted sessions and wanting to minimise the burden on day-to-day operation of the clinics.
@@
 
-[^3]: Whilst clients can have Core or Assisted memberships the former also have sessions that are supervised (i.e., their "Review" sessions noted in footnote$^1$) and the latter can have unsupervised sessions (i.e., they can attend the clinic outside of their scheduled sessions with the exercise scientist and train unsupervised).
+[^3]: Whilst members can have Core or Assisted memberships the former also have sessions that are supervised (i.e., their "Review" sessions noted in footnote$^1$) and the latter can have unsupervised sessions (i.e., they can attend the clinic outside of their scheduled sessions with the exercise scientist and train unsupervised).
 Thus we limited our sample to only normal training sessions which had been either unsupervised or supervised for both Core and Assisted members respectively.
@@
 All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland) including the A1 (hip extension), A2 (torso flexion), A3 (hip abduction), B1 (leg extension), B6 (leg press), B7 (seated leg curl), C3 (torso arm i.e., pulldown), C7 (seated row), D5 (arm cross i.e., pectoral fly), D6 (chest press), D7 (seated dip), F2/F2.1 (abdominal flexion), and K2 (supported supinated grip pullup).
+Note, not all participants performed exercise using each machine, but had at least some of these machines in their current training cards.
 For the experimental study participants were instructed to attend two training sessions at least one-week apart where they completed either their current Core session (i.e., unsupervised) or an Assisted session (i.e., supervised).
-During the Core session the participant completed their training session as prescribed with the only difference for the Assisted session being that an exercise scientist supervised them providing instruction and motivation, however, they did not interfere in any physical way during the set (e.g.
+During the Core session the participant completed their training session as prescribed with the only difference for the Assisted session being that an exercise scientist supervised them providing instruction and motivation; however, they did not interfere in any physical way during the set (e.g.
 to spot or assist in completing a failing repetition).
@@
 In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card prior to participation in the study.
-Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the time-under-load achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL.
+Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the TUL achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL acheived.
 Momentary failure was defined as per Steele et al.
-[-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was communicated to the participants.
-During each session for each exercise the trainees/trainers recorded their time-under-load achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose.
-In addition, they recorded their perceptions of effort and discomfort in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).
+[-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was clearly communicated to the participants.
+During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose.
+In addition, they recorded their RPE and RPD in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).
 
 The data generated from the sample of historical members was generated according to the standard protocol described above that all Kieser Australia members are prescribed.
-TUL was also recorded similarly using timers available about the clinic, r for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which has placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.
-Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale with (where the scale is displayed as scrollable and the verbal anchors are "6 - None", "7- Very, very light", "13 – Somewhat hard", "19 - Very, very hard", and "20 – Maximal Exertion"
-)
+TUL was also recorded similarly using timers available about the clinic, or for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which is placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.
+Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale where the scale is displayed as scrollable and the verbal anchors are "6 - None", "7- Very, very light", "13 – Somewhat hard", "19 - Very, very hard", and "20 – Maximal Exertion".
 
 ## Statistical Analysis
-
-```         
-    JAMES TO UPDATE LINKS/CITATION IN PARAGRAPH BELOW!!!!!!!!
-```
-
-All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/dqwh5/> or the corresponding GitHub repository <https://github.com/jamessteeleii/self_talk_meta_analysis_update>.
-We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://osf.io/ftajc>.
+All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/etb34/> or the corresponding GitHub repository <https://github.com/jamessteeleii/supervision_tul>.
+We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://osf.io/ew79g>.
 
@@
 All models were run with 2000 warmup and 2000 sampling iterations and four Monte Carlo Markov Chains.
-Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions.
-We fit two sets of models for the TUL and RPE outcomes, one on the prior sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the rating of perceived discomfort outcome using weakly regularising priors.
+Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions (see <https://osf.io/u7g8c>.
+We fit two sets of models for the TUL and RPE outcomes, one on the prior sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the RPD outcome using weakly regularising priors.
 These models are described below.
@@
 We then employed an ordered beta regression model [@kubinecOrderedBetaRegression2022] which employs a cutpoint process similar to ordered logistic models in order to model both the continuous responses on the $(0,1)$ interval reflecting some effort though not maximal, and the degenerate response on the bounds $[0,1]$ reflecting both no and maximal effort respectively.
-We approached the rating of perceived discomfort similarly as this scale was also anchored between no discomfort and the maximum imaginable.
+We approached the RPD similarly as this scale was also anchored between no discomfort and the maximum imaginable.
 
@@
 
-We did not have prior sample data for rating of perceived discomfort and so for this outcome we used the same model as above for the RPE outcomes but with wholly default weakly regularising priors.
+We did not have prior sample data for RPD and so for this outcome we used the same model as above for the RPE outcomes but with wholly default weakly regularising priors.
 
@@
 
-Our experimental sample model showed far greater difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds, in addition to the posterior estimates differing considerably from the prior sample estimates in terms of between condition contrasts for TUL and probability of stopping t 120 seconds, Assisted condition TUL, and both conditions probability of stopping at 120 seconds.
+Our experimental sample model showed far greater difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds, in addition to the posterior estimates differing considerably from the prior sample estimates in terms of between condition contrasts for TUL and probability of stopping at 120 seconds, Assisted condition TUL, and both conditions probability of stopping at 120 seconds.
 The estimated TUL for Core sessions was `r preds_tul$estimate[2]+120` [95%QI: `r preds_tul$conf.low[2]+120`, `r preds_tul$conf.high[2]+120`] seconds, for Assisted sessions was `r preds_tul$estimate[1]+120` [95%QI: `r preds_tul$conf.low[1]+120`, `r preds_tul$conf.high[1]+120`] seconds, and the between condition contrast (Core minus Assisted) was `r contrast_tul$estimate` [95%QI: `r contrast_tul$conf.low`, `r contrast_tul$conf.high`] seconds.
@@
 
-The rating of perceived discomfort results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions.
-The estimated rating of perceived discomfort as a percentage for Core sessions was `r preds_discomfort$estimate[2]*10` [95%QI: `r preds_discomfort$conf.low[2]*10`, `r preds_discomfort$conf.high[2]*10`], for Assisted sessions `r preds_discomfort$estimate[1]*10` [95%QI: `r preds_discomfort$conf.low[1]*10`, `r preds_discomfort$conf.high[1]*10`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort$estimate*10` [95%QI: `r contrast_discomfort$conf.low*10`, `r contrast_discomfort$conf.high*10`].
-The $\phi$ parameter for precision of the beta distribution was higher in the Core condition suggesting that rating of perceived discomfort reported was more clustered around the estimated rating of perceived discomfort values as compared to in the Assisted condition, though the contrast for this was less precise than in the RPE models.
-The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].
+The RPD results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions.
+The estimated RPD as a percentage for Core sessions was `r preds_discomfort$estimate[2]*10` [95%QI: `r preds_discomfort$conf.low[2]*10`, `r preds_discomfort$conf.high[2]*10`], for Assisted sessions `r preds_discomfort$estimate[1]*10` [95%QI: `r preds_discomfort$conf.low[1]*10`, `r preds_discomfort$conf.high[1]*10`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort$estimate*10` [95%QI: `r contrast_discomfort$conf.low*10`, `r contrast_discomfort$conf.high*10`].
+The $\phi$ parameter for precision of the beta distribution was higher in the Core condition suggesting that RPD reported was more clustered around the estimated RPD values as compared to in the Assisted condition, though the contrast for this was less precise than in the RPE models.
+The $\phi$ parameter for Core sessions was `r preds_discomfort_phi$estimate[2]` [95%QI: `r preds_discomfort_phi$conf.low[2]`, `r preds_discomfort_phi$conf.high[2]`], for Assisted sessions `r preds_discomfort_phi$estimate[1]` [95%QI: `r preds_discomfort_phi$conf.low[1]`, `r preds_discomfort_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort_phi$estimate` [95%QI: `r contrast_discomfort_phi$conf.low`, `r contrast_discomfort_phi$conf.high`].
 
@@
 Our study is a first to offer insights into the role of supervision (i.e., Core vs Assisted sessions) during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
-Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and rating of perceived discomfort.
+Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.
 Our key findings were that under real world settings in our prior sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range.
 There was a small difference in RPE reported with and without supervision under both real world and experimental conditions with members reporting higher RPE in Assisted training conditions, though RPE on the percentage points scale was notably higher during the experimental study compared with the prior sample for both Core and Assisted sessions.
-Lastly, rating of perceived discomfort was also higher in the experimental study under supervised Assisted sessions compared to Core sessions.
-These results in general support prior work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022a; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+Lastly, RPD was also higher in the experimental study under supervised Assisted sessions compared to Core sessions.
+These results in general support prior work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
 
@@
 In this *in situ* dataset we find for both Core and Assisted members there is clear evidence of targeting a specific TUL based on the prescribed TUL range of 90-120 seconds.
-For Core members this is perhaps unsurprising given that, in repetition count based RT prescription, when given the opportunity to self-select repetition number people tend to target ~10 repetitions regardless of the load they self-select and that also most RT prescriptions tend to emphasise that particular target [@steeleAreTraineesLifting2022].
-Thus, that members are also shooting for targets under this slightly different type of RT protocol suggests that this might be characteristic of most people training under real world unsupervised conditions.
+For Core members this is perhaps unsurprising given that, in repetition count based RT prescription, when given the opportunity to self-select repetition number people tend to target ~10 repetitions regardless of the load they self-select and that also most RT prescriptions tend to emphasise that particular target repetition number [@steeleAreTraineesLifting2022].
+Thus, that members are also shooting for targets under this TUL focused RT protocol suggests that this might be characteristic of most people training under real world unsupervised conditions.
 It is however surprising that this is also the case for Assisted members and that TUL differs very little between Core and Assisted sessions.
@@
 Thus it might be that during Assisted sessions trainees might be more likely to use loads more appropriate for the prescribed TUL range and thus train with closer proximity to failure as compared to under Core sessions in addition to greater clustering of RPE around higher values (indicated by the $\phi$ parameter in the ordered beta regression).
-This may be due to load progression protocols being more closely followed by supervising exercise scientists as compared to those training unsupervised where, unless using initially heavy loads or in previously well trained persons, most unsupervised trainees do not progress load sufficiently [@steeleAreTraineesLifting2022; @fisherRoleSupervisionResistance2022a].
-In future work we plan to explore differences longitudinally in load progression between Core and Assisted members.
-But, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our prior sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort and many perhaps "going through the motions" [@steeleHigherEffortbasedParadigm2017].
+This may be due to load progression protocols being more closely followed by supervising exercise scientists as compared to those training unsupervised where, unless using initially heavy loads or in previously well trained persons, most unsupervised trainees do not progress load sufficiently [@steeleAreTraineesLifting2022; @fisherRoleSupervisionResistance2022].
+In future work we plan to explore more thoroughly the causal effects of supervision in both load progression and also isometric strength outcomes collected between Core and Assisted members; but, an initial descriptive examination of load progression using linear-logarithmic growth models known to theoretically describe strength progression with time from RT well [@steeleLongTermTimeCourseStrength2023a; @latellaUsingPowerliftingAthletes2024; @gschneidnerEffectsLengthenedpartialRange2024] suggests that load progressions are typically similar between membership types (see https://osf.io/wcv6g)[^5].
+In fact, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our prior sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort whether being supervised or not and many perhaps just "going through the motions" [@steeleHigherEffortbasedParadigm2017].
+
+[^5]: Similarly to the other prior historical samples used, we took a sample of 1000 Core and 1000 Assisted members fitting a linear-logarithmic model of time in weeks interacting with membership type (Core vs Assisted) with random intercepts for location, machine, and participant id, and random slopes for log time for each level.
+The model was fit using Restricted Maximum Likelihood estimation with the `lme4` package [@batesFittingLinearMixedeffects2015].
 
@@
 Interestingly whilst the probability of targeting the 120 second threshold was similar in the prior sample between session types, in the experimental study this was far greater in the Core session condition.
-We speculate that it might be a result of many members, whom were typically not achieving the 120 second TUL upper threshold in their prior sessions, interpreting our explicit instructions in the study to continue to momentary failure as being to ensure that they continue to this point.
+We speculate that it might be a result of many members, whom were typically not achieving the 120 second TUL upper threshold in their prior sessions, interpreting our explicit instructions in the study to continue to momentary failure as being to ensure that they continue to this upper TUL range target.
 Whereas the wide range of TULs, and particularly some of the very high ones, might suggest that others more literally interpreted the instruction to continue to momentary failure.
-That under the explicit experimental study instructions to train to momentary failure members were likely to train with a greater effort under both conditions is reinforced by the greater RPE in percentage points compared to that reported in the prior sample.
 In addition to greater average RPE in both conditions, there was also greater clustering of RPE reported under the Assisted condition suggesting that members might have achieved closer proximity to failure whilst supervised.
-It was also the case that ratings of perceived discomfort were higher under supervision which, as previous evidence has suggested increases with greater proximity to failure [@refaloImprovedUnderstandingProximitytofailure2022a; @refaloInfluenceResistanceTraining2023], also supports that members trained with greater effort whilst supervised.
+It was also the case that ratings of perceived discomfort were higher under supervision which, as previous evidence has suggested increases with greater proximity to failure [@refaloImprovedUnderstandingProximitytofailure2022a; @refaloInfluenceResistanceTraining2023; @refaloEffectProximityToFailurePerceptual2025], also supports that members trained with greater effort whilst supervised.
+
+Of interest is the considerable difference between the *in situ* prior historical sample and the experimental study.
+That under the explicit experimental study instructions to train to momentary failure members were likely to train with a greater effort under both unsupervised and supervised conditions, as evidenced by the TUL results, is reinforced by the greater RPE in percentage points compared to that reported in the prior sample.
+Indeed, this may reflect some degree of the so called "Hawthorne Effect" in both members and the exercise scientists supervising sessions wherein both were more likely to train/supervise training to a higher degree of effort and thus performance whilst under the auspice of the explicit experimental study instructions and observation[^6].
+Whilst this so called effect is not particularly well understood in its specifics there is general agreement that both participants and researchers react differently under observational settings such as research [@mccambridgeSystematicReviewHawthorne2014; @paradisGoodStoryHawthorne2017] as seems to be observed here.
+
+[^6]: At Kieser Australia the exercise scientists undertake an "exam" / "assessment" after 3-months of being employed and anecdotally we have noted from staff conducting these observations that those exercise scientists will change how they run a session when they are being watched as part of this process.
 
 The limitations of the present results and conclusions should be noted.
-Firstly, the lack of difference in our prior sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors causal of the TUL during training (member characteristics, load used and load progression).
+Firstly, the lack of difference in our prior sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors possibly causal of the TUL during training (member characteristics, load used and load progression).
 However, the session RPE data does perhaps corroborate the inference that, whilst training with slightly greater effort during Assisted training, members *in situ* may still be training with suboptimal effort compared to what occurs under experimental study conditions when explicitly the instruction to train to momentary failure is reinforced.
-But, a further issue with this is that in that two separate tools/protocols for capturing RPE was used in the prior sample (session RPE with and adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
+But, a further issue with this is that two separate tools/protocols for capturing RPE were used in the prior sample (session RPE with an adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
 [-@steeleDifferentiationPerceivedEffort2017]).
-The adaptation to the session RPE scale in particular involved changes to verbal anchors and it was noticed in the raw distribution of data (see @fig-raw) that there was an abundance of values of 13 reported.
+Session RPE is historically captured using the Borg Category Ratio Scale [@fosterMonitoringTrainingAthletes1998] but for reasons unclear a choice was made when originally implementing this data capture to use the 6-20 scale.
+The adaptation to the 6-20 scale in particular involved changes to verbal anchors and it was noticed in the raw distribution of data (see @fig-raw) that there was an abundance of values of 13 reported.
 Whether or not the tendency for lower RPE in percentage points in the prior sample is necessarily caused by reporting behaviours due to scale construction or due to members actually training with lower effort *in situ* is not clear.
-It is planned that for future member data collection we will move towards an exercise by exercise approach using the Steele et al.
+It is planned that for future member data collection Kieser Australia will move towards an exercise by exercise approach using adaptations of the Steele et al.
 [-@steeleDifferentiationPerceivedEffort2017] scales.
-The similar degree of contrast between Core and Assisted conditions in both prior sample and experimental study data though does reinforce the role of supervision in influencing trainee effort.
+The similar degree of contrast between Core and Assisted conditions in both prior sample and experimental study data though does reinforce the role of supervision in influencing trainee effort, or at least reported perception of effort.
 Lastly, despite the impact of supervision on trainee effort, it is still relatively unclear the extent to which it influences adaptation to training.
@@
 [-@robinsonExploringDoseResponse2024] suggest that higher effort through load and/or proximity to failure is important for outcomes and  Fisher et al.
-[-@fisherRoleSupervisionResistance2022a] reported a moderate standardised mean effect in favour of supervision for strength adaptation (0.40 \[95%CI: 0.06, 0.74\]) but estimates of these effects are still relatively imprecise.
-Kieser Australia also collect isometric strength test data from their members and so in future work we intend to explore this longitudinally and generate causal estimates regarding the role of supervision during RT from a large sample of members.
+[-@fisherRoleSupervisionResistance2022] reported a moderate standardised mean effect in favour of supervision for strength adaptation (0.40 \[95%CI: 0.06, 0.74\]) but estimates of these effects are still relatively imprecise.
+Kieser Australia also collect isometric strength test data from their members and so in future work we intend to explore this longitudinally and generate causal estimates regarding the role of supervision during RT among other variables from a large sample of members.
 
@@
 Under real world settings there was little difference in exercise performance, yet in our experimental study there was a clear benefit to performance when under supervision.
-There was a small difference in rating of perceived effort reported with and without supervision under both real world and experimental conditions suggesting a that under supervision trainees train with greater proximity to failure, which was also supported by greater rating of perceived discomfort under supervision.
+There was a small difference in RPE reported with and without supervision under both real world and experimental conditions suggesting that under supervision trainees train with greater proximity to failure, which was also supported by greater RPD under supervision.
 These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
 
-      JAMES TO UPDATE THESE SECTIONS!!!!! BUT FEEL FREE TO ADD ANYTHING YOU THINK APPROPRIATE
+# Funding Information
+
+No direct funding was received for this study. Operational costs for its delivery were absorbed by Kieser Australia and so provided *in-kind* value.
 
@@
 
-Authors should report the contributions of each author in the a specific contribution section based on the guidelines set forth by the International Committee of Medical Journal Editors.
-
-Please indicate author contributions as clearly as possible, according to the following criteria:
-
--   Substantial contributions to conception and design
--   Acquisition of data
--   Analysis and interpretation of data
--   Drafting the article or revising it critically for important intellectual content
--   Final approval of the version to be published
+All authors conceived the idea for the project and designed the methods.
+TD coordinated and oversaw the data collection.
+JS conducted the statistical analyses and produced data visualisations.
+JS and JF drafted the initial manuscript.
+All authors contributed to editing the manuscript.
+All authors read and approved the final manuscript.
 
@@
 
-People who contributed to the work but do not fit our author criteria should be listed in the acknowledgements, along with their contributions.
-You must ensure that anyone named in the acknowledgements agrees to being so named.
-
-Funding sources should not be included in the acknowledgements, but in the section below.
-
-# Funding Information
-
-Please provide a list of the sources of funding, as well as the relevant grant numbers, where possible.
-List the authors associated with specific funding sources.
-You will also enter this information in a form during the submission process, but it must be repeated here.
+We would like to extend our sincere gratitude to all the members and exercise scientists at Kieser Australia without whom this project would not have been possible.
+We would also like to extend our thanks to Myles Moore for his feedback on the manuscript.
 
@@
 
-This should list the database(s) and, if appropriate, the respective accession numbers and DOIs for all data or supplementary material for the manuscript that has been made publicly available on a trusted digital repository.
-If no data, code, or supplementary material are available for this manuscript then the reason for this should be explained here.
+All data and code utilised for data preparation and analyses are available in either the Open Science Framework page for this project https://osf.io/etb34/ or the corresponding GitHub repository https://github.com/jamessteeleii/supervision_tul.
+Other supplementary analyses and plots are also available there.

### 2025-02-18 R/functions.R

@@ -411,6 +411,27 @@ fit_model_discomfort <- function(data) {
   )
 }
 
+# Model checks -----
+make_rhat_plot <- function(model) {
+  mod_rhat <- enframe(brms::rhat(model)) |>
+    filter(!str_detect(name, "^r_id"))
+  
+  rhat_main_params <- mod_rhat$value
+  
+  mcmc_rhat(rhat_main_params) +
+    scale_x_continuous(breaks = c(1, 1.01, 1.02, 1.03, 1.04, 1.05)) +
+    geom_vline(xintercept = 1.01,
+               linetype = "dashed",
+               alpha = 0.25)
+}
+
+make_trace_plot <- function(model) {
+  plot(model)
+}
+
+make_pp_check <- function(model) {
+  pp_check(model)
+}
 # Model post-processing -----
 
 # Time under load outcome
@@ -1142,4 +1163,95 @@ make_plot_combined_rpe_discomfort <- function(plot1, plot2, plot3, plot4) {
   
   wrap_elements(rpe_plots) / wrap_elements(discomfort_plots) 
   
+}
+
+# Additional descriptive model of load progression -----
+fit_loadprog_model <- function(data) {
+  
+  lmer(load ~ log1p(week) * core_assisted + (log1p(week) | location) +  (log1p(week) | machine) + (log1p(week) | id),
+                     data = data,
+                     REML = TRUE)
+  
+}
+
+get_preds_loadprog <- function(model) {
+  predictions(
+    model,
+    newdata = datagrid(
+      location = NA,
+      id = NA,
+      machine = NA,
+      week = 1:150,
+      core_assisted = c("core", "assisted")
+    ),
+    re.form = NA
+  )
+}
+
+get_slopes_loadprog <- function(model) {
+  slopes(
+    model,
+    newdata = datagrid(
+      location = NA,
+      id = NA,
+      machine = NA,
+      week = 1:150,
+      core_assisted = c("core", "assisted")
+    ),
+    re.form = NA,
+    variables = "week"
+  )
+}
+
+make_plot_loadprog <- function(preds, slopes) {
+  
+  pred_plot <- preds |>
+    mutate(
+      core_assisted = case_when(
+        core_assisted == "core" ~ "Core",
+        core_assisted == "assisted" ~ "Assisted"
+      )
+    ) |>
+    ggplot(aes(x=week, y=estimate, ymin=conf.low, ymax=conf.high, color=core_assisted, fill=core_assisted)) +
[…]

### 2025-02-18 _targets.R

@@ -17,6 +17,7 @@ tar_option_set(
     "ggdist",
     "tidybayes",
     "marginaleffects",
+    "lme4",
     "rstan",
     "patchwork",
     "here",
@@ -122,7 +123,89 @@ list(
     fit_model_discomfort(data)
   ),
   
-  # Model post procession -----
+  
+  
+  # Model checks -----
+  
+  # rhat plots
+  tar_target(
+    rhat_prior_sample_tul,
+    make_rhat_plot(model_prior_sample_tul)
+  ),
+  
+  tar_target(
+    rhat_prior_sample_rpe,
+    make_rhat_plot(model_prior_sample_rpe)
+  ),
+  
+  tar_target(
+    rhat_tul,
+    make_rhat_plot(model_tul)
+  ),
+  
+  tar_target(
+    rhat_rpe,
+    make_rhat_plot(model_rpe)
+  ),
+  
+  tar_target(
+    rhat_discomfort,
+    make_rhat_plot(model_discomfort)
+  ),
+  
+  # trace plots
+  tar_target(
+    trace_plot_prior_sample_tul,
+    make_trace_plot(model_prior_sample_tul)
+  ),
+  
+  tar_target(
+    trace_plot_prior_sample_rpe,
+    make_trace_plot(model_prior_sample_rpe)
+  ),
+  
+  tar_target(
+    trace_plot_tul,
+    make_trace_plot(model_tul)
+  ),
+  
+  tar_target(
+    trace_plot_rpe,
+    make_trace_plot(model_rpe)
+  ),
+  
+  tar_target(
+    trace_plot_discomfort,
+    make_trace_plot(model_discomfort)
+  ),
+  
+  # pp check
+  tar_target(
+    pp_check_prior_sample_tul,
+    make_pp_check(model_prior_sample_tul)
+  ),
+  
+  tar_target(
+    pp_check_prior_sample_rpe,
+    make_pp_check(model_prior_sample_rpe)
+  ),
+  
+  tar_target(
+    pp_check_tul,
+    make_pp_check(model_tul)
+  ),
+  
+  tar_target(
+    pp_check_rpe,
+    make_pp_check(model_rpe)
+  ),
+  
+  tar_target(
+    pp_check_discomfort,
+    make_pp_check(model_discomfort)
+  ),
+  
+  # Model post processing -----
   tar_target(
     pred_draws_tul,
     get_pred_draws_tul(model_tul, model_prior_sample_tul)
[…]

### 2025-02-18 pre_print/pre_print.qmd
DOI added to pre print

@@
 Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study.
-The prior sample included 1000 training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
+The prior sample included ~1000 members training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
 Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training.
-However, in our experimental study, supervision significantly improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
-RPE was slightly higher with supervision in both real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
+However, in our experimental study, supervision improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
+In percentage points RPE was slightly higher with supervision in both prior real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
 This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]).
@@
 # year: 2025 # defaults to current year
-keywords: [template, demo, exercise science] # optional
+keywords: [supervision, effort, resistance training, bayesian] # optional
 # optional link to Supplementary Material:

### 2025-03-04 submission/pre_print_jss_submission.qmd
Update preprint, and add journal submission outputs

--- 2025-02-18 pre_print/pre_print.qmd
@@
 ---
-articletitle: |
-  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
-# shorttitle: "Doe et al. (2022)" # usually auto-generated
-format: 
-  sportrxiv-pdf:
-    include-in-header:
-      text: |
-       \usepackage[font=scriptsize]{caption}
-       \usepackage{caption}
+title: "The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
 author:
@@
 abstract: |
-  Supervision during resistance training (RT) may enhance strength gains by optimizing trainee effort.
-We investigated supervision’s role in effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership).
-Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort (RPD).
-Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study.
-The prior sample included 1000 training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
-Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training.
-However, in our experimental study, supervision significantly improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
-RPE was slightly higher with supervision in both real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
-This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]).
-Overall, these results reinforce prior research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
-license-type: ccby # change if neccessary
-# year: 2025 # defaults to current year
-keywords: [template, demo, exercise science] # optional
-# optional link to Supplementary Material:
-suppl-link: https://osf.io/etb34/
-reference-section-title: References
-printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
+  Supervision during resistance training (RT) may optimise trainee effort.
+We examined this in private strength clinics, where members train either unsupervised (“Core”) or supervised by an exercise scientist (“Assisted”).
+Using retrospective analysis of real-world data from ~1000 members training sessions per group, and an experimental study with 45 Core members, we assessed supervision’s impact on time under load (TUL), and rating of perceived effort (RPE) and discomfort (RPD).
+Bayesian methods were used to integrate retrospective data priors and update to posteriors using data from the experimental study.
+Real-world retrospective data showed minimal performance differences, but the experimental study posterior estimates found supervision improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
+Supervision also slightly increased RPE in real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%])), indicating trainees pushed closer to failure.
+RPD was also higher with supervision.
+These findings reinforce prior research on supervision’s benefits, suggesting unsupervised trainees—especially in real-world conditions—train with suboptimal effort.
+format:
+  pdf: 
+    include-in-header:
+      text: '\usepackage{lineno}\linenumbers'
 bibliography: bibliography.bib  
@@
 ```
+
+## Author Affiliations
+James Steele: Kieser Australia, Melbourne, Australia; Steele Research Limited, Eastleigh, UK; Department of Sport and Health, Solent University, Southampton, UK
+James Fisher: Department of Sport and Health, Solent University, Southampton, UK; The Exercise Coach, Tennessee, USA
+Tim Dettmann: Kieser Australia, Melbourne, Australia
+
+Corresponding author: James Steele (james@steele-research.com)
 
@@
 \\
-X'\beta&=(\beta_0 + b_{0_{j[i],k[i],l[i]}}) + \beta_1 \text{Condition}_{ic} & \text{Vector of predictors} \\[4pt] 
-b_{0_j} &\sim \mathcal{N}(0, \sigma_{b_{0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in student}\ t\ \text{model model} \\[4pt]
-b_{0_k} &\sim \mathcal{N}(0, \sigma_{b_{0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in student}\ t\ \text{model model} \\[4pt]
-b_{0_l} &\sim \mathcal{N}(0, \sigma_{b_{0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in student}\ t\ \text{model model} \\[4pt]
-X'\beta_\phi&=(\beta_{\phi0} + b_{\phi0_{j[i],k[i],l[i]}}) + \beta_{\phi1} \text{Condition}_{ic} & \text{Vector of predictors} \\[4pt] 
-b_{\phi0_j} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in student}\ t\ \text{model model} \\[4pt]
-b_{\phi0_k} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in student}\ t\ \text{model model} \\[4pt]
-b_{\phi0_l} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in student}\ t\ \text{model model} \\[4pt]
+X'\beta&=(\beta_0 + b_{0_{j[i],k[i],l[i]}}) + \beta_1 \text{Condition}_{ic} & \text{Vector of predictors for mean parameter} \\[4pt] 
+b_{0_j} &\sim \mathcal{N}(0, \sigma_{b_{0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in mean parameter} \\[4pt]
+b_{0_k} &\sim \mathcal{N}(0, \sigma_{b_{0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in mean parameter} \\[4pt]
+b_{0_l} &\sim \mathcal{N}(0, \sigma_{b_{0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in mean parameter} \\[4pt]
+X'\beta_\phi&=(\beta_{\phi0} + b_{\phi0_{j[i],k[i],l[i]}}) + \beta_{\phi1} \text{Condition}_{ic} & \text{Vector of predictors for precision parameter} \\[4pt] 
+b_{\phi0_j} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{j}}}), \text{for location j} = 1,\ ...,\text{J} & \text{Variation in intercepts for location in precision parameter} \\[4pt]
+b_{\phi0_k} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{k}}}), \text{for member k} = 1,\ ...,\text{K} & \text{Variation in intercepts for member in precision parameter} \\[4pt]
+b_{\phi0_l} &\sim \mathcal{N}(0, \sigma_{b_{\phi0_{l}}}), \text{for machine l} = 1,\ ...,\text{L} & \text{Variation in intercepts for machine  in precision parameter} \\[4pt]
 \\
@@
 Other supplementary analyses and plots are also available there.
+
+# References

### 2025-03-04 submission/pre_print_jss_submission_anon.qmd
Update preprint, and add journal submission outputs

--- 2025-03-04 submission/pre_print_jss_submission.qmd
@@
 title: "The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
-author:
-  - name: James Steele
-    affiliations:
-      - ref: 1
-      - ref: 2
-      - ref: 3
-    orcid: 0000-0002-8003-0757
-    corresponding: true
-    email: james@steele-research.com
-  - name: James Fisher
-    affiliations:
-      - ref: 3
-      - ref: 4
-    orcid: 0000-0002-6013-8402
-    email: james.fisher.phd@outlook.com
-  - name: Tim Dettmann
-    affiliations:
-      - ref: 1
-    email: tim.dettmann@gmail.com
-affiliations:
-      - id: 1
-        name: Kieser Australia, Melbourne, Australia
-      - id: 2
-        name: Steele Research Limited, Eastleigh, UK
-      - id: 3
-        name: Department of Sport and Health, Solent University, Southampton, UK
-      - id: 4
-        name: The Exercise Coach, Tennessee, USA
 abstract: |
@@
 ```
-
-## Author Affiliations
-James Steele: Kieser Australia, Melbourne, Australia; Steele Research Limited, Eastleigh, UK; Department of Sport and Health, Solent University, Southampton, UK
-James Fisher: Department of Sport and Health, Solent University, Southampton, UK; The Exercise Coach, Tennessee, USA
-Tim Dettmann: Kieser Australia, Melbourne, Australia
-
-Corresponding author: James Steele (james@steele-research.com)
 
@@
 
-The experimental study component of this project was approved by the Southampton Solent University Health Exercise and Sport Science Ethics committee (id: STEELEAUG2016).
+The experimental study component of this project was approved by the a university ethics committee.
 All participants provided informed consent to participate in the study.
@@
 ## Statistical Analysis
-All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/etb34/> or the corresponding GitHub repository <https://github.com/jamessteeleii/supervision_tul>.
+All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/etb34/>.
 We cite all software and packages used in the analysis pipeline using the `grateful` package [@rodriguez-sanchezGratefulFacilitateCitation2023] which can be seen here: <https://osf.io/ew79g>.
@@
 
-# Contributions
-
-All authors conceived the idea for the project and designed the methods.
-TD coordinated and oversaw the data collection.
-JS conducted the statistical analyses and produced data visualisations.
-JS and JF drafted the initial manuscript.
-All authors contributed to editing the manuscript.
-All authors read and approved the final manuscript.
-
 # Acknowledgements
@@
 
-All data and code utilised for data preparation and analyses are available in either the Open Science Framework page for this project https://osf.io/etb34/ or the corresponding GitHub repository https://github.com/jamessteeleii/supervision_tul.
+All data and code utilised for data preparation and analyses are available in either the Open Science Framework page for this project https://osf.io/etb34/.
 Other supplementary analyses and plots are also available there.

### 2025-05-20 pre_print/response_to_reviewers.qmd
Revisions complete, flow chart added, preprint updated

---
title: "Response to Reviewers"
subtitle: "The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
format: pdf
bibliography: bibliography.bib  
---

Dear Professor Gruet,

Many thanks for considering our manuscript and placing it under review with *Peer Community in Health and Movement Sciences*. We are very grateful for your efforts in obtaining reviews. We are very impressed with the quality and constructiveness of the feedback provided by the reviewers which we feel has significantly improved the quality of our manuscript.

Note, because of the extensive rewrite of the manuscript due to the shift in focus we have not used track changes or highlighting throughout, though changes can be viewed in the GitHub repository commits and versions. However, please see below for our comments in response to some of the key points raised by the reviewers.

Again, we would like to thank all involved in the process of reviewing this manuscript and look forward to the next round of reviews in considering our revisions.

Many thanks

The authors

# Recommender and Reviewer Comments and Author Responses

## Recommender - Mathieu Gruet

### Recommender Comment

<!-- The reviewers have, however, identified several concerns that require your attention before the manuscript can proceed further. Most notably, the analysis methodology and presentation of results would benefit being slightly more accessible. As your research has considerable practical implications for resistance training coaches and strength clinic members, and considering that many of whom may lack the technical expertise to fully comprehend your analytical approach, greater clarity is essential. The actual technical complexity may also present challenges for researchers who wish to replicate or build upon your methods in similar or different contexts. -->

## Reviewer 1 - David Clarke

### Reviewer Comment

Minor comments on the writing:

While the paper is overall well written, I found that the paragraphs tended to be excessively lengthy. There were also a few awkward and run-on sentences and a few instances of colloquial writing that I recommend be cleaned up, as follows:

Page 3: start a new paragraph at “An acute randomized cross-over design…”

Page 4 and elsewhere: “We ended up with n = 45…” – “ended up” seems colloquial to me.

Page 6: The following sentence is awkward and excessively long: “Anecdotal reports from Kieser Australia staff had also prior to exploring the data suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.”

### Author Response

Thank you for your kinds words and for the suggestions. We have made the minor changes noted on page 3, page 4 and elsewhere (noting instead the "final sample size..."), and change the sentence on page 6 to *"Prior to exploring the data we had received anecdotal reports from Kieser Australia staff. These suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure."*

### Reviewer Comment

The repetition tempo and resulting TUL are major components of the resistance exercise prescription, yet the emphasis on these variables is poorly justified. Specifically, on page 3, the authors mention that the clients are prescribed repetitions lasting ~ 12 seconds, which consists of a 4:2:4:2 concentric:isometric:eccentric:isometric tempo. This tempo prescription is questionable for strength development, wherein powerful concentric motions of maximal effort should be emphasized (Spiering et al. 2023 J Str Cond Res). Pausing at the top of the movement (in the “lockout” position) is likewise questionable.

What effects can be expected from this tempo prescription on TUL compared to more common resistance exercise prescriptions? Would these effects affect the generalizability of this study?

To what extent do the clients comply with the tempo prescriptions?

What is the reliability and accuracy of the TUL recordings? The authors stated that “During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose.” For the retrospective analysis, the authors stated that “TUL was also recorded similarly using timers available about the clinic, or for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which is placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.” The self-report nature of this variable could bias the results.

### Author Response

Unfortunately this is not something that was necessarily chosen and justified from a research design perspective. Instead this is just the protocol that Kieser utilise across their facilities. Firstly However, we would note that, whilst most would argue that faster movement speeds are likely to maximise neuromuscular adaptations that contribute to overall strength development it is also not conclusive that repetition durations of those used here are necessarily suboptimal and a range of movement speeds have been found to be effective [@wilkInfluenceMovementTempo2021]. A recent meta-analysis did conclude that intentionally slow repetitions are sub-optimal [@hermesIntentionallySlowConcentric2023] though the dichotomisation of "fast" vs "slow" does limit the interpretation of their results (we are actually in discussions regarding whether to redo this meta-analysis with updated searches and the treatment of repetition duration as a continuous variables with interactions between concentric and eccentric phase durations). But more broadly speaking, the magnitude of standardised effect sizes reported in a recent large scale study from our group utilising another dataset [@steeleLongTermTimeCourseStrength2023a] from a commercial training provider who use an even longer repetition duration (~10:10s) were similar to those seen in large scale meta-analytic data we have collected regarding more typical resistance training programs compared with non-training controls [@steeleMetaanalysisVariationSport2023a]. Thus, it seems unlikely that there would be drastically different effects upon strength even with slower repetition durations. Lastly, it is also thought that the intended rather than actual speed of movement may be more important [@behmNarrativeReviewVelocitybased2025]. We actually intend to more systematically explore this at some point as we are shifting towards a protocol whereby, despite a deliberately slow initial repetition duration, after fatigue accumulates intraset the actual movement speed possible is reduced such that even slow repetition durations can be maintained whilst attempting to move fast.

Also, we did not necessarily have any specific theoretically driven hypotheses about the impact of supervision with specific manipulations of resistance training variables such as repetition duration. So we felt that this ecological dataset and context was suitable for our research questions, particularly given it offered a natural experiment of sorts as Kieser already implements supervised and unsupervised training.

Notably however we do not have data on the extent to which clients actually comply with these prescriptions (though informally from clinical observations the majority do comply with them approximately). We have added a comment in the limitations to note this in addition to also the reliability of the TUL recordings. The latter we are less concerned with however given that the sample size of the previous data is large (any measurement error due to self-report, assuming no systematic bias, will only affect the error term and thus precision of estimates), and that this is also the case with the experimental dataset when considering the multiple exercises utilised too.

*"Firstly, though anecdotal observations in the clinic generally confirm compliance broadly speaking, we do not have systematically captured data on whether the repetition durations prescribed (or other aspects of the prescribed training protocol) are complied with by members. Further, the measurement error for the self-recorded or exercise scientist recorded TUL is not known; though the large sample size for the previous data combined with the multiple observations per person in our experimental dataset minimised any non-systematic measurement error as this would merely affect the precision of estimates."*

### Reviewer Comment

Consider justifying the use of Bayesian inference over alternative frequentist methods. This justification should be upfront, perhaps in the Introduction or in the Approach to the Problem section of the Methods. In this specific case, the Bayesian methods are powerful for studying in an integrated manner the experimental and observational data. The authors should better emphasize this point as well as other benefits that the Bayesian approach uniquely afforded.

### Author Response

We have added the following to the beginning of the statistical methods section:

*"All analyses have been conducted within a Bayesian framework and are primarily focused upon estimation of the relevant parameters of interest from our models [@kruschkeBayesianNewStatistics2018]. A Bayesian framework was used specifically because it allows for the explicit updating of beliefs regarding prior probable values for parameters of interest with new evidence (i.e., data) to form updated posterior beliefs regarding probable parameter values. Given that we had prior empirical data from which we could generate our prior beliefs regarding probable values for parameters of interest, and we had collected new experimental data that could be used to update these, the Bayesian framework is ideally suited to this type of inferential problem. Contrastingly, within traditional frequentist inference prior beliefs are not specifically incorporated in this fashion. Further, our choice to utilise Bayesian approaches was also based upon the specific modelling choices made and the flexibility of current Bayesian software packages such as those used here to implement such models."*

### Reviewer Comment

Consider including a flowchart for the statistical analysis and then describing each element in the text. For example, something that communicates the following: Priors for observational data > posterior distributions for observational data > priors for experimental > posteriors experimental, etc. Relate the flowchart elements to the results, i.e., from which points different estimates were obtained.

### Author Response

This is a great suggestion and we have added a flow chart color coded to hopefully relate the results to the different stages of our modelling strategy.

### Reviewer Comment

Define all letters and symbols for the non-expert. For example, the degrees of freedom for the TUL model is specified as νy, but the “y” is undefined in the manuscript. Perhaps knowledge of y is obvious to the expert, but it eluded me.

### Author Response

Apologies, this we realise is not so clear in hindsight, but the $y$ subscript refers to the fact that this is a singular parameter estimated for the whole model. We have updated the entire statistical analysis section and included details for each model notation, and also tried to include a brief "lay" description of what the model is in essence doing to try to aid the reader in understanding.

Given this is a scientific manuscript though (and frankly, most researchers and practitioners don't fully read the analyses section most the time anyway... and I often question the extent to which many really understand the typical "basic" analyses they are more used to seeing), we have placed this in the online supplementary materials and included a footnote at the beginning of the statistical analysis section that links to this and reads:

*"Note, we thank the reviewers for noting the complexity of the analyses performed in the present study relative to what both researchers and practitioners are used to encountering in the field. We have done our best to provide both detailed explanation and justification of the choices made, describe what the models being used are doing, provided mathematical notation for them, and all code for the analyses conducted is available. We have also included in the online supplementary material a brief "lay" summary which was generated making use of ChatGPT to assist in developing these explanations. We prompted ChatGPT explicitly by providing it with the copied text, including equations, from our quarto manuscript files and asked it to provide a lay description. We then edited these. These can be seen here <https://osf.io/b9n4v>."*

### Reviewer Comment

On page 7, in the priors for γ1 and β1 in the TUL model, it is stated “lb = -60, lb = 60.” Should the second lb be “ub” for upper bound?

### Author Response

Thanks for spotting this. It has been updated.

### Reviewer Comment

The results are effectively described and visualized. However, the figure captions should begin with a title sentence, followed by the text that was written. Panel labels (a, b, c, etc.) should also be included.

### Author Response

As the figures already contain titles in the plots themselves we have retained only the caption. The purpose of having titles on the figures is to faciliate their reuses by others in presentations etc without having to use the figure caption. We have added the panel labels though for each row and column.

### Reviewer Comment

Throughout the manuscript but particularly in the Results, the authors use the term “prior sample” to describe the retrospective observational data. This terminology might cause confusion because the word “prior” also refers to the prior distributions needed for Bayesian modelling. I suggest renaming the sample “observational sample” or some other alternative.

### Author Response

We have updated it throughout to instead read "previous observational sample". We have also removed the use of "prior" in other instances that might cause confusion e.g., "...prior training...", "... prior research..."

### Reviewer Comment

The phrase “the 𝜙 parameter for precision of the beta distribution in our experimental sample was higher in the Assisted condition suggesting that <variable> reported was more clustered around the estimated <variable> values as compared to in the Core condition” was repeated three times, once for each of the outcome variables. While informative, this phrase reflects an interpretation and it is questionable as to whether it belongs in the Results. If the authors decide to include it, include it only once because it is repetitive.

### Author Response

We have removed this sentence as it is mentioned that this parameter relates to clustering of responses in the discussion already.

### Reviewer Comment

I proposed above some methodological questions and comments that the authors may wish to address in the revised Discussion. Another potential limiter to generalizability of this study is that the exercises are all machine based, and, as the authors noted in their previous meta-analysis, a major potential benefit of supervision is technical coaching (Fisher 2022 Int J Str Cond). If participants were to use free weights, they may further “underload” for safety considerations and for not knowing the “proper technique.” I encourage the authors to consider commenting on this aspect.

### Author Response

This is a good point and we have added a comment in the limitations regarding this - *"Additionally, the protocols were all performed using resistance machines thus minimising one potential element of supervision; technical coaching [@fisherRoleSupervisionResistance2022; @fisherSupervisionStrengthTrainingthe2025]. Thus the generalisability to other resistance training modalities is unclear."*

### Reviewer Comment

·       Page 13: The following sentence is awkward and excessively long: “Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.”

### Author Response

The sentence itself is quite short, though the additional content in parentheses makes it appear longer. Upon re-reading the text in parentheses though we realise that was awkward so have re-written it slightly - *"...(i.e., single sets of RT using resistance machines, performed to momentary failure using a fixed repetition duration and load that should permit momentary failure in a target TUL range of 90-120 seconds)..."* - we have kept it though as it serves as a reminder for the reader of the protocols used when they first beging reading the discussion.

## Reviewer 2 - Anonymous

### Reviewer Comment

The aim of the current manuscript was to examine the role of supervision on effort, defined here as time under load (TUL), rating of perceived exertion (RPE), and rating of perceived discomfort (RPD), during resistance training, using both retrospective and prospective analyses. The manuscript is well written and structured. I would particularly like to highlight the ecological validity of the approach. Randomized clinical trials often struggle to replicate real-world settings, so the chosen study design strengthens the external validity of the findings.

### Author Response

Thank you for the kind words.

### Reviewer Comment

The objective provided at the end of the introduction could be clearer. At that point in the text, "effort" is not well defined. Since the manuscript later defines effort as TUL, RPE, and RPD, it would be helpful to clarify this earlier in the manuscript.

<!-- ### Author Response -->

We do in the introduction note that, in resistance training, effort is proxied by proximity to momentary failure but have added this more explicitly now. *"Effort is conceptualised at the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance."*

### Reviewer Comment

While I understand the logistical reasons given for the sample size in the experimental analysis, but the relatively small sample may limit the generalizability of the findings considering the approach of real-world setting. It would be appropriate to include this as a limitation of the study.

### Author Response

We would argue that the sample size is in fact still relatively large compared to most studies in the field and that generalisability is not the primary concern regarding sample size here and we did have participants from across varied locations as well. However the first reviewer did not that, despite us having a varied selection of resistance machines enhancing the generalisability of effects in that regard, we did have only machine based exercise and there may be theoretical reasons to suspect our findings may not generalise to other modalities such as free weights. We have added this as a limitation.

### Reviewer Comment

Regarding the retrospective analysis, it is unclear why the authors chose two separate samples (one with TUL and another with RPE scores) instead of selecting a sample in which both variables were available. Since the aim was to analyze the influence of supervision on effort (as reflected by both TUL and RPE), could separate samples introduce bias (as no confounding factors could be taken into account)? I am not sure about this.

### Author Response

We do not in the methods why two separate samples were taken. This was due to not all members using the RPE recording function in their training app. Given we also did not have any explicit causal model from which to select covariates for adjustment, and that our primary use for the prior samples was to generate informative priors, we felt that the random samples achieved this without having to identify a more selective sample of only those who had recorded RPE.

### Reviewer Comment

The manuscript notes that participants recorded their own TUL using timers visible in the training space. While this aligns with the goal of examining real-world settings, it would be helpful to address the validity and reliability of this method. Could discrepancies between trainee- and trainer-reported data have affected the results? Were trainers generally more rigorous in data collection?

<!-- ### Author Response -->

This is definitely a limitation and one which reviewer 1 also noted. We have included this in our limitations section now to make the reader aware of this. However, notably from a reliability perspective this would primary influence parameter estimate precision and so the sample sizes, use of prior-posterior updating, enhances our precision here even in the face of such error. Also, if we were to speculate, we would think that the trainees were probably more likely to be biased upwards in their TUL recording (having to stop the exercise and then stop the timer on their app... though if using the timers in the clinics they can merely see these and note mentally what time they achieved to then record), but yet we still see at least in the experimental data that far greater TUL is acheived with supervision (and even in light of the very strong priors for there being no supervision effect based on the previous sample data).

### Reviewer Comment

How was momentary failure defined and monitored? Was it self-determined by participants in the Core group? If so, how might this have influenced the outcomes?

### Author Response

The fact that it is not explicitly monitored is part of the purpose of the study. In both the *in situ* settings of the clinic in terms of the protocols prescribed, and in the experimental study, members are informed that momentary failure is defined as per Steele et al. [-@steeleClarityReportingTerminology2017] and that this is the endpoint for their exercises. In this study, TUL and RPE offer some insight into the extent to which members are adhering to this. For TUL we'd expect, assuming following the load progression rules, some distribution of TUL between 90-120 seconds and not hugely exceeding this nor any spikes. For RPE we'd expect them to be providing maximal ratings if they are indeed training to momentary failure. Thus these are used as proxies to infer effort and also the effects of supervision upon this. We already try to highlight this also in the discussion where we note that the difference in TUL for Core conditions between the previous observational sample and the experimental sample may also express that *in situ* members, despite being prescribed to do so, are not training to momentary failure. But in the experimental setting where they knew they were part of a study they trained with a greater degree of effort indicated by greater TUL and RPE.

### Reviewer Comment

Were the Core and Assisted groups in the retrospective analysis performing the same exercises? If different exercises were used, this could have impacted TUL and RPE outcomes.

### Author Response

Exercises varied from member to member in both the previous observational sample and the experimental study data. However, this has been factored into the modelling allowing for random effects by exercise/machine in order to produce an overall average treatment effect (ATE) marginalised over these. We did not have any specific hypotheses around these though so did not explore them, but the random effects for machine suggested there was very little effect with standard deviation of only ~3-4 seconds TUL between machines.

### Reviewer Comment

For the retrospective data, how exactly was TUL entered into the Kieser database, especially for Core participants? Did participants self-report their TUL using a stopwatch?

### Author Response

We note already in the methods under protocols that the data captured either using the timers (i.e., Swiss engineered clocks with sweep movement for the second hand and positioned so that there is always one in eye line for every machine to enable recording of the TUL in seconds), mobile phone app, or trainers app is input to the training card which is then recorded and stored in the Kieser Australia Database. We have added some additional information now including the description of the timers noted above.

### Reviewer Comment

Is comparing only one supervised and one unsupervised session, one week apart, sufficient to draw firm conclusions? Given the potential for a Hawthorne effect (i.e., participants modifying behavior due to awareness of being observed), a single-session design may overestimate the true effect of supervision.
Could repeated sessions have offered a more stable measure?

### Author Response

This is certain to be the case that there will be differential effects over time for the effects of supervision, as would be the case for any training manipulations. As noted inn the discussion already, we are in the process of developing longitudinal research to examine this specifically though. In fact we are in the process currently of developing the protocols and analysis plans for an interrupted time-series with control design with internal replication in order to examine the effects in our members who have changed contracts over time i.e., went from Core to Assisted and vice versa. It is however a complex design and analysis and so, given the already complex approach taken here, we have kept this as a separate piece of follow on work to build further on this area.

### Reviewer Comment

If participants were not blinded to their TUL, is it possible that increased TUL influenced their reported RPE and RPD? Could it influence the outcomes considering the cross-over design?

### Author Response

This certainly could be a confounder for RPE/RPD responses, however this could be a confounder irrespective of whether participants knew their exact TUL or not assuming they have a reasonable perception of time i.e., they could tell if they went longer/shorter than a previous experience they could anchor against and the respond with RPE/RPD as either higher/lower correspondingly. But the randomised cross-over design for the experimental study at least allows for any effect of this to be removed.

### Reviewer Comment

The lack of significant differences in TUL in the retrospective analysis contrasts with findings from the experimental condition. The authors suggest this may be due to more appropriate load prescription in the Assisted group; however, this interpretation is not supported by the retrospective RPE data. If Assisted participants truly trained with a higher (and more optimal) load, one would expect their RPE scores to be notably higher, yet this was not observed.

### Author Response

This interpretation is more speculative and in reference to the Core vs Assisted contrast in the previous observational sample only. However, the Core vs Assisted RPE contrasts are actually of similar magnitudes in both the previous observational sample and the experimental data. So whilst the absolute magnitudes are still low for the previous sample, the difference between conditions is still similar to that seen in the experimental data. So, it is perhaps plausible that more appropriate loads during Assisted compared to Core may be a reason for this even if they are still suboptimal. But we do also somewhat test this explanation and note the exploratory examination of some of our load progression data which suggests that load progressions may be similar in the footnote. As noted above, we do plan to more carefully examine the longitudinal causal effects of supervision in our members looking at both load progression, and also the isometric strength test data we routinely collect which we believe may better tease out this as a possible explanation for the *in situ* results observed.

### Reviewer Comment

Given the limitations presented by the authors, the results provided by the retrospective analysis, and the effect analyzed during the experimental analysis about the comparison of only one session of supervised versus unsupervised one-week apart, the conclusion that “these results in general support prior work highlighting the importance of supervision during RT” seems overstated. While I agree with the authors onthe potential influence of the Hawthorne effect, a more cautious and nuanced interpretation would better reflect the study’s findings and limitations.

### Author Response

Given that we are cautious in our interpretation of the previous observational data results alone, but that the experimental study provides us with causal estimates due to the randomisation and cross-over design, we do not feel that our results are overstated. The experimental findings do support the effects of supervision, but the contrast with the seeming lack of difference in TUL, and the smaller absolute RPEs, in *in situ* data suggests that this effect is difficult to causally identify against the background of potential confounders in observational data. Again, this is why in our next planned staged of this programme of work we are engaging with more appropriate causal designs such as the interrupted time-series with control design [@lopezbernalUseControlsInterrupted2018].

### Reviewer Comment

No conflict of interest is reported, yet the study analyzes the effect of the Kieser Method, a supervised high-intensity training model developed by the Kieser Training Company, with which two of the authors are affiliated. Given this connection, could there be a potential bias in the interpretation or preference for certain results? While this does not undermine the value of the study, acknowledging and addressing this possibility would strengthen the transparency and integrity of the manuscript.

### Author Response

Thank you for noting this. We apologies for not being more explicit. We had mentioned that Kieser had provided in-kind funding for this project and our affiliations as you note are clearly aligned. But we have added a section to be more clear regarding this. We also hope that our transparency in terms of the data, materials, code, and indeed the results (which to some extent suggest that our in clinic protocols are not in fact being adhered to... something we have fed back internally and are looking at how to address and also evaluate prospectively the approaches we take to this) also highlight that we are trying to minimise any affiliational bias as much as possible.

## Reviewer 3 - Anonymous

### Reviewer Comment

I have gone through the study and need the following recommendations for the final decision

### Author Response

Many thanks for your review.

### Reviewer Comment

1. The abstract is well-written and explained, but it would benefit from being divided into distinct sections, such as background, methodology, results, and conclusion.

### Author Response

We have not made this change for now as the journals we may potentially submit to upon recommendation (either Communications in Kinesiology or PeerJ) do not typically require this format. If a journal subsequently requires this we will make the change in line with their requirements.

### Reviewer Comment

2. There is only one line about RT. A few more lines at the start of the abstract would be beneficial for the reader.

### Author Response

The abstract is already quite lengthy and hence we have kept this general information to a minimum opting to get straight to summarising the study itself. The introduction provides more exposition on the broader topic.

### Reviewer Comment

3. The authors used the word "little" in "exercise performance differed little between supervised and unsupervised training." It would be better to use some technical words because it shows subjectivity rather objectivity. What is meant by little can be explained numerically.

### Author Response

It is hard to avoid using imprecise descriptors of the magnitude of difference and impossible to avoid making some subjective determination of whether a result is big enough or small enough to care or not care about relative to typical values of the outcome and other background information. Though, the manuscript and abstract provide the precise numerical estimates for the values we are admittedly subjectively interpreting here and so any reader can interpret these themselves deciding if they agree with the difference being "little".

### Reviewer Comment

4.The introduction is well written and it contains latest studies. but it should be extended more.

### Author Response

Without suggestion of what could be included to extend it we have not added anything further. The manuscript is already quite long and we believe we have provided the background as succinctly as possible in the introduction already and do not require any further detail.

### Reviewer Comment

5.The software used for the study should also be mentioned.

### Author Response

Any specific software used for protocols such as the mobile phone app is already mentioned, and we link out to an extensive list generated using the `grateful` package already containing all of the software used for analysis including version numbers - https://osf.io/ew79g. This is also included in the `renv` lockfile such that anyone reproducing our analyses can utilise the exact versions we had utilised.

### Reviewer Comment

6.The result section should be explained for the general readers to understand the technical outcomes. It would be better to make statements simple while explaining technical words.

### Author Response

As suggested by reviewer 1 we have added a lay explanation of the models, and have also added a visual to help with interpretation of the models and results.

### Reviewer Comment

7.The authors mentioned about the confidence intervals and quantile intervals. Are there any more measures to show the strength of outcomes like effect size ?

### Author Response

This is a common misonception in the field; namely that by *effect size* something like a standardised effect size e.g., Cohen's d is meant. All of the point estimates and the corresponding interval estimates are in fact effect size estimates in and of them selves and in the units of interest to a reader i.e., seconds, probability of stopping, percentage effort/discomfort. So the strength of the outcomes can be interpreted from these directly.

### Reviewer Comment

8.Overall, the study is new and well articulated and would provide reader a quality study.

### Authors Response

Thank you for the kinds words.

### Reviewer Comment

9.Actual real life contribution of this study should be highlighted in the conclusion section. What it can bring in real life?

### Authors Response

We would prefer to leave out inclusion of "real life contribution" such as practical recommendations from this work. In fact, the typical inclusion of practical recommendations by many applied journals in our field from a single study always feels unwarranted and results in overstretching of conclusions. Neither of the journals we are considering to submit to require practical recommendations of this kind. We would note in response though that a very real outcome from this work has for us at Kieser Australia to internally consider how we might adapt our operational procedures to try and ensure that our protocols as prescribed are in fact being followed by members and trainers. Our future work will look to evaluate this more closely as we implement changes.

# References

### 2025-05-20 pre_print/pre_print_v2.qmd
Revisions complete, flow chart added, preprint updated

--- 2025-02-18 pre_print/pre_print.qmd
@@
 articletitle: |
-  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
+  The effects of supervision upon effort during resistance training: A Bayesian analysis of previous observational data and an experimental study of private strength clinic members
 # shorttitle: "Doe et al. (2022)" # usually auto-generated
@@
 Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study.
-The prior sample included ~1000 members training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
+The previous observational sample included ~1000 members training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
 Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training.
 However, in our experimental study, supervision improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
-In percentage points RPE was slightly higher with supervision in both prior real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
+In percentage points RPE was slightly higher with supervision in both previous observational real-world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; contrast = -6% [95%QI: -8%, -4%]) and experimental settings (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; contrast = -6% [95%QI: -10%, -4%]), suggesting trainees push closer to failure under supervision.
 This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]).
-Overall, these results reinforce prior research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
+Overall, these results reinforce research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
 license-type: ccby # change if neccessary
@@
 It has been speculated [@fisherRoleSupervisionResistance2022] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort.
-Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
+Effort is conceptualised at the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance.
+
+A recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
 Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., they select \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024].
@@
 When considering the average self-selected load of \~53% 1RM coupled with a typical \~10 repetitions per set equating to proximities to failure \>10 repetitions [@steeleAreTraineesLifting2022] it seems plausible that unsupervised trainees may train with a less-than-optimal load, repetitions, and resultant effort to optimise strength outcomes.
-Ultimately, and in combination, supervision should provide trainees with feedback of their past performance, whether technical, effort- or program-based, and guidance toward their future performance (Fisher, under review).
+Ultimately, and in combination, supervision should provide trainees with feedback of their past performance, whether technical, effort- or program-based, and guidance toward their future performance [@fisherSupervisionStrengthTrainingthe2025].
 
@@
 All Kieser Australia members are prescribed the same protocol.
-This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action).
+This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension, not "locked-out", on the involved musculature at the end of each concentric and eccentric muscle action).
 Core trainees are prompted to progress load by \~5% for the next session once a TUL of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist.
 Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session.
+
 An acute randomised cross-over design was employed to examine participants completing an RT session with and without direct supervision with one week apart (i.e., a Core vs Assisted session).
@@
 All participants provided informed consent to participate in the study.
-Data in the prior sample was used in de-identified form and with member consent for data to be used for research purposes.
+Data in the previous observational sample was used in de-identified form and with member consent for data to be used for research purposes.
 
@@
 
-The study was advertised at five Kieser Australia strength clinics and we sought to recruit members who had \> 6 months prior training experience at the strength clinics, were healthy (no clinical conditions on their member record as recorded by a physiotherapist or exercise physiologist), both males and females (age +18 years), and without any current condition for which RT would be contraindicated.
+The study was advertised at five Kieser Australia strength clinics and we sought to recruit members who had \> 6 months previous training experience at the strength clinics, were healthy (no clinical conditions on their member record as recorded by a physiotherapist or exercise physiologist), both males and females (age +18 years), and without any current condition for which RT would be contraindicated.
 They were recruited from the existing member pool of Core members.
@@
 As such it was decided that a target of 50 participants across clinics would be acceptable.
-We ended up with n = `r length(unique(data$id))` participants recruited across the five clinics.
+The final sample size was n = `r length(unique(data$id))` participants recruited across the five clinics.
 
-### Prior Sample
+### Previous Observational Sample
 
@@
 We opted to generate reasonably large samples to ensure precise prior estimates, though did not use the entirety of the historical data (\~50000 members, \>10 years of data) so as to reduce computation time for Bayesian modelling.
-For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.
-Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`).
+For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of previous training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.
+Thus our final sample size was n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`).
 For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.
@@
 Each participant had an existing training programme card.
-In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card prior to participation in the study.
-Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the TUL achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL acheived.
+In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card before participation in the study.
+Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the TUL achieved for the purposes of the study i.e., if they realised during the set that their previous load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL acheived.
 Momentary failure was defined as per Steele et al.
 [-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was clearly communicated to the participants.
-During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose.
+During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using standard timers situated around the clinic in view during training specifically for this purpose (these are essentially Swiss engineered clocks with sweep movement for the second hand and positioned so that there is always one in eye line for every machine to enable recording of the TUL in seconds).
+These TUL recordings where then input to their training cards.
 In addition, they recorded their RPE and RPD in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).
@@
 The data generated from the sample of historical members was generated according to the standard protocol described above that all Kieser Australia members are prescribed.
-TUL was also recorded similarly using timers available about the clinic, or for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which is placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.
+TUL was also recorded similarly using timers available about the clinic which are then entered onto their training cards, or for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which is placed on a stand on the resistance machines with a start/stop recording button used to track TUL for each exercise in their training card which is programmed to the application.
 Session RPE was recorded using either the mobile application for Core members, or via the application used by staff during Assisted sessions and used a modified version of the Borg 6-20 scale where the scale is displayed as scrollable and the verbal anchors are "6 - None", "7- Very, very light", "13 – Somewhat hard", "19 - Very, very hard", and "20 – Maximal Exertion".
+All of this data is directly collected on training cards manually, or via the mobile phone app, or trainers app, is recorded and stored in the Kieser Australia backend database.
 
-## Statistical Analysis
+## Statistical Analysis[^4]
+
+[^4]: Note, we thank the reviewers for noting the complexity of the analyses performed in the present study relative to what both researchers and practitioners are used to encountering in the field.
+We have done our best to provide both detailed explanation and justification of the choices made, describe what the models being used are doing, provided mathematical notation for them, and all code for the analyses conducted is available.
+We have also included in the online supplementary material a brief "lay" summary which was generated making use of ChatGPT to assist in developing these explanations.
+We prompted ChatGPT explicitly by providing it with the copied text, including equations, from our quarto manuscript files and asked it to provide a lay description.
+We then edited these.
+These can be seen here <https://osf.io/b9n4v>.
+
 All code utilised for data preparation and analyses are available in either the Open Science Framework page for this project <https://osf.io/etb34/> or the corresponding GitHub repository <https://github.com/jamessteeleii/supervision_tul>.
@@
 
-As noted, the project was not pre-registered but involved exploratory analysis of the experimental and prior datasets.
-All analyses have been conducted within a Bayesian framework and all posterior estimates and their precision, along with conclusions based upon them, are interpreted continuously and probabilistically, considering priors, data quality, and all within the context of each outcome and the assumptions of the model employed as the estimator [@kruschkeBayesianNewStatistics2018].
+As noted, the project was not pre-registered but involved exploratory analysis of the experimental and previous observational datasets.
+All analyses have been conducted within a Bayesian framework and are primarily focused upon estimation of the relevant parameters of interest from our models [@kruschkeBayesianNewStatistics2018].
+A Bayesian framework was used specifically because it allows for the explicit updating of beliefs regarding prior probable values for parameters of interest with new evidence (i.e., data) to form updated posterior beliefs regarding probable parameter values.
+Given that we had previous observational data from which we could generate our prior beliefs regarding probable values for parameters of interest, and we had collected new experimental data that could be used to update these, the Bayesian framework is ideally suited to this type of inferential problem.
+Contrastingly, within traditional frequentist inference prior beliefs are not specifically incorporated in this fashion.
+Further, our choice to utilise Bayesian approaches was also based upon the specific modelling choices made and the flexibility of current Bayesian software packages such as those used here to implement such models.
+To aid the reader in understanding the flow of our modelling strategy we refer them to @fig-flow which is also colour coded to reflect the prior and posterior models in the results section
+
+```{r}
+#| label: fig-flow
+#| fig-width: 10
+#| fig-height: 7.5
+#| fig-cap: Flow chart displaying the modelling strategy employed in the present study incorporating both previous observational and experimental study data (grey), modelling of the previous study data using weakly informative priors (white) to produce prior models and estimates (light blue) to use in modelling the experimental data (orange).
+
+targets::tar_load(flow_chart)
+
+flow_chart
+
+```
+
+All posterior estimates and their precision, along with conclusions based upon them, are interpreted continuously and probabilistically, considering priors, data quality, and all within the context of each outcome and the assumptions of the model employed as the estimator [@kruschkeBayesianNewStatistics2018].
 All models were run with 2000 warmup and 2000 sampling iterations and four Monte Carlo Markov Chains.
 Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions (see <https://osf.io/u7g8c>.
-We fit two sets of models for the TUL and RPE outcomes, one on the prior sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the RPD outcome using weakly regularising priors.
+We fit two sets of models for the TUL and RPE outcomes, one on the previous observational sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the RPD outcome using weakly regularising priors.
 These models are described below.
@@
 
-Upon inspecting the raw distribution of the prior sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (see @fig-raw).
-Anecdotal reports from Kieser Australia staff had also prior to exploring the data suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.
-We did not however expect this to be the same for the prior sample of Assisted members.
+Upon inspecting the raw distribution of the previous observational sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (see @fig-raw).
+Before to exploring the data we had received anecdotal reports from Kieser Australia staff.
+These suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.
+We did not however expect this to be the same for the previous observational sample of Assisted members.
 As such, it suggested to us that there might be two processes underlying the observed data: a tendency to target a specific TUL threshold and to stop the exercise at that point, or to otherwise continue the exercise to momentary failure (or to stop the exercise based on some other process such as tolerable proximity to failure).
-Thus, we opted to use a hurdle type model comprising a Bernoulli distribution for the probability of having a TUL of 120 seconds, and a student $t$ distribution for all other values of TUL[^4].
+Thus, we opted to use a hurdle type model comprising a Bernoulli distribution for the probability of having a TUL of 120 seconds, and a student $t$ distribution for all other values of TUL[^5].
 We expected this may also be the case in our experimental dataset given we recruited Core members and so utilised the same type of model for this.
@@
 
-[^4]: A custom family for the hurdle-student $t$ model was produced for use with the `brms` R package, adapted from the hurdle-normal distribution developed by Heiss [-@heissGuideModelingOutcomes].
+[^5]: A custom family for the hurdle-student $t$ model was produced for use with the `brms` R package, adapted from the hurdle-normal distribution developed by Heiss [-@heissGuideModelingOutcomes].
 
-For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept.
+For both the previous observational sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept.
 The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used (note, due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset).
 These fixed and random effects were modelled as predictors for both the hurdle and student $t$ components of the model.
-TUL was centred at zero for modelling (i.e., all values of TUL$_i$ has 120 subtracted from them prior to modelling).
+TUL was centred at zero for modelling (i.e., all values of TUL$_i$ has 120 subtracted from them before to modelling).
 Formally the model(s) for $i^{th}$ TUL across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows:
@@
 \begin{aligned}
-\text{TUL}_{ic} &\sim \operatorname{Hurdle\,log-student}t(\pi_{ic_{j[i],k[i],l[i]}}, \nu_y, \mu_{ic_{j[i],k[i],l[i]}}, \sigma_y) \ ...\ \text{or alternatively,} \\
+\text{TUL}_{ic} &\sim \operatorname{Hurdle\,log-student}t(\pi_{ic_{j[i],k[i],l[i]}}, \nu, \mu_{ic_{j[i],k[i],l[i]}}, \sigma) \ ...\ \text{or alternatively,} \\
 \text{TUL}_{ic} &\sim
@@
 0 & \text{with probability } \pi_{ic_{j[i],k[i],l[i]}} \\[4pt]
-\operatorname{Student}t(\mu_{ic_{j[i],k[i],l[i]}}, \nu_y, \sigma_y) & \text{with probability } 1 - \pi_{ic_{j[i],k[i],l[i]}}
+\operatorname{Student}t(\mu_{ic_{j[i],k[i],l[i]}}, \nu, \sigma) & \text{with probability } 1 - \pi_{ic_{j[i],k[i],l[i]}}
 \end{Bmatrix}
@@
 $$ {#eq-hurdle-model}
-Weakly regularising priors were employed for prior sample dataset.
+where, for $\textrm{logit}(\pi_{ic})$ i.e., the hurdle portion of the model, $\pi_{ic}$ refers to the probability of TUL being 120 seconds, $\gamma_0$ refers to the intercept for $\textrm{logit}(\pi_{ic})$ which in this case is the Core condition, $\gamma_{0_{j[i],k[i],l[i]}}$ refers to the relevant random intercept for $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine, and $\gamma_1\textrm{Condition}_{ic}$ is the difference between the Core and Assisted conditions.
+Similarly for $\mu_{ic}$ i.e., the student $t$ portion of the model, $\mu_{ic}$ refers to the mean TUL (i.e., location parameter), $\beta_0$ refers to the intercept for $\mu_{ic}$ which in this case is the Core condition, $b_{0_{j[i],k[i],l[i]}}$ refers to the relevant random intercept for $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine, and $\beta_1\textrm{Condition}_{ic}$ is the difference between the Core and Assisted conditions.
+The $\sigma$ term refers to the standard deviation for the student $t$ portion of the model, and the other $\sigma$ terms refer to the standard deviation of the random effects for their respect models (i.e., standard deviation in intercepts for probability of stopping at 120 seconds and TUL for location, member, and machine).
+Finally, $\nu$ refers to the estimated degrees of freedom for the student $t$ portion of the model.
+
+<!-- To provide a simple description of what this model does; the model first estimates the probability of whether a given TUL from a given person training on a given machine at a given location is equal to 120 seconds.
+It also estimates If the TUL is not equal to 120 seconds then it contributes to estimating  -->
+
+Weakly regularising priors were employed for previous observational sample dataset.
 These were the default priors in the `brms` package for all intercept terms which are scaled to the expected response value when all predictors are at their means and use a student $t$ distribution with degrees of freedom $\nu=3$, and variance terms similarly scaled with $\mu=0$.
@@
 \beta_0 &\sim \text{Student}\ t(3, 3, 13.3) & \text{Prior for intercept in student}\ t\ \text{model i.e., Assisted condition} \\
-\gamma_1,\beta_1 &\sim \text{Student}\ t(3, 0, 15, \text{lb}=-60, \text{lb}=60) & \text{Prior for Condition effects in both models i.e., Core minus Assisted condition} \\
-\sigma_y,\sigma_{\gamma_{0_{j}}},\sigma_{\gamma_{0_{k}}},\sigma_{\gamma_{0_{l}}},\sigma_{b_{0_{j}}},\sigma_{b_{0_{k}}},\sigma_{b_{0_{l}}} \ &\sim \text{Student}\ t(3, 0, 13.3, \text{lb}=0) & \text{Prior for all variability parameters} \\
-\nu_y &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
+\gamma_1,\beta_1 &\sim \text{Student}\ t(3, 0, 15, \text{lb}=-60, \text{ub}=60) & \text{Prior for Condition effects in both models i.e., Core minus Assisted condition} \\
+\sigma,\sigma_{\gamma_{0_{j}}},\sigma_{\gamma_{0_{k}}},\sigma_{\gamma_{0_{l}}},\sigma_{b_{0_{j}}},\sigma_{b_{0_{k}}},\sigma_{b_{0_{l}}} \ &\sim \text{Student}\ t(3, 0, 13.3, \text{lb}=0) & \text{Prior for all variability parameters} \\
+\nu &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
 \end{aligned}
@@
 
-We then took posterior parameter estimates from the prior sample model and used these to generate informative priors for the experimental data models which were as follows:
+We then took posterior parameter estimates from the previous observational sample model and used these to generate informative priors for the experimental data models which were as follows:
 
@@
 \beta_1 &\sim \text{Student}\ t(3.28, 0.07, 0.58) & \text{Prior for Condition effects in student}\ t\ \text{model i.e., Core minus Assisted condition} \\
-\sigma_y &\sim \text{Student}\ t(3, 10.59, 0.19, \text{lb}=0) & \text{Prior for residual variability parameter} \\
+\sigma &\sim \text{Student}\ t(3, 10.59, 0.19, \text{lb}=0) & \text{Prior for residual variability parameter} \\
 \sigma_{\gamma_{0_{j}}} &\sim \text{Student}\ t(3, 0.31, 0.08, \text{lb}=0) & \text{Prior for location variability parameter in hurdle model} \\
@@
 \sigma_{b_{0_{l}}} &\sim \text{Student}\ t(3, 3.72, 0.93, \text{lb}=0) & \text{Prior for machine variability parameter in student}\ t\ \text{model} \\
-\nu_y &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
+\nu &\sim \text{Gamma}(2,1) & \text{Prior for student}\ t\ \text{degrees of freedom} \\
 \end{aligned}
@@
 
-Despite the use of two different scales and protocols for data collection of perception of effort in our samples i.e., a session RPE using the 6-20 Borg scale for the prior sample and immediately post each exercise RPE using a 0-10 scale, both scales were anchored at their limits as in essence no effort (6 or 0) or maximal effort (20 or 10).
-As such, and as the intention was to utilise the prior sample data to assist in producing empirical prior distributions for our experimental data models, we opted to rescale both outcomes to lie on the $(0,1)$ interval such that they reflected the percent effort that was reported as perceived [@steeleWhatPerceptionEffort2020].
+Despite the use of two different scales and protocols for data collection of perception of effort in our samples i.e., a session RPE using the 6-20 Borg scale for the previous observational sample and immediately post each exercise RPE using a 0-10 scale, both scales were anchored at their limits as in essence no effort (6 or 0) or maximal effort (20 or 10).
+As such, and as the intention was to utilise the previous observational sample data to assist in producing empirical prior distributions for our experimental data models, we opted to rescale both outcomes to lie on the $(0,1)$ interval such that they reflected the percent effort that was reported as perceived [@steeleWhatPerceptionEffort2020].
 We then employed an ordered beta regression model [@kubinecOrderedBetaRegression2022] which employs a cutpoint process similar to ordered logistic models in order to model both the continuous responses on the $(0,1)$ interval reflecting some effort though not maximal, and the degenerate response on the bounds $[0,1]$ reflecting both no and maximal effort respectively.
@@
 
-For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept.
-The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used for the experimental data model only (note, due to the session RPE reported in the prior sample data and that we sampled only one session we do not have member or machine random effects, and due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset).
+For both the previous observational sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept.
+The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used for the experimental data model only (note, due to the session RPE reported in the previous observational sample data and that we sampled only one session we do not have member or machine random effects, and due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset).
 These fixed and random effects were modelled as predictors for both the mean and precision parameters of the ordered beta regression model.
 As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval.
-Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows (note that the model for the prior dataset omits the random effect for machine):
+Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows (note that the model for the previous observational dataset omits the random effect for machine):
 
@@
 $$ {#eq-ordbeta-model}
+where, $\alpha_{ic}$ refers to the probability of the rescaled RPE being 0, $\delta_{ic}$ refers to the probability of the rescaled RPE being in the interval $(0,1)$, and $\gamma_{ic}$ refers to the probability of the rescaled RPE being 1.
+$\beta_0$ refers to the intercept for the mean rescaled RPE which in this case is the Core condition, $b_{0_{j[i],k[i],l[i]}}$ refers to the relevant random intercept for $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine, and $\beta_1\textrm{Condition}_{ic}$ is the difference between the Core and Assisted conditions means.
+Similarly, $\beta_{\phi0}$ refers to the intercept for the precision of the rescaled RPE which in this case is the Core condition, $b_{{\phi0}_{j[i],k[i],l[i]}}$ refers to the relevant random intercept for $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine, and $\beta_{\phi0}\textrm{Condition}_{ic}$ is the difference between the Core and Assisted conditions precision.
 
-Weakly regularising priors were employed for prior sample dataset.
+Weakly regularising priors were employed for previous observational sample dataset.
 These were the default priors in the `ordbetareg` package for all intercept terms which are scaled to the expected response value when all predictors are at their means and use a student $t$ distribution with degrees of freedom $\nu=3$, and variance terms similarly scaled with $\mu=0$.
-We left the flat prior on the fixed effect for condition given we were primarily interested in letting the prior dataset speak for itself in generating a posterior distribution to use as prior in the experimental data model.
-As noted, the prior sample model only included the random effects for location.
+We left the flat prior on the fixed effect for condition given we were primarily interested in letting the previous observational dataset speak for itself in generating a posterior distribution to use as prior in the experimental data model.
+As noted, the previous observational sample model only included the random effects for location.
 The priors were as follows:
@@
 \beta_{\phi0} &\sim \text{Student}\ t(3, 0, 2.5) & \text{Prior for intercept of precision parameter i.e., Assisted condition} \\
-\beta_1,\beta_{\phi1} &\sim \text{Uniform}(\text{lb}=-\infty,u=\infty) & \text{Prior for Condition effects in both location and precision parameters i.e., Core minus Assisted condition} \\
+\beta_1,\beta_{\phi1} &\sim \text{Uniform}(\text{lb}=-\infty,\text{ub}=\infty) & \text{Prior for Condition effects in both location and precision parameters i.e., Core minus Assisted condition} \\
 \sigma_{b_{0_{j}}},\sigma_{b_{\phi0_{j}}} \ &\sim \text{Student}\ t(3, 0, 2.5, \text{lb}=0) & \text{Prior for location variability parameters} \\
@@
 
-We then took posterior parameter estimates from the prior sample model and used these to generate informative priors for the experimental data model which were as follows:
+We then took posterior parameter estimates from the previous observational sample model and used these to generate informative priors for the experimental data model which were as follows:
 
@@
 
-We did not have prior sample data for RPD and so for this outcome we used the same model as above for the RPE outcomes but with wholly default weakly regularising priors.
+We did not have previous observational sample data for RPD and so for this outcome we used the same model as above for the RPE outcomes but with wholly default weakly regularising priors.
 
 # Results
-In the experimental dataset we ended up with `r nrow(data)/2` observations for each outcome across for each of the two conditions.
-Our prior sample for TUL after filtering to the same sample of machines came from `r length(unique(prior_data_tul$id))` members (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`) from `r length(unique(prior_data_tul$location))` locations encompassing `r nrow(prior_data_tul)` observations.
-The distributions of raw data from both the prior sample and experimental datasets can be seen in @fig-raw.
+In the experimental dataset there were a final sample of `r nrow(data)/2` observations for each outcome across for each of the two conditions.
+Our previous observational sample for TUL after filtering to the same sample of machines came from `r length(unique(prior_data_tul$id))` members (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`) from `r length(unique(prior_data_tul$location))` locations encompassing `r nrow(prior_data_tul)` observations.
+The distributions of raw data from both the previous observational sample and experimental datasets can be seen in @fig-raw.
 ```{r}
@@
 #| fig-height: 7
-#| fig-cap: The top two panels show the distributions as histograms of both time under load and session rating of perceived effort in the prior sample of data, and the bottom three panels show the paired responses for time under load in addition to the paired responses and histograms for the rating of perceived effort and discomfort in the current experimental data (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel and horizontal dashed lines in the bottom left panel).
+#| fig-cap: The top two panels show the distributions as histograms of both time under load and session rating of perceived effort in the previous observational sample of data, and the bottom three panels show the paired responses for time under load in addition to the paired responses and histograms for the rating of perceived effort and discomfort in the current experimental data (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel and horizontal dashed lines in the bottom left panel).
 
@@
 ```
-In our prior sample model there was little difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds.
+In our previous observational sample model there was little difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds.
 The estimated TUL for Core sessions was `r preds_prior_tul$estimate[2]+120` [95%QI: `r preds_prior_tul$conf.low[2]+120`, `r preds_prior_tul$conf.high[2]+120`] seconds, for Assisted sessions was `r preds_prior_tul$estimate[1]+120` [95%QI: `r preds_prior_tul$conf.low[1]+120`, `r preds_prior_tul$conf.high[1]+120`] seconds, and the between condition contrast (Core minus Assisted) was `r contrast_prior_tul$estimate` [95%QI: `r contrast_prior_tul$conf.low`, `r contrast_prior_tul$conf.high`] seconds.
@@
 
-Our experimental sample model showed far greater difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds, in addition to the posterior estimates differing considerably from the prior sample estimates in terms of between condition contrasts for TUL and probability of stopping at 120 seconds, Assisted condition TUL, and both conditions probability of stopping at 120 seconds.
+Our experimental sample model showed far greater difference between Core and Assisted sessions in terms of both mean TUL and the probability of stopping at 120 seconds, in addition to the posterior estimates differing considerably from the previous observational sample estimates in terms of between condition contrasts for TUL and probability of stopping at 120 seconds, Assisted condition TUL, and both conditions probability of stopping at 120 seconds.
 The estimated TUL for Core sessions was `r preds_tul$estimate[2]+120` [95%QI: `r preds_tul$conf.low[2]+120`, `r preds_tul$conf.high[2]+120`] seconds, for Assisted sessions was `r preds_tul$estimate[1]+120` [95%QI: `r preds_tul$conf.low[1]+120`, `r preds_tul$conf.high[1]+120`] seconds, and the between condition contrast (Core minus Assisted) was `r contrast_tul$estimate` [95%QI: `r contrast_tul$conf.low`, `r contrast_tul$conf.high`] seconds.
@@
 #| fig-height: 7
-#| fig-cap: The top two panels show the prior (i.e., generated from the prior sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of time under load (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel) and the bottom panels show the probabilities of stopping at 120 seconds.
+#| fig-cap: The top two panels show the prior (i.e., generated from the previous observational sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of time under load (note, the prescribed target time under load range of 90-120 seconds is indicated by the vertical dashed lines in the top left panel) and the bottom panels show the probabilities of stopping at 120 seconds.
 
@@
 ```
-In our prior sample model there was a small difference between Core and Assisted sessions in RPE.
+In our previous observational sample model there was a small difference between Core and Assisted sessions in RPE.
 The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_rpe$estimate)` [95%QI: `r scales::percent(contrast_prior_rpe$conf.low)`, `r scales::percent(contrast_prior_rpe$conf.high)`].
-The $\phi$ parameter for precision of the beta distribution was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition.
 The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].
 
-Our experimental sample model showed a similar magnitude of difference between Core and Assisted sessions in RPE, though the posterior estimates for RPEs were considerably greater compared with the prior sample estimates.
+Our experimental sample model showed a similar magnitude of difference between Core and Assisted sessions in RPE, though the posterior estimates for RPEs were considerably greater compared with the previous observational sample estimates.
 The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_rpe$estimate)` [95%QI: `r scales::percent(contrast_rpe$conf.low)`, `r scales::percent(contrast_rpe$conf.high)`].
-Similarly to the prior data, the $\phi$ parameter for precision of the beta distribution in our experimental sample was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition.
 The $\phi$ parameter for Core sessions was `r preds_rpe_phi$estimate[2]` [95%QI: `r preds_rpe_phi$conf.low[2]`, `r preds_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_rpe_phi$estimate[1]` [95%QI: `r preds_rpe_phi$conf.low[1]`, `r preds_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_rpe_phi$estimate` [95%QI: `r contrast_rpe_phi$conf.low`, `r contrast_rpe_phi$conf.high`].
@@
 The estimated RPD as a percentage for Core sessions was `r preds_discomfort$estimate[2]*10` [95%QI: `r preds_discomfort$conf.low[2]*10`, `r preds_discomfort$conf.high[2]*10`], for Assisted sessions `r preds_discomfort$estimate[1]*10` [95%QI: `r preds_discomfort$conf.low[1]*10`, `r preds_discomfort$conf.high[1]*10`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort$estimate*10` [95%QI: `r contrast_discomfort$conf.low*10`, `r contrast_discomfort$conf.high*10`].
-The $\phi$ parameter for precision of the beta distribution was higher in the Core condition suggesting that RPD reported was more clustered around the estimated RPD values as compared to in the Assisted condition, though the contrast for this was less precise than in the RPE models.
 The $\phi$ parameter for Core sessions was `r preds_discomfort_phi$estimate[2]` [95%QI: `r preds_discomfort_phi$conf.low[2]`, `r preds_discomfort_phi$conf.high[2]`], for Assisted sessions `r preds_discomfort_phi$estimate[1]` [95%QI: `r preds_discomfort_phi$conf.low[1]`, `r preds_discomfort_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort_phi$estimate` [95%QI: `r contrast_discomfort_phi$conf.low`, `r contrast_discomfort_phi$conf.high`].
@@
 #| fig-height: 7
-#| fig-cap: The top two panels show the prior (i.e., generated from the prior sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of rating of perceived effort on the percentage scale, and the bottom panels show posterior disributions only for rating of perceived discomfort transformed back to the arbitrary unit scale (0,10).
+#| fig-cap: The top two panels show the prior (i.e., generated from the previous observational sample) and posterior (i.e., the updated distributions after observing the current experimental data) distributions with median (point) and 95% quantile intervals (error bar) of rating of perceived effort on the percentage scale, and the bottom panels show posterior disributions only for rating of perceived discomfort transformed back to the arbitrary unit scale (0,10).
 
@@
 Our study is a first to offer insights into the role of supervision (i.e., Core vs Assisted sessions) during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
-Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.
-Our key findings were that under real world settings in our prior sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range.
-There was a small difference in RPE reported with and without supervision under both real world and experimental conditions with members reporting higher RPE in Assisted training conditions, though RPE on the percentage points scale was notably higher during the experimental study compared with the prior sample for both Core and Assisted sessions.
+Under standardised training protocol prescription (i.e., single sets of RT using resistance machines, performed to momentary failure using a fixed repetition duration and load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.
+Our key findings were that under real world settings in our previous observational sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range.
+There was a small difference in RPE reported with and without supervision under both real world and experimental conditions with members reporting higher RPE in Assisted training conditions, though RPE on the percentage points scale was notably higher during the experimental study compared with the previous observational sample for both Core and Assisted sessions.
 Lastly, RPD was also higher in the experimental study under supervised Assisted sessions compared to Core sessions.
-These results in general support prior work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+These results in general support previous work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
 
-We will first discuss the findings from the prior sample of Core and Assisted members.
+We will first discuss the findings from the previous observational sample of Core and Assisted members.
 In this *in situ* dataset we find for both Core and Assisted members there is clear evidence of targeting a specific TUL based on the prescribed TUL range of 90-120 seconds.
@@
 This may be due to load progression protocols being more closely followed by supervising exercise scientists as compared to those training unsupervised where, unless using initially heavy loads or in previously well trained persons, most unsupervised trainees do not progress load sufficiently [@steeleAreTraineesLifting2022; @fisherRoleSupervisionResistance2022].
-In future work we plan to explore more thoroughly the causal effects of supervision in both load progression and also isometric strength outcomes collected between Core and Assisted members; but, an initial descriptive examination of load progression using linear-logarithmic growth models known to theoretically describe strength progression with time from RT well [@steeleLongTermTimeCourseStrength2023a; @latellaUsingPowerliftingAthletes2024; @gschneidnerEffectsLengthenedpartialRange2024] suggests that load progressions are typically similar between membership types (see https://osf.io/wcv6g)[^5].
-In fact, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our prior sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort whether being supervised or not and many perhaps just "going through the motions" [@steeleHigherEffortbasedParadigm2017].
+In future work we plan to explore more thoroughly the causal effects of supervision in both load progression and also isometric strength outcomes collected between Core and Assisted members; but, an initial descriptive examination of load progression using linear-logarithmic growth models known to theoretically describe strength progression with time from RT well [@steeleLongTermTimeCourseStrength2023a; @latellaUsingPowerliftingAthletes2024; @gschneidnerEffectsLengthenedpartialRange2024] suggests that load progressions are typically similar between membership types (see https://osf.io/wcv6g)[^6].
+In fact, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our previous observational sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort whether being supervised or not and many perhaps just "going through the motions" [@steeleHigherEffortbasedParadigm2017].
 
-[^5]: Similarly to the other prior historical samples used, we took a sample of 1000 Core and 1000 Assisted members fitting a linear-logarithmic model of time in weeks interacting with membership type (Core vs Assisted) with random intercepts for location, machine, and participant id, and random slopes for log time for each level.
+[^6]: Similarly to the other previous observational historical samples used, we took a sample of 1000 Core and 1000 Assisted members fitting a linear-logarithmic model of time in weeks interacting with membership type (Core vs Assisted) with random intercepts for location, machine, and participant id, and random slopes for log time for each level.
 The model was fit using Restricted Maximum Likelihood estimation with the `lme4` package [@batesFittingLinearMixedeffects2015].
 
-In contrast to the prior sample, our experimental study clearly showed a benefit to TUL performance with supervision suggesting that it might enhance trainee effort.
+In contrast to the previous observational sample, our experimental study clearly showed a benefit to TUL performance with supervision suggesting that it might enhance trainee effort.
 Further, given we recruited existing Core members and had them use their current training loads, it also reinforces that members may be typically selecting loads that are too light and also training with considerable distance from momentary failure.
-Though the prescribed protocol *in situ* in the Kieser Australia clinics is for trainees to perform each exercise to momentary failure, despite the average TUL under Core conditions being similar between prior sample and experimental sample (see @fig-tul), we can see in @fig-raw that many were capable of achieving very long TUL before reaching momentary failure in both the Core and Assisted sessions.
-Interestingly whilst the probability of targeting the 120 second threshold was similar in the prior sample between session types, in the experimental study this was far greater in the Core session condition.
-We speculate that it might be a result of many members, whom were typically not achieving the 120 second TUL upper threshold in their prior sessions, interpreting our explicit instructions in the study to continue to momentary failure as being to ensure that they continue to this upper TUL range target.
+Though the prescribed protocol *in situ* in the Kieser Australia clinics is for trainees to perform each exercise to momentary failure, despite the average TUL under Core conditions being similar between previous observational sample and experimental sample (see @fig-tul), we can see in @fig-raw that many were capable of achieving very long TUL before reaching momentary failure in both the Core and Assisted sessions.
+Interestingly whilst the probability of targeting the 120 second threshold was similar in the previous observational sample between session types, in the experimental study this was far greater in the Core session condition.
+We speculate that it might be a result of many members, whom were typically not achieving the 120 second TUL upper threshold in their previous sessions, interpreting our explicit instructions in the study to continue to momentary failure as being to ensure that they continue to this upper TUL range target.
 Whereas the wide range of TULs, and particularly some of the very high ones, might suggest that others more literally interpreted the instruction to continue to momentary failure.
@@
 
-Of interest is the considerable difference between the *in situ* prior historical sample and the experimental study.
-That under the explicit experimental study instructions to train to momentary failure members were likely to train with a greater effort under both unsupervised and supervised conditions, as evidenced by the TUL results, is reinforced by the greater RPE in percentage points compared to that reported in the prior sample.
-Indeed, this may reflect some degree of the so called "Hawthorne Effect" in both members and the exercise scientists supervising sessions wherein both were more likely to train/supervise training to a higher degree of effort and thus performance whilst under the auspice of the explicit experimental study instructions and observation[^6].
+Of interest is the considerable difference between the *in situ* previous observational historical sample and the experimental study.
+That under the explicit experimental study instructions to train to momentary failure members were likely to train with a greater effort under both unsupervised and supervised conditions, as evidenced by the TUL results, is reinforced by the greater RPE in percentage points compared to that reported in the previous observational sample.
+Indeed, this may reflect some degree of the so called "Hawthorne Effect" in both members and the exercise scientists supervising sessions wherein both were more likely to train/supervise training to a higher degree of effort and thus performance whilst under the auspice of the explicit experimental study instructions and observation[^7].
 Whilst this so called effect is not particularly well understood in its specifics there is general agreement that both participants and researchers react differently under observational settings such as research [@mccambridgeSystematicReviewHawthorne2014; @paradisGoodStoryHawthorne2017] as seems to be observed here.
 
-[^6]: At Kieser Australia the exercise scientists undertake an "exam" / "assessment" after 3-months of being employed and anecdotally we have noted from staff conducting these observations that those exercise scientists will change how they run a session when they are being watched as part of this process.
+[^7]: At Kieser Australia the exercise scientists undertake an "exam" / "assessment" after 3-months of being employed and anecdotally we have noted from staff conducting these observations that those exercise scientists will change how they run a session when they are being watched as part of this process.
 
 The limitations of the present results and conclusions should be noted.
-Firstly, the lack of difference in our prior sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors possibly causal of the TUL during training (member characteristics, load used and load progression).
+Firstly, though anecdotal observations in the clinic generally confirm compliance broadly speaking, we do not have systematically captured data on how accurately the repetition durations prescribed (or other aspects of the prescribed training protocol) are complied with by members.
+Further, the measurement error for the self-recorded or exercise scientist recorded TUL is not known; though the large sample size for the previous observational data, combined with the multiple observations per person in that and our experimental dataset, minimises the impact of any non-systematic measurement error as this would merely affect the precision of estimates.
+Additionally, the protocols were all performed using resistance machines thus minimising one potential element of supervision; technical coaching [@fisherRoleSupervisionResistance2022; @fisherSupervisionStrengthTrainingthe2025].
+Thus the generalisability to other resistance training modalities is unclear.
+The lack of difference in our previous observational sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors possibly causal of the TUL during training (member characteristics, load used and load progression).
 However, the session RPE data does perhaps corroborate the inference that, whilst training with slightly greater effort during Assisted training, members *in situ* may still be training with suboptimal effort compared to what occurs under experimental study conditions when explicitly the instruction to train to momentary failure is reinforced.
-But, a further issue with this is that two separate tools/protocols for capturing RPE were used in the prior sample (session RPE with an adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
+But, a further issue with this is that two separate tools/protocols for capturing RPE were used in the previous observational sample (session RPE with an adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
 [-@steeleDifferentiationPerceivedEffort2017]).
@@
 The adaptation to the 6-20 scale in particular involved changes to verbal anchors and it was noticed in the raw distribution of data (see @fig-raw) that there was an abundance of values of 13 reported.
-Whether or not the tendency for lower RPE in percentage points in the prior sample is necessarily caused by reporting behaviours due to scale construction or due to members actually training with lower effort *in situ* is not clear.
+Whether or not the tendency for lower RPE in percentage points in the previous observational sample is necessarily caused by reporting behaviours due to scale construction or due to members actually training with lower effort *in situ* is not clear.
 It is planned that for future member data collection Kieser Australia will move towards an exercise by exercise approach using adaptations of the Steele et al.
 [-@steeleDifferentiationPerceivedEffort2017] scales.
-The similar degree of contrast between Core and Assisted conditions in both prior sample and experimental study data though does reinforce the role of supervision in influencing trainee effort, or at least reported perception of effort.
+The similar degree of contrast between Core and Assisted conditions in both previous observational sample and experimental study data though does reinforce the role of supervision in influencing trainee effort, or at least reported perception of effort.
 Lastly, despite the impact of supervision on trainee effort, it is still relatively unclear the extent to which it influences adaptation to training.
@@
 There was a small difference in RPE reported with and without supervision under both real world and experimental conditions suggesting that under supervision trainees train with greater proximity to failure, which was also supported by greater RPD under supervision.
-These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+These results in general support previous work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+
+# Conflict of Interest
+All authors work in the health industry providing paid consultancy. During the time of the project TD was employed by Kieser Australia, and JS is at the time of writing employed by Kieser Australia.
 
@@
 We would like to extend our sincere gratitude to all the members and exercise scientists at Kieser Australia without whom this project would not have been possible.
-We would also like to extend our thanks to Myles Moore for his feedback on the manuscript.
+We would also like to extend our thanks to Dr Myles Moore for his feedback on the manuscript.
 

### 2025-05-20 R/functions.R

@@ -786,14 +786,16 @@ make_plot_combined_data <- function(plot1, plot2, plot3, plot4, plot5) {
     (plot1) + theme(strip.text = element_blank()) | (plot2)
   ) +
     plot_annotation(title = "Prior Sample Data Distributions",
-                    subtitle = "Time Under Load and Session Rating of Perceived Effort") +
+                    subtitle = "Time Under Load and Session Rating of Perceived Effort",
+                    tag_levels = "1", tag_prefix = "A") +
     plot_layout(axes = "collect")
   
   data_plots <- (
     (plot3) | (plot4) | (plot5)
   ) +
     plot_annotation(title = "Current Experimental Data Distributions",
-                    subtitle = "Time Under Load, Rating of Perceived Effort, and Rating of Perceived Discomfort") +
+                    subtitle = "Time Under Load, Rating of Perceived Effort, and Rating of Perceived Discomfort",
+                    tag_levels = "1", tag_prefix = "B") +
     plot_layout(axes = "collect_x")
   
   
@@ -976,13 +978,15 @@ make_plot_combined_tul <- function(plot1, plot2, plot3, plot4) {
     (plot1 + guides(fill = "none", color = "none")) | (plot2 + guides(fill = "none", color = "none"))
   ) +
     plot_annotation(title = "Time Under Load",
-                    subtitle = "Global Grand Means for Predictions and Contrasts")
+                    subtitle = "Global Grand Means for Predictions and Contrasts",
+                    tag_levels = "1", tag_prefix = "A")
   
   tul_hu_plots <- (
     (plot3 + guides(fill = "none", color = "none")) | (plot4)
   ) +
     plot_annotation(title = "Stopping at 120 seconds?",
-                    subtitle = "Global Grand Means for Predictions and Contrasts") + 
+                    subtitle = "Global Grand Means for Predictions and Contrasts",
+                    tag_levels = "1", tag_prefix = "B") + 
     plot_layout(guides = "collect") & 
     theme(legend.position = "bottom")
   
@@ -1150,7 +1154,8 @@ make_plot_combined_rpe_discomfort <- function(plot1, plot2, plot3, plot4) {
     (plot1 + guides(fill = "none", color = "none")) | (plot2)
   ) +
     plot_annotation(title = "Rating of Perceived Effort",
[…]

### 2025-05-20 _targets.R

@@ -415,5 +415,24 @@ list(
       w = 7.5,
       h = 5
     )
+  ),
+  
+  # Add flow chart for modelling strategy -----
+  tar_target(
+    flow_chart,
+    make_flow_chart()
+  ),
+  
+  tar_target(
+    flow_chart_tiff,
+    ggsave(
+      "plots/flow_chart.tiff", 
+      flow_chart, 
+      device = "tiff",
+      dpi = 300,
+      w = 10,
+      h = 7.5
+    )
   )
+  
 )

### 2025-05-20 R/functions.R
Make sure flow chart colors are correct

@@ -1266,7 +1266,6 @@ make_plot_loadprog <- function(preds, slopes) {
 make_flow_chart <- function() {
   # Data for nodes
   nodes <- data.frame(
-    id = 1:8,
     label = c(
       "Previous observational data\n(i.e., TUL, session RPE)",
       "Experimental study data\n(i.e., TUL, RPE, RPD)",

### 2025-09-22 pre_print/response_to_reviewers.qmd
Second round of manuscript revisions

@@
 title: "Response to Reviewers"
-subtitle: "The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
+subtitle: "The effects of supervision on effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members"
 format: pdf
@@
 
-Many thanks for considering our manuscript and placing it under review with *Peer Community in Health and Movement Sciences*.
-We are very grateful for your efforts in obtaining reviews.
-We are very impressed with the quality and constructiveness of the feedback provided by the reviewers which we feel has significantly improved the quality of our manuscript.
+Many thanks for considering our manuscript and placing it under further review with *Peer Community in Health and Movement Sciences*.
 
-Note, because of the extensive rewrite of the manuscript due to the shift in focus we have not used track changes or highlighting throughout, though changes can be viewed in the GitHub repository commits and versions.
-However, please see below for our comments in response to some of the key points raised by the reviewers.
+We apologise sincerely for our delay in addressing these and resubmitting.
+Unfortunately other commitments and workloads had prevented us from returning to this.
+We have attempted to address or rebut the additional comments provided by the reviewers.
 
@@
 
-# Recommender and Reviewer Comments and Author Responses
-
-## Recommender - Mathieu Gruet
-
-### Recommender Comment
-
-<!-- The reviewers have, however, identified several concerns that require your attention before the manuscript can proceed further.
-Most notably, the analysis methodology and presentation of results would benefit being slightly more accessible.
-As your research has considerable practical implications for resistance training coaches and strength clinic members, and considering that many of whom may lack the technical expertise to fully comprehend your analytical approach, greater clarity is essential.
-The actual technical complexity may also present challenges for researchers who wish to replicate or build upon your methods in similar or different contexts.
--->
+# Reviewer Comments and Author Responses
 
@@
 
-Minor comments on the writing:
-
-While the paper is overall well written, I found that the paragraphs tended to be excessively lengthy.
-There were also a few awkward and run-on sentences and a few instances of colloquial writing that I recommend be cleaned up, as follows:
-
-Page 3: start a new paragraph at “An acute randomized cross-over design…”
-
-Page 4 and elsewhere: “We ended up with n = 45…” – “ended up” seems colloquial to me.
-
-Page 6: The following sentence is awkward and excessively long: “Anecdotal reports from Kieser Australia staff had also prior to exploring the data suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.”
+Title - Suggest changing “upon” to “on.”
 
@@
 
-Thank you for your kinds words and for the suggestions.
-We have made the minor changes noted on page 3, page 4 and elsewhere (noting instead the "final sample size..."), and change the sentence on page 6 to *"Prior to exploring the data we had received anecdotal reports from Kieser Australia staff.
-These suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure."*
+We have made this change.
 
@@
 
-The repetition tempo and resulting TUL are major components of the resistance exercise prescription, yet the emphasis on these variables is poorly justified.
-Specifically, on page 3, the authors mention that the clients are prescribed repetitions lasting ~ 12 seconds, which consists of a 4:2:4:2 concentric:isometric:eccentric:isometric tempo.
-This tempo prescription is questionable for strength development, wherein powerful concentric motions of maximal effort should be emphasized (Spiering et al.
-2023 J Str Cond Res).
-Pausing at the top of the movement (in the “lockout” position) is likewise questionable.
-
-What effects can be expected from this tempo prescription on TUL compared to more common resistance exercise prescriptions? Would these effects affect the generalizability of this study?
-
-To what extent do the clients comply with the tempo prescriptions?
-
-What is the reliability and accuracy of the TUL recordings?
-The authors stated that “During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose.” For the retrospective analysis, the authors stated that “TUL was also recorded similarly using timers available about the clinic, or for some sessions using a more recently developed mobile phone application (Kieser Konnect, Kieser Australia) which is placed on a stand on the resistance machines and used to track TUL for each exercise in their training card which is programmed to the application.” The self-report nature of this variable could bias the results.
+Abstract - TUL, RPE, and RPD are measures of effort, not exercise performance.
+There are several instances in which “exercise performance” is used in place of “effort,” wherein only the latter should be used.
+Be consistent with the constructs and operational definitions.
 
@@
 
-Unfortunately this is not something that was necessarily chosen and justified from a research design perspective.
-Instead this is just the protocol that Kieser utilise across their facilities.
-Firstly However, we would note that, whilst most would argue that faster movement speeds are likely to maximise neuromuscular adaptations that contribute to overall strength development it is also not conclusive that repetition durations of those used here are necessarily suboptimal and a range of movement speeds have been found to be effective [@wilkInfluenceMovementTempo2021].
-A recent meta-analysis did conclude that intentionally slow repetitions are sub-optimal [@hermesIntentionallySlowConcentric2023] though the dichotomisation of "fast" vs "slow" does limit the interpretation of their results (we are actually in discussions regarding whether to redo this meta-analysis with updated searches and the treatment of repetition duration as a continuous variables with interactions between concentric and eccentric phase durations).
-But more broadly speaking, the magnitude of standardised effect sizes reported in a recent large scale study from our group utilising another dataset [@steeleLongTermTimeCourseStrength2023a] from a commercial training provider who use an even longer repetition duration (~10:10s) were similar to those seen in large scale meta-analytic data we have collected regarding more typical resistance training programs compared with non-training controls [@steeleMetaanalysisVariationSport2023a].
-Thus, it seems unlikely that there would be drastically different effects upon strength even with slower repetition durations.
-Lastly, it is also thought that the intended rather than actual speed of movement may be more important [@behmNarrativeReviewVelocitybased2025].
-We actually intend to more systematically explore this at some point as we are shifting towards a protocol whereby, despite a deliberately slow initial repetition duration, after fatigue accumulates intraset the actual movement speed possible is reduced such that even slow repetition durations can be maintained whilst attempting to move fast.
+In the present study exercise performance as TUL is used as an operationalisation to draw inferences about effort.
+As such, in places where we are drawing broader inferences about the concept of effort as informed by the exercise performance we use "effort", but in others we use more accurately to the operationalisation "exercise performance" or "TUL".
+We have added the following the abstract to clarify this:
 
-Also, we did not necessarily have any specific theoretically driven hypotheses about the impact of supervision with specific manipulations of resistance training variables such as repetition duration.
-So we felt that this ecological dataset and context was suitable for our research questions, particularly given it offered a natural experiment of sorts as Kieser already implements supervised and unsupervised training.
+*"...exercise performance, measured as time under load (TUL) where a longer TUL at a given load is indicative of greater effort..."*
 
-Notably however we do not have data on the extent to which clients actually comply with these prescriptions (though informally from clinical observations the majority do comply with them approximately).
-We have added a comment in the limitations to note this in addition to also the reliability of the TUL recordings.
-The latter we are less concerned with however given that the sample size of the previous data is large (any measurement error due to self-report, assuming no systematic bias, will only affect the error term and thus precision of estimates), and that this is also the case with the experimental dataset when considering the multiple exercises utilised too.
+We have also added to the introduction the text in parentheses at the end of this sentence to clarify this conceptually:
 
-*"Firstly, though anecdotal observations in the clinic generally confirm compliance broadly speaking, we do not have systematically captured data on whether the repetition durations prescribed (or other aspects of the prescribed training protocol) are complied with by members.
-Further, the measurement error for the self-recorded or exercise scientist recorded TUL is not known; though the large sample size for the previous data combined with the multiple observations per person in our experimental dataset minimised any non-systematic measurement error as this would merely affect the precision of estimates."*
+*"Effort is conceptualised at the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance at a given load (i.e., with more repetitions or a longer time under load)."*
+
+And clarified in the methods:
+
+*"The TUL performed (where a longer TUL was interpreted as indicating greater effort), and ratings of perceptions of effort (RPE) and discomfort (RPD) were recorded and compared between both conditions."*
 
@@
 
-Consider justifying the use of Bayesian inference over alternative frequentist methods.
-This justification should be upfront, perhaps in the Introduction or in the Approach to the Problem section of the Methods.
-In this specific case, the Bayesian methods are powerful for studying in an integrated manner the experimental and observational data.
-The authors should better emphasize this point as well as other benefits that the Bayesian approach uniquely afforded.
+Abstract - Second sentence: “We investigated supervision’s role in effort…” – “in” should be “on”
 
@@
 
-We have added the following to the beginning of the statistical methods section:
-
-*"All analyses have been conducted within a Bayesian framework and are primarily focused upon estimation of the relevant parameters of interest from our models [@kruschkeBayesianNewStatistics2018].
-A Bayesian framework was used specifically because it allows for the explicit updating of beliefs regarding prior probable values for parameters of interest with new evidence (i.e., data) to form updated posterior beliefs regarding probable parameter values.
-Given that we had prior empirical data from which we could generate our prior beliefs regarding probable values for parameters of interest, and we had collected new experimental data that could be used to update these, the Bayesian framework is ideally suited to this type of inferential problem.
-Contrastingly, within traditional frequentist inference prior beliefs are not specifically incorporated in this fashion.
-Further, our choice to utilise Bayesian approaches was also based upon the specific modelling choices made and the flexibility of current Bayesian software packages such as those used here to implement such models."*
+We have made this change.
 
@@
 
-Consider including a flowchart for the statistical analysis and then describing each element in the text.
-For example, something that communicates the following: Priors for observational data > posterior distributions for observational data > priors for experimental > posteriors experimental, etc.
-Relate the flowchart elements to the results, i.e., from which points different estimates were obtained.
+The Introduction describes the existing knowledge of the effect of supervision on resistance training outcomes, specifically strength, and summarizes the study’s design and methods.
+While the essential content is present, it could be more clearly discussed.
+I suggest reorganizing the introduction as follows:
+• The first paragraph is good as is. 
+• I would present the hypothesized mechanisms together up front, i.e., trainers enforce load progression and help motivate the exercisers to apply the proper amount of effort (i.e., closer proximity to momentary muscular failure).
+• I would then discuss the evidence for each hypothesized mechanism, which is mostly indirect. 
+• The second-to-last paragraph should summarize the existing gaps in knowledge, i.e., the data concerning supervision effects are still limited, especially in ecological settings, and the potential mechanisms by which supervision might enhance strength are likewise unclear.
+• The final paragraph should more explicitly state the study purpose and along with a hypothesis.
+For example, “The purpose of this study was to examine the effects of supervision on client effort in resistance training in a more ecological setting.
+We hypothesized that … I would then introduce the methods as the authors have done by stating that “We tested the hypothesis by …”
 
@@
 
-This is a great suggestion and we have added a flow chart color coded to hopefully relate the results to the different stages of our modelling strategy.
+We thank the reviewer for their suggested revision to the introduction.
+We have however opted not to make any major rewrites to it as we do feel the major rationale is presented clearly already, and we have not added any hypothesis as our analyses were not aimed at any kind of statistical hypothesis testing but instead exploratory and descriptive parameter estimation.
+Lastly, given the suggested restructure is largely one of preference, this is somewhat of a pragmatic choice on our part to not make changes because of available time for making revisions given other workload demands currently (and that we're already pretty late with making revisions in response to this second round of review - our apologies!).
 
@@
 
-Define all letters and symbols for the non-expert.
-For example, the degrees of freedom for the TUL model is specified as νy, but the “y” is undefined in the manuscript.
-Perhaps knowledge of y is obvious to the expert, but it eluded me.
+Minor comments on the writing: There were several errors and suboptimal wording choices present. 
+• The word “recent” (typically followed by “meta-analysis”) is repetitively used throughout.
+Just state the result and cite the paper.
+For example, the second sentence, “In a recent systematic review and meta-analysis from Fisher et al.
+(2022) there was a moderate standardised mean effect (0.40 [95%CI: 0.06, 0.74]) of supervised vs unsupervised RT on strength outcomes synthesised from ten studies” could be more succinctly stated as “A meta-analysis of 10 studies demonstrated a moderate benefit of supervision on strength outcomes (standardized mean effect = 0.40 [95%CI: 0.06, 0.74]; Fisher et al., 2022).”
+• Error, second paragraph: “Effort is conceptualized at…” – “at” should be “as” 
+• Grammar, third paragraph: “But a caveat is that there was limited data…” – the word “data” is plural, so “was” should be “were.” There are other instances of this error in the manuscript.
 
@@
 
-Apologies, this we realise is not so clear in hindsight, but the $y$ subscript refers to the fact that this is a singular parameter estimated for the whole model.
-We have updated the entire statistical analysis section and included details for each model notation, and also tried to include a brief "lay" description of what the model is in essence doing to try to aid the reader in understanding.
-
-Given this is a scientific manuscript though (and frankly, most researchers and practitioners don't fully read the analyses section most the time anyway...
-and I often question the extent to which many really understand the typical "basic" analyses they are more used to seeing), we have placed this in the online supplementary materials and included a footnote at the beginning of the statistical analysis section that links to this and reads:
-
-*"Note, we thank the reviewers for noting the complexity of the analyses performed in the present study relative to what both researchers and practitioners are used to encountering in the field.
-We have done our best to provide both detailed explanation and justification of the choices made, describe what the models being used are doing, provided mathematical notation for them, and all code for the analyses conducted is available.
-We have also included in the online supplementary material a brief "lay" summary which was generated making use of ChatGPT to assist in developing these explanations.
-We prompted ChatGPT explicitly by providing it with the copied text, including equations, from our quarto manuscript files and asked it to provide a lay description.
-We then edited these.
-These can be seen here <https://osf.io/b9n4v>."*
+We have however made these minor changes as noted for the existing introduction.
 
@@
 
-On page 7, in the priors for γ1 and β1 in the TUL model, it is stated “lb = -60, lb = 60.” Should the second lb be “ub” for upper bound?
+As with the Introduction, the Methods are thoroughly described but could be better organized to highlight critical features.
+In particular, the operational definitions of effort in this study were TUL, RPE, and RPD.
+These operational definitions should be more explicitly defined and justified as measures of effort.
+Currently, their description is an afterthought buried in a paragraph: “The TUL performed, and ratings of perceptions of effort (RPE) and discomfort (RPD) were recorded and compared between both conditions.” Furthermore, whether session RPE or RPE post each exercise is not made clear until the end of the first paragraph in the Protocols section (“In addition, they recorded their RPE and RPD in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions.”) Such critical details about the measures should be mentioned upfront in their own paragraph.
 
@@
 
-Thanks for spotting this. It has been updated.
+As above with the introduction revisions, we have opted not to make further structural changes as we believe that the required information is contained in the methods section.
+As noted above though we have more clearly noted that, given the conceptualisation of effort, the exercise performance as TUL is indicative as an operationalisation of effort in the present study.
 
@@
 
-The results are effectively described and visualized.
-However, the figure captions should begin with a title sentence, followed by the text that was written.
-Panel labels (a, b, c, etc.) should also be included.
+I question the following sentence: “This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.” To me, it seems a bit disingenuous to state the analysis is exploratory when there are meta-analyses of supervision effects on RT outcomes and a clear testable hypothesis that the study addresses using a very detailed Bayesian analysis.
+Why was the experimental intervention not pre-registered?
+Was this a deliberate choice?
+Could the study be retrospectively registered?
+(I have seen others do so.)
 
@@
 
-As the figures already contain titles in the plots themselves we have retained only the caption.
-The purpose of having titles on the figures is to faciliate their reuses by others in presentations etc without having to use the figure caption.
-We have added the panel labels though for each row and column.
+We have tried to be as transparent as possible with this study.
+The original experimental study was conducted several years previously, and before any of the authors were fully cognisant of the value of pre-registration.
+Further, we did not have a sufficiently precise statistical hypothesis to test for which we could for example have performed a power analysis (Frequentist or Bayesian) to help justify sample size for the experimental hypothesis.
+We would prefer not to retrospectively register given this would be merely performative at this stage and does not serve the primary purpose of pre-registration (i.e., the transparent reporting of research to enable the third-party evaluation of the severity of tests for claims made based on hypothesis tests).
+In retrospect we could have looked to convert the standardised effect size metrics to the relevant scales for the parameter estimates presented contrasting between conditions, and then conduct either an analysis akin to examining the overlap with the region of practical equivalence (ROPE), or calculated Bayes Factors for this.
+But we feel this would be disingenuous to add (and also take more time than we have currently) as we never planned to do this a priori.
+As such, having the full dataset open and being transparent about the nature of the study is appropriate in our opinion (notably, these responses will also be publicly available so people can see why we made this choice).
 
@@
 
-Throughout the manuscript but particularly in the Results, the authors use the term “prior sample” to describe the retrospective observational data.
-This terminology might cause confusion because the word “prior” also refers to the prior distributions needed for Bayesian modelling.
-I suggest renaming the sample “observational sample” or some other alternative.
+Experimental approach section: This sentence is run-on: “This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a ~ 12 seconds repetition duration (i.e., ~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension, not “locked-out”, on the involved musculature at the end of each concentric and eccentric muscle action).”
 
@@
 
-We have updated it throughout to instead read "previous observational sample".
-We have also removed the use of "prior" in other instances that might cause confusion e.g., "...prior training...", "...
-prior research..."
+We have split this sentence now.
+
+*"This consists of a single set of the resistance machine exercises prescribed on their current training programme card.
+A load is used that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension, not "locked-out", on the involved musculature at the end of each concentric and eccentric muscle action)."*
 
@@
 
-The phrase “the 𝜙 parameter for precision of the beta distribution in our experimental sample was higher in the Assisted condition suggesting that <variable> reported was more clustered around the estimated <variable> values as compared to in the Core condition” was repeated three times, once for each of the outcome variables.
-While informative, this phrase reflects an interpretation and it is questionable as to whether it belongs in the Results.
-If the authors decide to include it, include it only once because it is repetitive.
+Previous observational sample subsection: The following sentence is run on, contains several awkward passages, and contains punctuation errors: “For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively, took the first training session after at least 6 months of previous training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.” Suggest rewriting as a numbered list of inclusion criteria for the TUL data.
 
@@
 
-We have removed this sentence as it is mentioned that this parameter relates to clustering of responses in the discussion already.
+We have rewritten this as:
+
+*"For TUL we limited data to Core and Assisted members training sessions that either were not, or were, led by an exercise scientist respectively[^3].
+We took the first training session after at least 6 months of previous training at Kieser Australia had been completed by each member.
+Then we randomly sampled 1000 Core and 1000 Assisted members and filtered to the resistance machines used by members in the experimental study."*
 
@@
 
-I proposed above some methodological questions and comments that the authors may wish to address in the revised Discussion.
-Another potential limiter to generalizability of this study is that the exercises are all machine based, and, as the authors noted in their previous meta-analysis, a major potential benefit of supervision is technical coaching (Fisher 2022 Int J Str Cond).
-If participants were to use free weights, they may further “underload” for safety considerations and for not knowing the “proper technique.” I encourage the authors to consider commenting on this aspect.
+Protocols section:  
+• The first paragraph is overlong – a new paragraph could begin at the “Participants were instructed to ensure…”  
+• Typo: “These TUL recordings where then input…” – “where” should be “were”
 
@@
 
-This is a good point and we have added a comment in the limitations regarding this - *"Additionally, the protocols were all performed using resistance machines thus minimising one potential element of supervision; technical coaching [@fisherRoleSupervisionResistance2022; @fisherSupervisionStrengthTrainingthe2025].
-Thus the generalisability to other resistance training modalities is unclear."*
+We have made both of these changes.
 
@@
 
-·       Page 13: The following sentence is awkward and excessively long: “Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.”
+Statistical Analysis:  
+• “All posterior estimates and their precision,…” Precision should be pluralized as “precisions” 
+• Paragraph starting with “For both the previous observational sample…” The authors define “session type” (core or assisted) but then refer to it as “condition” in the model description.
+Suggest helping the reader by using consistent terminology.
 
@@
 
-The sentence itself is quite short, though the additional content in parentheses makes it appear longer.
-Upon re-reading the text in parentheses though we realise that was awkward so have re-written it slightly - *"...(i.e., single sets of RT using resistance machines, performed to momentary failure using a fixed repetition duration and load that should permit momentary failure in a target TUL range of 90-120 seconds)..."* - we have kept it though as it serves as a reminder for the reader of the protocols used when they first beging reading the discussion.
+We have pluralised "precisions" and also added *"...session type (i.e., condition)..."* to clarify the terminology.
+
+### Reviewer Comment
+
+Figure 2 title: “Prior Sample Data Distributions” should be changed to “Previous observational sample histograms”
+
+### Author Response
+
+We have changed the title.
+
+### Reviewer Comment
+
+An issue I sensed with the first version of the paper but failed to articulate in my original comments was the lack of validation of the modeling approach.
+The study features many assumptions and modeling decisions, and the reader is left to wonder about the extent to which the conclusions are robust to them.
+A comment on the sensitivity of the conclusions to the modeling decisions would be appropriate in the Discussion.
+
+### Author Response
+
+Whilst we agree to some extent that all elements of model selection and the assumptions underlying them should be well justified and explained the standards are often double in this regard - researchers in our field using more traditional t-test or ANOVA type analyses (and which probably don't even understand the assumptions underlying them) don't get asked to justify these.
+We've explained quite clearly, and also provided the lay summary for our modelling choices.
+Adding a comment to the discussion about the possible sensitivity of our conclusions to our choice of models is in one sense tautological as our results, as are any using a single modelling approach, conditional on that model (the exception being in the case of Bayesian Model Averaging) but also that making that point leads to the question why not take "multiverse" type approach and try out lots of different models to see what happens.
+In principle this would be great, but as we've noted we don't have time to continue working on this in that fashion.
+If readers are curious then the data is openly available and they are welcome to examine whether a different, perhaps more justifiable, modelling choice would lead to substantively different estimates.
+
+### Reviewer Comment
+
+In addition, diagnostic analyses were performed and included as a supplementary file, which I commend, but they were not mentioned in the Results or Discussion.
+A quick sentence or two summarizing the diagnostics would help assure the reader about the model validities.
+
+### Author Response
+
+We have added to the analysis section where we link to the diagnostic plots:
+
+*"Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged (all were $\hat{R}<1.05$ indicating convergence)..."*
+
+### Reviewer Comment
+
+As I alluded to above, the study purpose could be more explicitly articulated in the first paragraph of the Discussion.
+I suggest modifying the first sentence as follows: “Our study is a first to offer insights into the role of supervision on exerciser effort during RT in an ecologically valid real-world setting.
+These insights were derived from a unique and strong study design involving observational and experimental components.”
+
+I suggest modifying the second sentence as follows to focus on the outcomes measures as operational definitions of effort: “…we assessed effort as the TUL, RPE, and RPD.”
+
+### Author Response
+
+We have changed these sentence to the ones suggested.
+
+### Reviewer Comment
+
+Some paragraphs are overlong. The limitations paragraph, for example, takes up a whole page. It could be broken up into single paragraphs for each of the ~4 limitations discussed by the authors.
+
+### Author Response
+
+We have broken up the limitations paragraph.
 
@@
 
-The aim of the current manuscript was to examine the role of supervision on effort, defined here as time under load (TUL), rating of perceived exertion (RPE), and rating of perceived discomfort (RPD), during resistance training, using both retrospective and prospective analyses.
-The manuscript is well written and structured.
-I would particularly like to highlight the ecological validity of the approach.
-Randomized clinical trials often struggle to replicate real-world settings, so the chosen study design strengthens the external validity of the findings.
+I encourage the authors to revise the conclusion so that it: 
+1. Maintains the statement that under real-world settings there was little difference in exercise performance. 
+2. Clearly states that the observed benefit in the experimental setup was based on an acute effect of a single session. 
+3. Avoids speculation and focuses on direct interpretations of the data.
 
@@
 
-Thank you for the kind words.
+We have made changes to the conclusions to try to more accurately reflect the findings as suggested (and also minor change to the abstract). The conclusion now reads:
 
-### Reviewer Comment
-
-The objective provided at the end of the introduction could be clearer.
-At that point in the text, "effort" is not well defined.
-Since the manuscript later defines effort as TUL, RPE, and RPD, it would be helpful to clarify this earlier in the manuscript.
-
-<!-- ### Author Response -->
-
-We do in the introduction note that, in resistance training, effort is proxied by proximity to momentary failure but have added this more explicitly now.
-*"Effort is conceptualised at the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance."*
-
-### Reviewer Comment
-
-While I understand the logistical reasons given for the sample size in the experimental analysis, but the relatively small sample may limit the generalizability of the findings considering the approach of real-world setting.
-It would be appropriate to include this as a limitation of the study.
-
-### Author Response
-
-We would argue that the sample size is in fact still relatively large compared to most studies in the field and that generalisability is not the primary concern regarding sample size here and we did have participants from across varied locations as well.
-However the first reviewer did not that, despite us having a varied selection of resistance machines enhancing the generalisability of effects in that regard, we did have only machine based exercise and there may be theoretical reasons to suspect our findings may not generalise to other modalities such as free weights.
-We have added this as a limitation.
-
-### Reviewer Comment
-
-Regarding the retrospective analysis, it is unclear why the authors chose two separate samples (one with TUL and another with RPE scores) instead of selecting a sample in which both variables were available.
-Since the aim was to analyze the influence of supervision on effort (as reflected by both TUL and RPE), could separate samples introduce bias (as no confounding factors could be taken into account)?
-I am not sure about this.
-
-### Author Response
-
-We do not in the methods why two separate samples were taken.
-This was due to not all members using the RPE recording function in their training app.
-Given we also did not have any explicit causal model from which to select covariates for adjustment, and that our primary use for the prior samples was to generate informative priors, we felt that the random samples achieved this without having to identify a more selective sample of only those who had recorded RPE.
-
-### Reviewer Comment
-
-The manuscript notes that participants recorded their own TUL using timers visible in the training space.
-While this aligns with the goal of examining real-world settings, it would be helpful to address the validity and reliability of this method.
-Could discrepancies between trainee- and trainer-reported data have affected the results?
-Were trainers generally more rigorous in data collection?
-
-<!-- ### Author Response -->
-
-This is definitely a limitation and one which reviewer 1 also noted.
-We have included this in our limitations section now to make the reader aware of this.
-However, notably from a reliability perspective this would primary influence parameter estimate precision and so the sample sizes, use of prior-posterior updating, enhances our precision here even in the face of such error.
-Also, if we were to speculate, we would think that the trainees were probably more likely to be biased upwards in their TUL recording (having to stop the exercise and then stop the timer on their app...
-though if using the timers in the clinics they can merely see these and note mentally what time they achieved to then record), but yet we still see at least in the experimental data that far greater TUL is acheived with supervision (and even in light of the very strong priors for there being no supervision effect based on the previous sample data).
-
-### Reviewer Comment
-
-How was momentary failure defined and monitored? Was it self-determined by participants in the Core group? If so, how might this have influenced the outcomes?
-
-### Author Response
-
-The fact that it is not explicitly monitored is part of the purpose of the study.
-In both the *in situ* settings of the clinic in terms of the protocols prescribed, and in the experimental study, members are informed that momentary failure is defined as per Steele et al.
-[-@steeleClarityReportingTerminology2017] and that this is the endpoint for their exercises.
-In this study, TUL and RPE offer some insight into the extent to which members are adhering to this.
-For TUL we'd expect, assuming following the load progression rules, some distribution of TUL between 90-120 seconds and not hugely exceeding this nor any spikes.
-For RPE we'd expect them to be providing maximal ratings if they are indeed training to momentary failure.
-Thus these are used as proxies to infer effort and also the effects of supervision upon this.
-We already try to highlight this also in the discussion where we note that the difference in TUL for Core conditions between the previous observational sample and the experimental sample may also express that *in situ* members, despite being prescribed to do so, are not training to momentary failure.
-But in the experimental setting where they knew they were part of a study they trained with a greater degree of effort indicated by greater TUL and RPE.
-
-### Reviewer Comment
-
-Were the Core and Assisted groups in the retrospective analysis performing the same exercises? If different exercises were used, this could have impacted TUL and RPE outcomes.
-
-### Author Response
-
-Exercises varied from member to member in both the previous observational sample and the experimental study data.
-However, this has been factored into the modelling allowing for random effects by exercise/machine in order to produce an overall average treatment effect (ATE) marginalised over these.
-We did not have any specific hypotheses around these though so did not explore them, but the random effects for machine suggested there was very little effect with standard deviation of only ~3-4 seconds TUL between machines.
-
-### Reviewer Comment
-
-For the retrospective data, how exactly was TUL entered into the Kieser database, especially for Core participants? Did participants self-report their TUL using a stopwatch?
-
-### Author Response
-
-We note already in the methods under protocols that the data captured either using the timers (i.e., Swiss engineered clocks with sweep movement for the second hand and positioned so that there is always one in eye line for every machine to enable recording of the TUL in seconds), mobile phone app, or trainers app is input to the training card which is then recorded and stored in the Kieser Australia Database.
-We have added some additional information now including the description of the timers noted above.
-
-### Reviewer Comment
-
-Is comparing only one supervised and one unsupervised session, one week apart, sufficient to draw firm conclusions?
-Given the potential for a Hawthorne effect (i.e., participants modifying behavior due to awareness of being observed), a single-session design may overestimate the true effect of supervision.
-Could repeated sessions have offered a more stable measure?
-
-### Author Response
-
-This is certain to be the case that there will be differential effects over time for the effects of supervision, as would be the case for any training manipulations.
-As noted inn the discussion already, we are in the process of developing longitudinal research to examine this specifically though.
-In fact we are in the process currently of developing the protocols and analysis plans for an interrupted time-series with control design with internal replication in order to examine the effects in our members who have changed contracts over time i.e., went from Core to Assisted and vice versa.
-It is however a complex design and analysis and so, given the already complex approach taken here, we have kept this as a separate piece of follow on work to build further on this area.
-
-### Reviewer Comment
-
-If participants were not blinded to their TUL, is it possible that increased TUL influenced their reported RPE and RPD? Could it influence the outcomes considering the cross-over design?
-
-### Author Response
-
-This certainly could be a confounder for RPE/RPD responses, however this could be a confounder irrespective of whether participants knew their exact TUL or not assuming they have a reasonable perception of time i.e., they could tell if they went longer/shorter than a previous experience they could anchor against and the respond with RPE/RPD as either higher/lower correspondingly.
-But the randomised cross-over design for the experimental study at least allows for any effect of this to be removed.
-
-### Reviewer Comment
-
-The lack of significant differences in TUL in the retrospective analysis contrasts with findings from the experimental condition.
-The authors suggest this may be due to more appropriate load prescription in the Assisted group; however, this interpretation is not supported by the retrospective RPE data.
-If Assisted participants truly trained with a higher (and more optimal) load, one would expect their RPE scores to be notably higher, yet this was not observed.
-
-### Author Response
-
-This interpretation is more speculative and in reference to the Core vs Assisted contrast in the previous observational sample only.
-However, the Core vs Assisted RPE contrasts are actually of similar magnitudes in both the previous observational sample and the experimental data.
-So whilst the absolute magnitudes are still low for the previous sample, the difference between conditions is still similar to that seen in the experimental data.
-So, it is perhaps plausible that more appropriate loads during Assisted compared to Core may be a reason for this even if they are still suboptimal.
-But we do also somewhat test this explanation and note the exploratory examination of some of our load progression data which suggests that load progressions may be similar in the footnote.
-As noted above, we do plan to more carefully examine the longitudinal causal effects of supervision in our members looking at both load progression, and also the isometric strength test data we routinely collect which we believe may better tease out this as a possible explanation for the *in situ* results observed.
-
-### Reviewer Comment
-
-Given the limitations presented by the authors, the results provided by the retrospective analysis, and the effect analyzed during the experimental analysis about the comparison of only one session of supervised versus unsupervised one-week apart, the conclusion that “these results in general support prior work highlighting the importance of supervision during RT” seems overstated.
-While I agree with the authors onthe potential influence of the Hawthorne effect, a more cautious and nuanced interpretation would better reflect the study’s findings and limitations.
-
-### Author Response
-
-Given that we are cautious in our interpretation of the previous observational data results alone, but that the experimental study provides us with causal estimates due to the randomisation and cross-over design, we do not feel that our results are overstated.
-The experimental findings do support the effects of supervision, but the contrast with the seeming lack of difference in TUL, and the smaller absolute RPEs, in *in situ* data suggests that this effect is difficult to causally identify against the background of potential confounders in observational data.
-Again, this is why in our next planned staged of this programme of work we are engaging with more appropriate causal designs such as the interrupted time-series with control design [@lopezbernalUseControlsInterrupted2018].
-
-### Reviewer Comment
-
-No conflict of interest is reported, yet the study analyzes the effect of the Kieser Method, a supervised high-intensity training model developed by the Kieser Training Company, with which two of the authors are affiliated.
-Given this connection, could there be a potential bias in the interpretation or preference for certain results?
-While this does not undermine the value of the study, acknowledging and addressing this possibility would strengthen the transparency and integrity of the manuscript.
-
-### Author Response
-
-Thank you for noting this.
-We apologies for not being more explicit.
-We had mentioned that Kieser had provided in-kind funding for this project and our affiliations as you note are clearly aligned.
-But we have added a section to be more clear regarding this.
-We also hope that our transparency in terms of the data, materials, code, and indeed the results (which to some extent suggest that our in clinic protocols are not in fact being adhered to...
-something we have fed back internally and are looking at how to address and also evaluate prospectively the approaches we take to this) also highlight that we are trying to minimise any affiliational bias as much as possible.
-
-## Reviewer 3 - Anonymous
-
-### Reviewer Comment
-
-I have gone through the study and need the following recommendations for the final decision
-
-### Author Response
-
-Many thanks for your review.
-
-### Reviewer Comment
-
-1. The abstract is well-written and explained, but it would benefit from being divided into distinct sections, such as background, methodology, results, and conclusion.
-
-### Author Response
-
-We have not made this change for now as the journals we may potentially submit to upon recommendation (either Communications in Kinesiology or PeerJ) do not typically require this format.
-If a journal subsequently requires this we will make the change in line with their requirements.
-
-### Reviewer Comment
-
-2. There is only one line about RT. A few more lines at the start of the abstract would be beneficial for the reader.
-
-### Author Response
-
-The abstract is already quite lengthy and hence we have kept this general information to a minimum opting to get straight to summarising the study itself.
-The introduction provides more exposition on the broader topic.
-
-### Reviewer Comment
-
-3.
-The authors used the word "little" in "exercise performance differed little between supervised and unsupervised training." It would be better to use some technical words because it shows subjectivity rather objectivity.
-What is meant by little can be explained numerically.
-
-### Author Response
-
-It is hard to avoid using imprecise descriptors of the magnitude of difference and impossible to avoid making some subjective determination of whether a result is big enough or small enough to care or not care about relative to typical values of the outcome and other background information.
-Though, the manuscript and abstract provide the precise numerical estimates for the values we are admittedly subjectively interpreting here and so any reader can interpret these themselves deciding if they agree with the difference being "little".
-
-### Reviewer Comment
-
-4.The introduction is well written and it contains latest studies. but it should be extended more.
-
-### Author Response
-
-Without suggestion of what could be included to extend it we have not added anything further.
-The manuscript is already quite long and we believe we have provided the background as succinctly as possible in the introduction already and do not require any further detail.
-
-### Reviewer Comment
-
-5.The software used for the study should also be mentioned.
-
-### Author Response
-
-Any specific software used for protocols such as the mobile phone app is already mentioned, and we link out to an extensive list generated using the `grateful` package already containing all of the software used for analysis including version numbers - https://osf.io/ew79g.
-This is also included in the `renv` lockfile such that anyone reproducing our analyses can utilise the exact versions we had utilised.
-
-### Reviewer Comment
-
-6.The result section should be explained for the general readers to understand the technical outcomes. It would be better to make statements simple while explaining technical words.
-
-### Author Response
-
-As suggested by reviewer 1 we have added a lay explanation of the models, and have also added a visual to help with interpretation of the models and results.
-
-### Reviewer Comment
-
-7.The authors mentioned about the confidence intervals and quantile intervals. Are there any more measures to show the strength of outcomes like effect size ?
-
-### Author Response
-
-This is a common misonception in the field; namely that by *effect size* something like a standardised effect size e.g., Cohen's d is meant.
-All of the point estimates and the corresponding interval estimates are in fact effect size estimates in and of them selves and in the units of interest to a reader i.e., seconds, probability of stopping, percentage effort/discomfort.
-So the strength of the outcomes can be interpreted from these directly.
-
-### Reviewer Comment
-
-8.Overall, the study is new and well articulated and would provide reader a quality study.
-
-### Authors Response
-
-Thank you for the kinds words.
-
-### Reviewer Comment
-
-9.Actual real life contribution of this study should be highlighted in the conclusion section. What it can bring in real life?
-
-### Authors Response
-
-We would prefer to leave out inclusion of "real life contribution" such as practical recommendations from this work.
-In fact, the typical inclusion of practical recommendations by many applied journals in our field from a single study always feels unwarranted and results in overstretching of conclusions.
-Neither of the journals we are considering to submit to require practical recommendations of this kind.
-We would note in response though that a very real outcome from this work has for us at Kieser Australia to internally consider how we might adapt our operational procedures to try and ensure that our protocols as prescribed are in fact being followed by members and trainers.
-Our future work will look to evaluate this more closely as we implement changes.
-
-# References
+*"In real world settings there was little difference in exercise performance suggesting similar degrees of effort during both supervised and unsupervised conditions.
+Yet, in our experimental study there was a clear benefit to performance when under supervision suggesting trainees performed with a higher degree of effort.
+There was a small difference in RPE reported with and without supervision under both real world and experimental conditions suggesting that under supervision trainees train with greater proximity to failure, which was also supported by greater RPD under supervision; though in real world conditions the RPE reported suggested that effort may be sub-optimal irrespective of condition.
+These results in general support previous work highlighting the importance of supervision during RT and that under experimental conditions trainees likely train with greater effort when supervised."*

### 2025-09-22 pre_print/response_to_reviewers.tex
Second round of manuscript revisions

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Response to Reviewers},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Response to Reviewers}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{The effects of supervision on effort during resistance
training: A Bayesian analysis of prior data and an experimental study of
private strength clinic members}
\author{}
\date{}

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, breakable, interior hidden, sharp corners, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

Dear Professor Gruet,

Many thanks for considering our manuscript and placing it under further
review with \emph{Peer Community in Health and Movement Sciences}.

We apologise sincerely for our delay in addressing these and
resubmitting. Unfortunately other commitments and workloads had
prevented us from returning to this. We have attempted to address or
rebut the additional comments provided by the reviewers.

Again, we would like to thank all involved in the process of reviewing
this manuscript and look forward to the next round of reviews in
considering our revisions.

Many thanks

The authors

\hypertarget{reviewer-comments-and-author-responses}{%
\section{Reviewer Comments and Author
Responses}\label{reviewer-comments-and-author-responses}}

\hypertarget{reviewer-1---david-clarke}{%
\subsection{Reviewer 1 - David Clarke}\label{reviewer-1---david-clarke}}

\hypertarget{reviewer-comment}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment}}

Title - Suggest changing ``upon'' to ``on.''

\hypertarget{author-response}{%
\subsubsection{Author Response}\label{author-response}}

We have made this change.

\hypertarget{reviewer-comment-1}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-1}}

Abstract - TUL, RPE, and RPD are measures of effort, not exercise
performance. There are several instances in which ``exercise
performance'' is used in place of ``effort,'' wherein only the latter
should be used. Be consistent with the constructs and operational
definitions.

\hypertarget{author-response-1}{%
\subsubsection{Author Response}\label{author-response-1}}

In the present study exercise performance as TUL is used as an
operationalisation to draw inferences about effort. As such, in places
where we are drawing broader inferences about the concept of effort as
informed by the exercise performance we use ``effort'', but in others we
use more accurately to the operationalisation ``exercise performance''
or ``TUL''. We have added the following the abstract to clarify this:

\emph{``\ldots exercise performance, measured as time under load (TUL)
where a longer TUL at a given load is indicative of greater
effort\ldots{}''}

We have also added to the introduction the text in parentheses at the
end of this sentence to clarify this conceptually:

\emph{``Effort is conceptualised at the relation of task demands to the
capacity to meet those demands (Steele 2020) and so in RT is determined
by both the load utilised and also the proximity to momentary failure
due to the fatigue (i.e., reduction in capacity) that occurs with
continued performance at a given load (i.e., with more repetitions or a
longer time under load).''}

And clarified in the methods:

\emph{``The TUL performed (where a longer TUL was interpreted as
indicating greater effort), and ratings of perceptions of effort (RPE)
and discomfort (RPD) were recorded and compared between both
conditions.''}

\hypertarget{reviewer-comment-2}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-2}}

Abstract - Second sentence: ``We investigated supervision's role in
effort\ldots{}'' -- ``in'' should be ``on''

\hypertarget{author-response-2}{%
\subsubsection{Author Response}\label{author-response-2}}

We have made this change.

\hypertarget{reviewer-comment-3}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-3}}

The Introduction describes the existing knowledge of the effect of
supervision on resistance training outcomes, specifically strength, and
summarizes the study's design and methods. While the essential content
is present, it could be more clearly discussed. I suggest reorganizing
the introduction as follows: • The first paragraph is good as is. • I
would present the hypothesized mechanisms together up front, i.e.,
trainers enforce load progression and help motivate the exercisers to
apply the proper amount of effort (i.e., closer proximity to momentary
muscular failure). • I would then discuss the evidence for each
hypothesized mechanism, which is mostly indirect. • The second-to-last
paragraph should summarize the existing gaps in knowledge, i.e., the
data concerning supervision effects are still limited, especially in
ecological settings, and the potential mechanisms by which supervision
might enhance strength are likewise unclear.\\
• The final paragraph should more explicitly state the study purpose and
along with a hypothesis. For example, ``The purpose of this study was to
examine the effects of supervision on client effort in resistance
training in a more ecological setting. We hypothesized that \ldots{} I
would then introduce the methods as the authors have done by stating
that ``We tested the hypothesis by \ldots{}''

\hypertarget{author-response-3}{%
\subsubsection{Author Response}\label{author-response-3}}

We thank the reviewer for their suggested revision to the introduction.
We have however opted not to make any major rewrites to it as we do feel
the major rationale is presented clearly already, and we have not added
any hypothesis as our analyses were not aimed at any kind of statistical
hypothesis testing but instead exploratory and descriptive parameter
estimation. Lastly, given the suggested restructure is largely one of
preference, this is somewhat of a pragmatic choice on our part to not
make changes because of available time for making revisions given other
workload demands currently (and that we're already pretty late with
making revisions in response to this second round of review - our
apologies!).

\hypertarget{reviewer-comment-4}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-4}}

Minor comments on the writing: There were several errors and suboptimal
wording choices present. • The word ``recent'' (typically followed by
``meta-analysis'') is repetitively used throughout. Just state the
result and cite the paper. For example, the second sentence, ``In a
recent systematic review and meta-analysis from Fisher et al.~(2022)
there was a moderate standardised mean effect (0.40 {[}95\%CI: 0.06,
0.74{]}) of supervised vs unsupervised RT on strength outcomes
synthesised from ten studies'' could be more succinctly stated as ``A
meta-analysis of 10 studies demonstrated a moderate benefit of
supervision on strength outcomes (standardized mean effect = 0.40
{[}95\%CI: 0.06, 0.74{]}; Fisher et al., 2022).'' • Error, second
paragraph: ``Effort is conceptualized at\ldots{}'' -- ``at'' should be
``as'' • Grammar, third paragraph: ``But a caveat is that there was
limited data\ldots{}'' -- the word ``data'' is plural, so ``was'' should
be ``were.'' There are other instances of this error in the manuscript.

\hypertarget{author-response-4}{%
\subsubsection{Author Response}\label{author-response-4}}

We have however made these minor changes as noted for the existing
introduction.

\hypertarget{reviewer-comment-5}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-5}}

As with the Introduction, the Methods are thoroughly described but could
be better organized to highlight critical features. In particular, the
operational definitions of effort in this study were TUL, RPE, and RPD.
These operational definitions should be more explicitly defined and
justified as measures of effort. Currently, their description is an
afterthought buried in a paragraph: ``The TUL performed, and ratings of
perceptions of effort (RPE) and discomfort (RPD) were recorded and
compared between both conditions.'' Furthermore, whether session RPE or
RPE post each exercise is not made clear until the end of the first
paragraph in the Protocols section (``In addition, they recorded their
RPE and RPD in that order immediately upon completing the exercise using
previously validated scales for differentiating these perceptions.'')
Such critical details about the measures should be mentioned upfront in
their own paragraph.

\hypertarget{author-response-5}{%
\subsubsection{Author Response}\label{author-response-5}}

As above with the introduction revisions, we have opted not to make
further structural changes as we believe that the required information
is contained in the methods section. As noted above though we have more
clearly noted that, given the conceptualisation of effort, the exercise
performance as TUL is indicative as an operationalisation of effort in
the present study.

\hypertarget{reviewer-comment-6}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-6}}

I question the following sentence: ``This study was not pre-registered
and as described below the sample size was justified based on logistical
concerns and the analysis is considered exploratory.'' To me, it seems a
bit disingenuous to state the analysis is exploratory when there are
meta-analyses of supervision effects on RT outcomes and a clear testable
hypothesis that the study addresses using a very detailed Bayesian
analysis. Why was the experimental intervention not pre-registered? Was
this a deliberate choice? Could the study be retrospectively registered?
(I have seen others do so.)

\hypertarget{author-response-6}{%
\subsubsection{Author Response}\label{author-response-6}}

We have tried to be as transparent as possible with this study. The
original experimental study was conducted several years previously, and
before any of the authors were fully cognisant of the value of
pre-registration. Further, we did not have a sufficiently precise
statistical hypothesis to test for which we could for example have
performed a power analysis (Frequentist or Bayesian) to help justify
sample size for the experimental hypothesis. We would prefer not to
retrospectively register given this would be merely performative at this
stage and does not serve the primary purpose of pre-registration (i.e.,
the transparent reporting of research to enable the third-party
evaluation of the severity of tests for claims made based on hypothesis
tests). In retrospect we could have looked to convert the standardised
effect size metrics to the relevant scales for the parameter estimates
presented contrasting between conditions, and then conduct either an
analysis akin to examining the overlap with the region of practical
equivalence (ROPE), or calculated Bayes Factors for this. But we feel
this would be disingenuous to add (and also take more time than we have
currently) as we never planned to do this a priori. As such, having the
full dataset open and being transparent about the nature of the study is
appropriate in our opinion (notably, these responses will also be
publicly available so people can see why we made this choice).

\hypertarget{reviewer-comment-7}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-7}}

Experimental approach section: This sentence is run-on: ``This consists
of a single set of the resistance machine exercises prescribed on their
current training programme card using a load that should permit them to
reach momentary failure within a time-under-load (TUL) of 90-120 seconds
(though an upper limit of 180 seconds TUL is enforced to avoid machines
being occupied for too long on the clinic floor preventing other members
from using them) using a \textasciitilde{} 12 seconds repetition
duration (i.e., \textasciitilde4:4 seconds concentric:eccentric actions
with a 2 isometric second hold whilst still under load with tension, not
``locked-out'', on the involved musculature at the end of each
concentric and eccentric muscle action).''

\hypertarget{author-response-7}{%
\subsubsection{Author Response}\label{author-response-7}}

We have split this sentence now.

\emph{``This consists of a single set of the resistance machine
exercises prescribed on their current training programme card. A load is
used that should permit them to reach momentary failure within a
time-under-load (TUL) of 90-120 seconds (though an upper limit of 180
seconds TUL is enforced to avoid machines being occupied for too long on
the clinic floor preventing other members from using them) using a
\textasciitilde{} 12 seconds repetition duration (i.e.,
\textasciitilde4:4 seconds concentric:eccentric actions with a 2
isometric second hold whilst still under load with tension,
not''locked-out'', on the involved musculature at the end of each
concentric and eccentric muscle action).''}

\hypertarget{reviewer-comment-8}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-8}}

Previous observational sample subsection: The following sentence is run
on, contains several awkward passages, and contains punctuation errors:
``For TUL we limited data to Core and Assisted members training sessions
that either were not led by an exercise scientist or were respectively,
took the first training session after at least 6 months of previous
training at Kieser Australia had been completed by each member, randomly
sampled 1000 Core and 1000 Assisted members and then filtered to the
resistance machines used in by members in the experimental study so that
we had a selection of members across varied clinic locations and
completing sessions with a selection of resistance machine exercises and
had TUL data for each exercise.'' Suggest rewriting as a numbered list
of inclusion criteria for the TUL data.

\hypertarget{author-response-8}{%
\subsubsection{Author Response}\label{author-response-8}}

We have rewritten this as:

\emph{``For TUL we limited data to Core and Assisted members training
sessions that either were not, or were, led by an exercise scientist
respectively{[}\^{}3{]}. We took the first training session after at
least 6 months of previous training at Kieser Australia had been
completed by each member. Then we randomly sampled 1000 Core and 1000
Assisted members and filtered to the resistance machines used by members
in the experimental study.''}

\hypertarget{reviewer-comment-9}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-9}}

Protocols section:\\
• The first paragraph is overlong -- a new paragraph could begin at the
``Participants were instructed to ensure\ldots{}''\\
• Typo: ``These TUL recordings where then input\ldots{}'' -- ``where''
should be ``were''

\hypertarget{author-response-9}{%
\subsubsection{Author Response}\label{author-response-9}}

We have made both of these changes.

\hypertarget{reviewer-comment-10}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-10}}

Statistical Analysis:\\
• ``All posterior estimates and their precision,\ldots{}'' Precision
should be pluralized as ``precisions'' • Paragraph starting with ``For
both the previous observational sample\ldots{}'' The authors define
``session type'' (core or assisted) but then refer to it as
``condition'' in the model description. Suggest helping the reader by
using consistent terminology.

\hypertarget{author-response-10}{%
\subsubsection{Author Response}\label{author-response-10}}

We have pluralised ``precisions'' and also added \emph{``\ldots session
type (i.e., condition)\ldots{}''} to clarify the terminology.

\hypertarget{reviewer-comment-11}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-11}}

Figure 2 title: ``Prior Sample Data Distributions'' should be changed to
``Previous observational sample histograms''

\hypertarget{author-response-11}{%
\subsubsection{Author Response}\label{author-response-11}}

We have changed the title.

\hypertarget{reviewer-comment-12}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-12}}

An issue I sensed with the first version of the paper but failed to
articulate in my original comments was the lack of validation of the
modeling approach. The study features many assumptions and modeling
decisions, and the reader is left to wonder about the extent to which
the conclusions are robust to them. A comment on the sensitivity of the
conclusions to the modeling decisions would be appropriate in the
Discussion.

\hypertarget{author-response-12}{%
\subsubsection{Author Response}\label{author-response-12}}

Whilst we agree to some extent that all elements of model selection and
the assumptions underlying them should be well justified and explained
the standards are often double in this regard - researchers in our field
using more traditional t-test or ANOVA type analyses (and which probably
don't even understand the assumptions underlying them) don't get asked
to justify these. We've explained quite clearly, and also provided the
lay summary for our modelling choices. Adding a comment to the
discussion about the possible sensitivity of our conclusions to our
choice of models is in one sense tautological as our results, as are any
using a single modelling approach, conditional on that model (the
exception being in the case of Bayesian Model Averaging) but also that
making that point leads to the question why not take ``multiverse'' type
approach and try out lots of different models to see what happens. In
principle this would be great, but as we've noted we don't have time to
continue working on this in that fashion. If readers are curious then
the data is openly available and they are welcome to examine whether a
different, perhaps more justifiable, modelling choice would lead to
substantively different estimates.

\hypertarget{reviewer-comment-13}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-13}}

In addition, diagnostic analyses were performed and included as a
supplementary file, which I commend, but they were not mentioned in the
Results or Discussion. A quick sentence or two summarizing the
diagnostics would help assure the reader about the model validities.

\hypertarget{author-response-13}{%
\subsubsection{Author Response}\label{author-response-13}}

We have added to the analysis section where we link to the diagnostic
plots:

\emph{``Trace plots were produced along with \(\hat{R}\) values to
examine whether chains had converged (all were \(\hat{R}<1.05\)
indicating convergence)\ldots{}''}

\hypertarget{reviewer-comment-14}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-14}}

As I alluded to above, the study purpose could be more explicitly
articulated in the first paragraph of the Discussion. I suggest
modifying the first sentence as follows: ``Our study is a first to offer
insights into the role of supervision on exerciser effort during RT in
an ecologically valid real-world setting. These insights were derived
from a unique and strong study design involving observational and
experimental components.''

I suggest modifying the second sentence as follows to focus on the
outcomes measures as operational definitions of effort: ``\ldots we
assessed effort as the TUL, RPE, and RPD.''

\hypertarget{author-response-14}{%
\subsubsection{Author Response}\label{author-response-14}}

We have changed these sentence to the ones suggested.

\hypertarget{reviewer-comment-15}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-15}}

Some paragraphs are overlong. The limitations paragraph, for example,
takes up a whole page. It could be broken up into single paragraphs for
each of the \textasciitilde4 limitations discussed by the authors.

\hypertarget{author-response-15}{%
\subsubsection{Author Response}\label{author-response-15}}

We have broken up the limitations paragraph.

\hypertarget{reviewer-2---anonymous}{%
\subsection{Reviewer 2 - Anonymous}\label{reviewer-2---anonymous}}

\hypertarget{reviewer-comment-16}{%
\subsubsection{Reviewer Comment}\label{reviewer-comment-16}}

I encourage the authors to revise the conclusion so that it: 1.
Maintains the statement that under real-world settings there was little
difference in exercise performance. 2. Clearly states that the observed
benefit in the experimental setup was based on an acute effect of a
single session. 3. Avoids speculation and focuses on direct
interpretations of the data.

\hypertarget{author-response-16}{%
\subsubsection{Author Response}\label{author-response-16}}

We have made changes to the conclusions to try to more accurately
reflect the findings as suggested (and also minor change to the
abstract). The conclusion now reads:

\emph{``In real world settings there was little difference in exercise
performance suggesting similar degrees of effort during both supervised
and unsupervised conditions. Yet, in our experimental study there was a
clear benefit to performance when under supervision suggesting trainees
performed with a higher degree of effort. There was a small difference
in RPE reported with and without supervision under both real world and
experimental conditions suggesting that under supervision trainees train
with greater proximity to failure, which was also supported by greater
RPD under supervision; though in real world conditions the RPE reported
suggested that effort may be sub-optimal irrespective of condition.
These results in general support previous work highlighting the
importance of supervision during RT and that under experimental
conditions trainees likely train with greater effort when supervised.''}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-steeleWhatPerceptionEffort2020}{}}%
Steele, James. 2020. {``What Is (Perception of) Effort? {Objective} and
Subjective Effort During Attempted Task Performance.''} PsyArXiv.
\url{https://doi.org/10.31234/osf.io/kbyhm}.

\end{CSLReferences}

\end{document}

### 2025-09-22 pre_print/pre_print_v3.qmd
Second round of manuscript revisions

--- 2025-05-20 pre_print/pre_print_v2.qmd
@@
 articletitle: |
-  The effects of supervision upon effort during resistance training: A Bayesian analysis of previous observational data and an experimental study of private strength clinic members
+  The effects of supervision on effort during resistance training: A Bayesian analysis of previous observational data and an experimental study of private strength clinic members
 # shorttitle: "Doe et al. (2022)" # usually auto-generated
@@
   Supervision during resistance training (RT) may enhance strength gains by optimizing trainee effort.
-We investigated supervision’s role in effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership).
-Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort (RPD).
+We investigated supervision’s role on effort during RT in a unique setting with private strength clinics, where members train either unsupervised (“Core” membership) or supervised by a qualified exercise scientist (“Assisted” membership).
+Using both retrospective analysis of member training records and a prospective experimental study, we examined supervision’s impact on exercise performance, measured as time under load (TUL) where a longer TUL at a given load is indicative of greater effort, rating of perceived effort (RPE), and rating of perceived discomfort (RPD).
 Bayesian methods were applied, using empirically informed prior distributions from retrospective data to model the experimental study.
 The previous observational sample included ~1000 members training sessions from each membership type, while the experimental study involved 45 Core members performing both supervised and unsupervised sessions in randomized order, using their current training loads to momentary failure.
-Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training.
+Our findings suggest that, in real-world settings (in situ), exercise performance differed little between supervised and unsupervised training with trainees likely using suboptimal effort under both conditions.
 However, in our experimental study, supervision improved TUL (Core = 125.12 [95%QI: 113.70, 131.90] sec; Assisted = 147.35 [95%QI: 134.29, 154.81] sec; contrast = -22.10 [95%QI: -26.60, -17.61] sec).
@@
 This was further supported by higher RPD during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3]; Assisted = 7.5 [95%QI: 6.5, 8.3]; contrast = -1.2 [95%QI: -1.6, -0.9]).
-Overall, these results reinforce research on the benefits of supervision in RT, indicating that unsupervised trainees—especially in real-world conditions—likely train with suboptimal effort.
+Overall, these results reinforce research on the benefits of supervision in RT, indicating that under experimental conditions trainees likely train with greater effort when supervised.
 license-type: ccby # change if neccessary
@@
 Recent work has highlighted the importance of supervision in optimising strength outcomes from resistance training (RT).
-In a recent systematic review and meta-analysis from Fisher et al.
+In a systematic review and meta-analysis from Fisher et al.
 [-@fisherRoleSupervisionResistance2022] there was a moderate standardised mean effect (0.40 \[95%CI: 0.06, 0.74\]) of supervised vs unsupervised RT on strength outcomes synthesised from ten studies.
@@
 It has been speculated [@fisherRoleSupervisionResistance2022] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort.
-Effort is conceptualised at the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance.
+Effort is conceptualised as the relation of task demands to the capacity to meet those demands [@steeleWhatPerceptionEffort2020] and so in RT is determined by both the load utilised and also the proximity to momentary failure due to the fatigue (i.e., reduction in capacity) that occurs with continued performance at a given load (i.e., with more repetitions or a longer time under load)).
 
-A recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
+A meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022].
 Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., they select \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024].
@@
 
-However, a recent meta-analysis of the dose-response relationship of proximity to momentary failure for strength and hypertrophy outcomes highlights that, while there is increased hypertrophy with closer proximity to failure, there is not a clear relationship for strength [@robinsonExploringDoseResponse2024].
-But a caveat is that there was limited data for proximities to failure \>10 repetitions, and that the models were adjusted for load (average loads were typically \~75-85% of one repetition maximum).
+However, meta-analysis of the dose-response relationship of proximity to momentary failure for strength and hypertrophy outcomes highlights that, while there is increased hypertrophy with closer proximity to failure, there is not a clear relationship for strength [@robinsonExploringDoseResponse2024].
+But a caveat is that there were limited data for proximities to failure \>10 repetitions, and that the models were adjusted for load (average loads were typically \~75-85% of one repetition maximum).
 Thus, the results of this meta-analysis apply to proximity to failure *after* selecting load (i.e., intraset effort) and a higher relative load requires a greater effort all else being equal.
@@
 All Kieser Australia members are prescribed the same protocol.
-This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension, not "locked-out", on the involved musculature at the end of each concentric and eccentric muscle action).
+This consists of a single set of the resistance machine exercises prescribed on their current training programme card.
+A load is used that should permit them to reach momentary failure within a time-under-load (TUL) of 90-120 seconds (though an upper limit of 180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 12 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 2 isometric second hold whilst still under load with tension, not "locked-out", on the involved musculature at the end of each concentric and eccentric muscle action).
 Core trainees are prompted to progress load by \~5% for the next session once a TUL of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist.
@@
 Participants were instructed to complete a single set of the resistance machine exercises on their current training programme card using their current training loads to momentary failure.
-The TUL performed, and ratings of perceptions of effort (RPE) and discomfort (RPD) were recorded and compared between both conditions.
+The TUL performed (where a longer TUL was interpreted as indicating greater effort), and ratings of perceptions of effort (RPE) and discomfort (RPD) were recorded and compared between both conditions.
 This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.
@@
 We opted to generate reasonably large samples to ensure precise prior estimates, though did not use the entirety of the historical data (\~50000 members, \>10 years of data) so as to reduce computation time for Bayesian modelling.
-For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of previous training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise.
+For TUL we limited data to Core and Assisted members training sessions that either were not, or were, led by an exercise scientist respectively[^3].
+We took the first training session after at least 6 months of previous training at Kieser Australia had been completed by each member.
+Then we randomly sampled 1000 Core and 1000 Assisted members and filtered to the resistance machines used by members in the experimental study.
 Thus our final sample size was n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`).
@@
 In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card before participation in the study.
+
 Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the TUL achieved for the purposes of the study i.e., if they realised during the set that their previous load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL acheived.
@@
 During each session for each exercise the trainees/trainers recorded their TUL achieved as they would do for their usual sessions using standard timers situated around the clinic in view during training specifically for this purpose (these are essentially Swiss engineered clocks with sweep movement for the second hand and positioned so that there is always one in eye line for every machine to enable recording of the TUL in seconds).
-These TUL recordings where then input to their training cards.
+These TUL recordings were then input to their training cards.
 In addition, they recorded their RPE and RPD in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).
@@
 
-All posterior estimates and their precision, along with conclusions based upon them, are interpreted continuously and probabilistically, considering priors, data quality, and all within the context of each outcome and the assumptions of the model employed as the estimator [@kruschkeBayesianNewStatistics2018].
+All posterior estimates and their precisions, along with conclusions based upon them, are interpreted continuously and probabilistically, considering priors, data quality, and all within the context of each outcome and the assumptions of the model employed as the estimator [@kruschkeBayesianNewStatistics2018].
 All models were run with 2000 warmup and 2000 sampling iterations and four Monte Carlo Markov Chains.
-Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged, and posterior predictive checks for each model were also examined to understand the model implied distributions (see <https://osf.io/u7g8c>.
+Trace plots were produced along with $\hat{R}$ values to examine whether chains had converged (all were $\hat{R}<1.05$ indicating convergence), and posterior predictive checks for each model were also examined to understand the model implied distributions (see <https://osf.io/u7g8c>).
 We fit two sets of models for the TUL and RPE outcomes, one on the previous observational sample of data using weakly regularising priors in order to generate posterior distributions to inform the priors used in the other model using the experimental data, and a single model on the experimental data for the RPD outcome using weakly regularising priors.
@@
 Upon inspecting the raw distribution of the previous observational sample of data for TUL we noticed that there was a spike at 120 seconds concomitant with the top of the target TUL range prescribed to members under both Core and Assisted conditions (see @fig-raw).
-Before to exploring the data we had received anecdotal reports from Kieser Australia staff.
+Before exploring the data we had received anecdotal reports from Kieser Australia staff.
 These suggested to us that, despite the prescription to train to momentary failure, many Core members were instead selecting loads and training only to the upper 120 second TUL range threshold irrespective of proximity to failure.
@@
 
-For both the previous observational sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept.
+For both the previous observational sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type (i.e., condition), Core or Assisted, where Assisted was coded as the intercept.
 The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used (note, due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset).
@@
 # Discussion
-Our study is a first to offer insights into the role of supervision (i.e., Core vs Assisted sessions) during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
-Under standardised training protocol prescription (i.e., single sets of RT using resistance machines, performed to momentary failure using a fixed repetition duration and load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and RPD.
+Our study is a first to offer insights into the role of supervision on exerciser effort during RT in an ecologically valid real-world setting.
+These insights were derived from a unique and strong study design involving observational and experimental components.
+Under standardised training protocol prescription (i.e., single sets of RT using resistance machines, performed to momentary failure using a fixed repetition duration and load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined effort as TUL, RPE, and RPD.
 Our key findings were that under real world settings in our previous observational sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range.
@@
 Thus the generalisability to other resistance training modalities is unclear.
+
 The lack of difference in our previous observational sample of members regarding TUL does not necessarily indicate that even during Assisted training sessions members are training sub-optimally as this is a comparison unadjusted for other factors possibly causal of the TUL during training (member characteristics, load used and load progression).
 However, the session RPE data does perhaps corroborate the inference that, whilst training with slightly greater effort during Assisted training, members *in situ* may still be training with suboptimal effort compared to what occurs under experimental study conditions when explicitly the instruction to train to momentary failure is reinforced.
+
 But, a further issue with this is that two separate tools/protocols for capturing RPE were used in the previous observational sample (session RPE with an adapted Borg 6-20 scale) vs the experimental study (scales taken from  Steele et al.
@@
 The similar degree of contrast between Core and Assisted conditions in both previous observational sample and experimental study data though does reinforce the role of supervision in influencing trainee effort, or at least reported perception of effort.
+
 Lastly, despite the impact of supervision on trainee effort, it is still relatively unclear the extent to which it influences adaptation to training.
@@
 # Conclusions
-To our knowledge this study is the first examine the role of supervision during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study.
-Under real world settings there was little difference in exercise performance, yet in our experimental study there was a clear benefit to performance when under supervision.
-There was a small difference in RPE reported with and without supervision under both real world and experimental conditions suggesting that under supervision trainees train with greater proximity to failure, which was also supported by greater RPD under supervision.
-These results in general support previous work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
+In real world settings there was little difference in exercise performance suggesting similar degrees of effort during both supervised and unsupervised conditions.
+Yet, in our experimental study there was a clear benefit to performance when under supervision suggesting trainees performed with a higher degree of effort.
+There was a small difference in RPE reported with and without supervision under both real world and experimental conditions suggesting that under supervision trainees train with greater proximity to failure, which was also supported by greater RPD under supervision; though in real world conditions the RPE reported suggested that effort may be sub-optimal irrespective of condition.
+These results in general support previous work highlighting the importance of supervision during RT and that under experimental conditions trainees likely train with greater effort when supervised.
 
 # Conflict of Interest
-All authors work in the health industry providing paid consultancy. During the time of the project TD was employed by Kieser Australia, and JS is at the time of writing employed by Kieser Australia.
+All authors work in the health industry providing paid consultancy.
+During the time of the project TD was employed by Kieser Australia, and JS is at the time of writing employed by Kieser Australia.
+JS is also a recommender for both Peer Community in Health and Movement Sciences (where this manuscript is under review), and Peer Community in Registered Reports.
 

### 2025-09-22 R/functions.R

@@ -785,7 +785,7 @@ make_plot_combined_data <- function(plot1, plot2, plot3, plot4, plot5) {
   prior_plots <- (
     (plot1) + theme(strip.text = element_blank()) | (plot2)
   ) +
-    plot_annotation(title = "Prior Sample Data Distributions",
+    plot_annotation(title = "Previous Observational Sample Data Distributions",
                     subtitle = "Time Under Load and Session Rating of Perceived Effort",
                     tag_levels = "1", tag_prefix = "A") +
     plot_layout(axes = "collect")
